### 2024-03-22 paper/datasheet_template.qmd
adding starter code files

---
title: "Datasheet for Air quality index by municipality, Leading causes of death
and Average PM2.5 concentrated data dataset"
author: 
  - Vanshika Vanshika
  - Shivank Goel
  - Navya Hooda
thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This datasheet documents the datasets used in a study analyzing the impact of air pollution on health outcomes in Alberta, Canada. The data for the study is sourced from the Alberta government's open data portal. It spans several years and includes variables like AQHI readings, PM2.5 concentrations, and mortality rates linked to specific health conditions. This dataset is helpful for its concise representation to study the environmental health in Alberta. It offers insights into public health implications of air pollution. The data cleaning and processing done using tools like R, ensures its reliability and accuracy. This datasheet is an essential guide to the dataset's structure, collection process, and potential applications."

format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*

The dataset was created to investigate the correlation between air quality, particularly PM2.5 levels, and mortality rates due to respiratory and cardiac illnesses in Alberta. The aim was to understand how air pollution impacts public health in this region.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*

The datasets were created by the Alberta Government, specifically by Jobs, Economy and Northern Development, Service Alberta, and Environment and Parks departments.

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*

The creation of these datasets was likely funded by the Alberta government. No specific grant details were provided.

4. *Any other comments?*

These datasets provide critical insights into environmental and public health concerns in Alberta and are vital for ongoing research and policy development in these areas.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*

The instances represent various types of environmental and public health data. The first dataset details the Air Quality Health Indices by municipality in Alberta. The second dataset provides information on the leading causes of death in Alberta. The third dataset offers data on PM2.5 concentrations in Alberta.

2. *How many instances are there in total (of each type, if appropriate)?*

The exact number of instances is not specified, but each dataset covers multiple years of data across various locations in Alberta.

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*

The datasets seem to be comprehensive for their respective scopes, covering broadly geographical areas within Alberta. It seems to be representative of the larger set of environmental and health data within the province.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*

Each dataset has specific labels or targets. For the air quality index, the labels include AQHI values and municipal regions. The causes of death dataset labels include types of illnesses and mortality counts. The PM2.5 dataset's labels consist of concentration levels, locations, and time periods

5. *Is there a label or target associated with each instance? If so, please provide a description.*

Each dataset has specific labels or targets. For the air quality index, the labels include AQHI values and municipal regions. The causes of death dataset labels include types of illnesses and mortality counts. The PM2.5 dataset's labels consist of concentration levels, locations, and time periods.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*

There is no specific mention of missing information. However, as with most large datasets, there may be gaps due to data unavailability or limitations in data collection.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*

The datasets do not explicitly mention relationships between individual instances, but they can be inferred. For example, higher AQHI and PM2.5 values may correlate with increased mortality rates from certain diseases.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*

There are no specific recommended data splits mentioned. Data splits would depend on the specific research or analysis goals.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*

Specific errors or sources of noise are not mentioned. However, as with any environmental data, there may be variabilities and inconsistencies due to natural fluctuations and measurement limitations.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*

The datasets are self-contained and available through the Alberta open data portal. They are subject to the Open Government Licence - Alberta, which allows for reuse and distribution.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*

No confidential or personally identifiable information is included in the dataset.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*

There is no content in these datasets that would be considered offensive or insulting. The data is factual, focusing on environmental and health statistics.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*

The datasets do not specify sub-populations like age or gender. They are more focused on geographical and temporal aspects of air quality and health outcomes.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*

Individual identification is not possible with these datasets as they deal with aggregated data at the municipal level or higher.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*

The datasets do not contain sensitive personal data. They are focused on environmental measures and aggregated health outcomes.

16. *Any other comments?*
	N/A

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*

The data were directly observed and collected through environmental monitoring stations and health records. They are likely verified through standard government data validation processes, though specific details are not provided.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*

The air quality and PM2.5 data were likely collected using environmental monitoring stations equipped with sensors. The causes of death data were probably compiled from health records.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*

The datasets seem to be comprehensive collections rather than samples. They likely  represent the entire population of interest (i.e., all municipalities in Alberta for air quality and all recorded causes of death).

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*

Data collection was likely conducted by government employees or contractors. Specific details about their compensation are not provided.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*

The data collection covers several years, as indicated by the datasets (e.g., 2001-2021 for the leading causes of death).

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*

There is no specific mention of ethical review processes.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*

The data were collected from environmental monitoring stations and health records, not directly from individuals. It is a secondary data source.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*

Individual notification is not applicable, as the data do not pertain to individual persons.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*

Individual consent is not relevant for this type of data, as it does not involve personal data collection from individuals.

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*

As individual consent was not required for this data collection.

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*

There is no information provided regarding an impact analysis on data subjects, likely because the datasets deal with  environmental and health data and not individual-specific data.

12. *Any other comments?*

The data collection process is extensive and critical for understanding the environmental and public health in Alberta. The absence of personal data reduces privacy concerns.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*

There is likely some level of preprocessing and cleaning involved, ensuring consistency, and possibly dealing with missing values. However, specific details are not provided in the information given.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*

There is no mention of whether the raw data is saved alongside the processed data.

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*

Details about the specific software or tools used for preprocessing, cleaning, or labeling are not provided.

4. *Any other comments?*

Proper data preprocessing and cleaning are essential for the accuracy and reliability of analysis results. The datasets, as released, should be ready for use in various analyses and research.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*

The datasets have likely been used for academic research and study before. They have been used to analyze the relationship between mortality rates and causes of certain diseases.

2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*

No specific repository is mentioned, but the datasets are available through the Alberta open data portal.

3. *What (other) tasks could the dataset be used for?*

The data could be used for further environmental health studies, public health planning, and educational purposes.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*

The users should consider potential biases and reporting in data collection.Also they should be aware of geographic
coverage to avoid misinterpretations.

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*

The data should  not be used for identifying or targeting specific individuals.

6. *Any other comments?*

The datasets are valuable for understanding trends and correlations in environmental and public health. They should be used appropriately, considering their scope and limitations.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*

The datasets are publicly available and can be accessed by third parties. They are distributed under the Open Government Licence - Alberta, which allows for use and distribution by anyone.

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*

The datasets are available for download through the Alberta open data portal. There is no mention of a DOI or distribution through other platforms like GitHub.

3. *When will the dataset be distributed?*

The datasets are already available for public access.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*

The datasets are distributed under the Open Government Licence - Alberta, which permits a wide range of uses. The license details can be found on the Alberta open data portal.

5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*

There is no mention of any third-party IP restrictions on these datasets.

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*

There are no specified export controls or regulatory restrictions beyond the Open Government Licence - Alberta.

7. *Any other comments?*

The open access nature of these datasets makes them a valuable resource for a wide range of users, from researchers to policymakers.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*

The datasets are maintained by the respective departments of the Alberta government – Jobs, Economy and Northern Development, Service Alberta, and Environment and Parks.

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*

Contact information is likely available through the Alberta government’s open data portal.

3. *Is there an erratum? If so, please provide a link or other access point.*

There is no mention of an erratum.

4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*

The datasets are updated annually. Specific communication channels for updates are not mentioned.

5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*

There is no specific information about the maintenance of older versions of the dataset.

6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*

This is not applicable as the datasets do not contain individual-level data.

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*

There is no mentioned mechanism for external contributions to the dataset. The data are sourced from government monitoring and records, so external contributions might not be applicable.

8. *Any other comments?*

Regular maintenance and updates are crucial for the continued relevance and accuracy of these datasets.

### 2024-03-22 paper/paper.qmd
adding starter code files

---
title: "Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air pollution"
subtitle: "Based on Data Collected From The Province Of Alberta"
author: 
  - Vanshika Vanshika
  - Shivank Goel
  - Navya Hooda
thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(readxl)
library(here)

deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
data <- read.csv(here("data/analysis_data/cleaned_chart_data.csv"))
data_peak <- read.csv(here("data/analysis_data/cleaned_peak_data.csv"))
data_lung <- read.csv(here("data/analysis_data/merged_data.csv"))
data_heart <- read.csv(here("data/analysis_data/merged_heart_data.csv"))
```

# Introduction

The impact of air pollution on human health has increasingly become a global concern, with respiratory illnesses being a significant consequence of poor air quality. The National Library of Medicine reports that approximately 4 million people die prematurely each year from chronic respiratory diseases linked to air pollution @ncbi. 
The World Health Organization also suggests that air pollution poses a risk for all-cause mortality and specific diseases, especially the exposure to air pollution is strongly linked with outcomes such as stroke, ischemic heart disease, chronic obstructive pulmonary disease, lung cancer, pneumonia, and cataracts @who.

The pollutants most significantly concerning for public health, are particulate matter (PM), carbon monoxide (CO), ozone (O3), nitrogen dioxide (NO2), and sulfur dioxide (SO2). Of these, fine particulate matter is particularly worrisome due to its ability to penetrate the lungs deeply, enter the bloodstream, and reach organs, leading to systemic damage to tissues and cells.

Specifically, PM2.5 is of greater concern due to its harmful properties and dangers. PM2.5 is airborne particulate matter (PM2.5) and is not a single pollutant but rather a mixture of many chemical species and aerosols. PM2.5 is associated with the greatest proportion of adverse health effects related to air pollution, both in the United States and worldwide. Long-term (months to years) exposure to PM2.5 has been linked to premature death, particularly in people who have chronic heart or lung diseases, and reduced lung function growth in children @pm.

In this paper, we analyze data from Alberta, Canada, and aim to explore the estimand of the number of deaths that can be correlated to the Air Quality Health Index (AQHI) and PM2.5 quantities. Specifically, we assess how AQHI relates to the prevalence of respiratory and cardiac illnesses in this specific region, focusing on four major types: Chronic Obstructive Pulmonary Disease, Ischemic Heart Disease, Acute Myocardial Infarction, and Lung Cancer. We use the mortality rate in Alberta data to select respiratory-related illnesses and air quality health index data for Alberta through their provincial open data portal. To analyze respiratory illnesses, we use air pollution measured through the AQHI and specifically the particulate matter 2.5 (PM2.5) air pollutant as a predictor of mortality related to respiratory illnesses. In total, we will be assessing the overall relationship between AQHI and respiratory illnesses, the relationship between PM2.5 and respiratory illnesses, as well as the relevance of PM2.5 levels in lung and heart disease values as a means to draw any notable significance of PM2.5 in different types of illnesses.

Using negative binomial regression, this study seeks to uncover trends and correlations between AQHI and the prevalence of respiratory illnesses. Negative binomial regression is a type of generalized linear model (GLM) that is specifically designed to model count data. In our paper, we are using it to predict the number of deaths, based on the year and illness, for the pollutants present in the air.
By analyzing this relationship, we provide valuable insights that can inform policymakers, healthcare professionals, and the public about the impact of air pollution on respiratory health in Alberta. This research aims to contribute to a better understanding of the health effects of air pollution and to support the development of targeted strategies for air quality improvement and public health protection in Alberta. Our analytical framework explores whether the variables assessed seem to correlate, if at all, to the total deaths in respiratory-related illnesses through 2012-2022.

This paper is organized as follows: In the Data section, we outline the sources of three different datasets, detail the data-cleaning processes applied to each dataset, and describe any data merging procedures used to prepare the data for input into various models. The Results section focuses on analyzing trends and correlations between AQHI, PM2.5, and respiratory and cardiac illnesses. Additionally, we discuss the trends and patterns identified by our model, along with the correlation analysis between these variables. In the Discussion section, we present our overall findings, discuss any biases and weaknesses in the data that may have influenced these findings, and explain our approach to analyzing these limitations.

Overall, this research has the potential to inform public health strategies and interventions aimed at reducing respiratory illnesses and improving overall health in Alberta and beyond.

# Data {#sec-data}

## Data Source and Collection:

The study relies on datasets obtained from the provincial open databases of Alberta, accessible through the official website @alberta. Three key datasets were utilized to extract relevant variables for analysis, aiming to uncover the relationship between air quality and mortality rates in Alberta.
The analysis begins with the leading causes of death dataset for Alberta, sourced from the provincial open data portal @deaths. This dataset provides insights into mortality rates associated with various illnesses, facilitating the examination of trends related to respiratory and heart-related illnesses.
To explore potential correlations between air quality and mortality rates, the study incorporates the Air Quality Health Index (AQHI) dataset for Alberta, sourced from the provincial open data portal @air. This dataset offers comprehensive information on the AQHI across different municipalities in Alberta over multiple years.
Additionally, the study utilizes PM2.5 air pollutant concentration level data sourced from Alberta's official resources @pollution. This dataset provides detailed information on the concentration levels of PM2.5 pollutants over several years, offering valuable insights into air quality trends.
The following subsections outline the sources, collection methodologies, and data-cleaning procedures implemented to ensure the accuracy and reliability of the datasets used in the analysis. This meticulous approach ensures that the data is prepared for thorough analysis, facilitating the exploration of correlations between air quality indicators and mortality rates in Alberta.

Leading Causes of Death in Alberta Data: The disease data is found from the government of Alberta’s open data portal, and was last updated on September 22, 2023 and continues to be updated annually.  This dataset encompasses mortality data related to the top 30 common causes of death. It reports on types of diseases, causes of death, mortality denoted by total death counts, and ranking for 2000-2022.  Due to our focus on respiratory illnesses, in the leading cause of death dataset, we grouped diseases by categories. Our category of focus included filtering on illnesses like acute myocardial infarction, malignant neoplasms of the trachea, bronchus, and lung, other chronic obstructive pulmonary disease, and all other forms of chronic ischemic heart disease. Leading causes of death are measured and ranked by the top 30 most common death causes each specific year. The causes of death are classified based on the International Classification of Diseases 10th Edition.

AQHI Data: The second dataset we used is the air quality health index (AQHI) dataset found at the government of Alberta’s open data portal. This dataset contains AQHI by municipality for the years 2012-2022 and reports air quality health index, and health risk both quantitatively and qualitatively.  To use the AQHI dataset we employed simple data-cleaning practice to maintain descriptive variable names and readability. The data is measured by the percentage of hours for each year at a given air quality level, by municipality. The Air Quality Health Index is calculated based on the relative risks of a combination of common air pollutants that is known to harm human health. These pollutants are ozone (O3) at ground level, particulate matter (PM2.5), and nitrogen dioxide (NO2). Risks are defined as follows: 1-3 High Quality; 4-6 Moderate Quality; 7-9 Low Quality; 10+ Very Low Quality.

PM2.5 Data: We used the PM2.5 data set retrieved from Alberta.ca (Government of Alberta) which was last updated in April 2023. It reports on average PM2.5 concentration levels through the years 2000-2021 using a provincial average, the 10th percentile quantities, and the 90th percentile quantities, with a focus on 8 municipalities Edmonton, Fort McMurray, Grande Prairie, Lethbridge, Medicine Hat, and Red Deer respectively and lastly reports the Canadian Ambient Air Quality Standard (CAAQS) value.

The Alberta Air Zone report @report, which is linked to our dataset, provides a detailed explanation of the measurement and processing of the PM2.5 quantity. Alberta Air Zones divides Alberta into six air zones which are aligned with Alberta’s Land-use Framework regional boundaries. Ambient air quality in Alberta is monitored at continuous air monitoring stations located within these air zones. PM2.5 quantities are taken throughout these stations across Alberta, and they measure the quantities in µg (micrograms per cubic meter of air).

## Data Cleaning

We used R [@citeR] and @rohan for data cleaning and processing, utilizing packages like tidyverse [@tidy] for data manipulation and janitor [@jan] for cleaning column names. Other packages used includes `ggplot2` [@ggplot], `dplyr` [@dp], `readr` [@read], `tibble` [@tib], `janitor` [@jan],`reshape2` [@reshape], `knitr` [@knit], `ggbeeswarm` [@gg], `ggrepel` [@repel], `kableExtra`[@kable], `readxl`[@readxl], `MASS`[@mass], `rstanarm`[@rstan], `modelsummary`[@model] and `here` [@here].

The raw air quality data were preprocessed to remove inconsistencies and irrelevant information. Specifically, we filtered the dataset to include observations from the years 2012 to 2021, which are relevant to our analysis. Additionally, we merged this dataset with additional information on peak pollution levels for comprehensive analysis. 
Similar to the previous datasets, the raw mortality data underwent cleaning procedures to focus on specific causes relevant to our analysis. We filtered the dataset to include observations up to 2021 and merged it with additional information on air quality for correlation analysis. 
The raw data on AQHI were filtered to include observations from the years 2011 to 2021 for consistency with other datasets. Additionally, the data were aggregated at the municipal level for further analysis.

## Data Modifications

In this study, we constructed unique datasets by thoughtfully selecting and merging data from the Government of Alberta's open data portal, Alberta.ca, spanning the years 2012 to 2022. Our process involved merging variables from various datasets to create specific datasets tailored for model building and analysis.
One such dataset, 'cleaned_chart_data,' was created by merging variables such as causes of death, total deaths, provincial average PM2.5 levels, and CAAQS. This dataset was designed to facilitate our analysis of any significant correlations between these variables. A snapshot of this data is referenced in Table X. 
Additionally, we derived two other datasets, 'merged_data' and 'merged_heart_data,' by merging variables related to heart disease numbers, lung disease numbers, and provincial average PM2.5 values. These datasets were instrumental in examining the impact of PM2.5 on each type of illness, as previously discussed.
Overall, our methodology ensured the creation of comprehensive datasets that allowed for a detailed investigation into the relationships between PM2.5 levels and various health outcomes in Alberta.

```{r}
#| label: fig-all
#| fig-cap: Provincial Average PM2.5 Quantity Vs Total Deaths
#| echo: false
#| warning: false

# Plot the regression graph
ggplot(data, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average PM2.5", y = "Total Deaths") +
  theme_minimal()

```

```{r}
#| label: fig-lung
#| fig-cap: Lung Related Causes Mortality Rates vs Average PM2.5 Quantity 
#| echo: false
#| warning: false

# Plot the regression graph for all lungs causes
ggplot(data_lung, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths") +
  theme_minimal()
```

```{r}
#| label: fig-heart
#| fig-cap: Heart Related Causes Mortality Rates vs Average PM2.5 Quantity 
#| echo: false
#| warning: false
# Plot the regression graph for all heart causes
ggplot(data, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths") +
  theme_minimal()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-annual
#| fig-cap: Annual Trends of PM2.5 Concentrations in Alberta
#| echo: false
#| warning: false
#| message: false

# Assuming your data frame is named 'cleaned_avg_PM25_conc' and is already filtered and renamed.
# Convert factors to character to avoid issues in ggplot
data[] <- lapply(data, function(x) if(is.factor(x)) as.character(x) else x)
# Creating the plot with ggplot
p <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = provincial_average, colour = "Provincial Average")) +
  geom_line(aes(y = x10th_percentile, colour = "10th Percentile")) +
  geom_line(aes(y = x90th_percentile, colour = "90th Percentile")) +
  labs(x = "Years", y = "PM2.5 Concentration (μg/m³)") +
  scale_colour_manual(values = c("Provincial Average" = "blue", "10th Percentile" = "green", "90th Percentile" = "red")) +
  theme_minimal()

# Display the plot
p
```

```{r}
#| label: fig-four
#| fig-cap: Mortality Rates vs. Air Quality Health Index
#| echo: false
#| warning: false
#| message: false

library(MASS)

air_quality_data <- air_quality_data[air_quality_data$air_quality_health_index != "10+",]
air_quality_data$air_quality_health_index <- as.numeric(air_quality_data$air_quality_health_index)

# Calculate the average Air Quality Health Index per year
average_aqi_by_year <- air_quality_data %>%
  group_by(year) %>%
  summarize(average_aqi = mean(air_quality_health_index))

combined_data <- merge(deaths_data, average_aqi_by_year, by = "year")

# Plot scatter plot of mortality rates vs. air quality health index
ggplot(combined_data, aes(x = average_aqi, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~cause) +
  labs(
       x = "Average Air Quality Health Index",
       y = "Total Deaths")
```

```{r}
#| label: fig-cause
#| fig-cap: Total Deaths by Year for Each Cause
#| echo: false
#| warning: false
#| message: false

# Summarize total deaths for each disease by year
total_deaths_by_year <- deaths_data %>%
  group_by(year, cause) %>%
  summarize(total_deaths = sum(total_deaths))

# Plot for mortality data
ggplot(total_deaths_by_year, aes(x = as.factor(year), y = total_deaths, group = cause, color = cause)) +
  geom_line() +
  labs(
       x = "Year",
       y = "Total Deaths") +
  theme(legend.position = "top") +
  theme(legend.text = element_text(size = 5)) +
  guides(color = guide_legend(direction = "vertical"))
```

```{r}
#| label: fig-average_air
#| fig-cap: Average Air Quality Health Index by Year
#| echo: false
#| warning: false
#| message: false
# Ensure the 'year' column is in numeric format for proper line plotting
average_aqi_by_year$year <- as.numeric(as.character(average_aqi_by_year$year))

# Plotting with ggplot2
ggplot(average_aqi_by_year, aes(x = year, y = average_aqi)) +
  geom_line() +  # for lines
  geom_point() + # for points
  labs(
       x = "Year",
       y = "Average Air Quality Index") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) # Optional: Adjust x-axis breaks
```
Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align}
y_i | \mu_i, \phi &\sim \text{NegBin}(\mu_i, \phi) \\
\mu_i &= \exp(\alpha + \beta x_i) \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta &\sim \text{Normal}(0, 2.5) \\
\phi &\sim \text{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)
library(here)
library(ggplot2)
library(modelsummary)

# Assuming you've already read your models as before
neg_binom_model <- readRDS(file = here::here("models/neg_binom_model.rds"))
stan_neg_binom_model <- readRDS(file = here::here("models/stan_neg_binom_model.rds"))
stan_poisson_model <-readRDS(file = here::here("models/stan_poisson_model.rds"))

heart_model <- readRDS(file = here::here("models/heart_model.rds"))
lung_model <-readRDS(file = here::here("models/lung_model.rds"))

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

poisson_model <-
  readRDS(file = here::here("models/poisson_model.rds"))

# # Make predictions for heart and lung models
# heart_predictions <- predict(heart_model, newdata = data.frame(provincial_average = heart$provincial_average), se.fit = TRUE)
# lung_predictions <- predict(lung_model, newdata = data.frame(provincial_average = lung$provincial_average), se.fit = TRUE)
# # Create trend graphs
# # For example, let's plot predicted values against provincial average for heart data
# heart_pred_df <- data.frame(provincial_average = heart$provincial_average,
#                             total_deaths_pred = heart_predictions$fit,
#                             total_deaths_pred_upper = heart_predictions$fit + 1.96 * heart_predictions$se.fit,
#                             total_deaths_pred_lower = heart_predictions$fit - 1.96 * heart_predictions$se.fit)
# 
# # Plot
# ggplot(heart_pred_df, aes(x = provincial_average, y = total_deaths_pred)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper), alpha = 0.2) +
#   labs(x = "Provincial Average", y = "Predicted Total Deaths", title = "Predicted Total Deaths vs. Provincial Average (Heart Data)") +
#   theme_minimal()

# class(heart_model$provincial_average)
# class(lung_model$provincial_average)
# 
# # Create a range of values for the predictor variable
# provincial_average_range <- seq(min(heart$provincial_average, lung$provincial_average),
#                                 max(heart$provincial_average, lung$provincial_average),
#                                 length.out = 100)
# 
# # Make predictions for heart and lung models
# heart_predictions <- predict(heart_model, newdata = data.frame(provincial_average = provincial_average_range), se.fit = TRUE)
# lung_predictions <- predict(lung_model, newdata = data.frame(provincial_average = provincial_average_range), se.fit = TRUE)
# 
# # Create data frames for plotting
# heart_pred_df <- data.frame(provincial_average = provincial_average_range,
#                             total_deaths_pred = heart_predictions$fit,
#                             total_deaths_pred_upper = heart_predictions$fit + 1.96 * heart_predictions$se.fit,
#                             total_deaths_pred_lower = heart_predictions$fit - 1.96 * heart_predictions$se.fit)
# 
# lung_pred_df <- data.frame(provincial_average = provincial_average_range,
#                            total_deaths_pred = lung_predictions$fit,
#                            total_deaths_pred_upper = lung_predictions$fit + 1.96 * lung_predictions$se.fit,
#                            total_deaths_pred_lower = lung_predictions$fit - 1.96 * lung_predictions$se.fit)
# 
# # Plot predicted deaths for heart and lung models
# ggplot() +
#   geom_line(data = heart_pred_df, aes(x = provincial_average, y = total_deaths_pred, color = "Heart Data")) +
#   geom_ribbon(data = heart_pred_df, aes(x = provincial_average, ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper, fill = "Heart Data"), alpha = 0.2) +
#   geom_line(data = lung_pred_df, aes(x = provincial_average, y = total_deaths_pred, color = "Lung Data")) +
#   geom_ribbon(data = lung_pred_df, aes(x = provincial_average, ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper, fill = "Lung Data"), alpha = 0.2) +
#   labs(x = "Provincial Average", y = "Predicted Total Deaths", title = "Predicted Total Deaths due to Air Pollution") +
#   scale_color_manual(values = c("Heart Data" = "blue", "Lung Data" = "red")) +
#   scale_fill_manual(values = c("Heart Data" = "blue", "Lung Data" = "red"), guide = FALSE) +
#   theme_minimal()

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelcauses
#| tbl-cap: "Different causes of mortality and their death counts"
#| warning: false
library(modelsummary)

if (!requireNamespace("modelsummary", quietly = TRUE)) {
  install.packages("modelsummary")
}
library(modelsummary)

if (!requireNamespace("broom.mixed", quietly = TRUE)) {
  install.packages("broom.mixed")
}

library(broom.mixed)

coef_short_names <- 
  c("causeAll other forms of chronic ischemic heart disease" = "Ischemic Heart Disease",
    "causeMalignant neoplasms of trachea, bronchus and lung" = "Trachea/Bronchus/Lung Cancer",
    "Acute myocardial infarction" = "Heart Attack",
    "causeOther chronic obstructive pulmonary disease" = "COPD"
  )

# Summarize the models
modelsummary(
  list(
    "Poisson" = poisson_model,
    "Negative binomial" = first_model
  ),
  coef_map = coef_short_names,
  estimate_function = function(model) {
    broom.mixed::tidy(model, parameters = "all", conf.int = TRUE)
  }
)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelheart
#| tbl-cap: "Heart related causes death count and the air quality"
#| warning: false

modelsummary::modelsummary(
  list(
    "heart causes model" = heart_model
  ),
  #statistic = "mad",
  fmt = 2
)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modellung
#| tbl-cap: "Lung related causes death count and the air quality"
#| warning: false
modelsummary::modelsummary(
  list(
    "Lung Causes model" = lung_model
  ),
  #statistic = "mad",
  fmt = 2
)

```
# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckfirst
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme(legend.position = "bottom")

pp_check(poisson_model) +
  theme(legend.position = "bottom")
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckheartlung
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(heart_model) +
  theme(legend.position = "bottom")

pp_check(lung_model) +
  theme(legend.position = "bottom")

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")

plot(stan_neg_binom_model, "trace")

plot(stan_neg_binom_model, "rhat")
```

\newpage

# References

### 2024-03-22 paper/paper.tex
adding starter code files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

  \newcolumntype{d}{S[
    input-open-uncertainty=,
    input-close-uncertainty=,
    parse-numbers = false,
    table-align-text-pre=false,
    table-align-text-post=false
  ]}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air pollution},
  pdfauthor={Vanshika Vanshika; Shivank Goel; Navya Hooda},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Respiratory-Related Mortality Rates Show A Positive Correlation
With Increasing Air pollution\thanks{Code and data are available at:
https://github.com/shivankgoel003/Mortality-in-Alberta.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Based on Data Collected From The Province Of Alberta}
\author{Vanshika Vanshika \and Shivank Goel \and Navya Hooda}
\date{March 15, 2024}

\begin{document}
\maketitle
\begin{abstract}
First sentence. Second sentence. Third sentence. Fourth sentence.
\end{abstract}

\section{Introduction}\label{introduction}

The impact of air pollution on human health has increasingly become a
global concern, with respiratory illnesses being a significant
consequence of poor air quality. The National Library of Medicine
reports that approximately 4 million people die prematurely each year
from chronic respiratory diseases linked to air pollution Med (2020).
The World Health Organization also suggests that air pollution poses a
risk for all-cause mortality and specific diseases, especially the
exposure to air pollution is strongly linked with outcomes such as
stroke, ischemic heart disease, chronic obstructive pulmonary disease,
lung cancer, pneumonia, and cataracts Organization (2024).

The pollutants most significantly concerning for public health, are
particulate matter (PM), carbon monoxide (CO), ozone (O3), nitrogen
dioxide (NO2), and sulfur dioxide (SO2). Of these, fine particulate
matter is particularly worrisome due to its ability to penetrate the
lungs deeply, enter the bloodstream, and reach organs, leading to
systemic damage to tissues and cells.

Specifically, PM2.5 is of greater concern due to its harmful properties
and dangers. PM2.5 is airborne particulate matter (PM2.5) and is not a
single pollutant but rather a mixture of many chemical species and
aerosols. PM2.5 is associated with the greatest proportion of adverse
health effects related to air pollution, both in the United States and
worldwide. Long-term (months to years) exposure to PM2.5 has been linked
to premature death, particularly in people who have chronic heart or
lung diseases, and reduced lung function growth in children Board
(2024).

In this paper, we analyze data from Alberta, Canada, and aim to explore
the estimand of the number of deaths that can be correlated to the Air
Quality Health Index (AQHI) and PM2.5 quantities. Specifically, we
assess how AQHI relates to the prevalence of respiratory and cardiac
illnesses in this specific region, focusing on four major types: Chronic
Obstructive Pulmonary Disease, Ischemic Heart Disease, Acute Myocardial
Infarction, and Lung Cancer. We use the mortality rate in Alberta data
to select respiratory-related illnesses and air quality health index
data for Alberta through their provincial open data portal. To analyze
respiratory illnesses, we use air pollution measured through the AQHI
and specifically the particulate matter 2.5 (PM2.5) air pollutant as a
predictor of mortality related to respiratory illnesses. In total, we
will be assessing the overall relationship between AQHI and respiratory
illnesses, the relationship between PM2.5 and respiratory illnesses, as
well as the relevance of PM2.5 levels in lung and heart disease values
as a means to draw any notable significance of PM2.5 in different types
of illnesses.

Using negative binomial regression, this study seeks to uncover trends
and correlations between AQHI and the prevalence of respiratory
illnesses. Negative binomial regression is a type of generalized linear
model (GLM) that is specifically designed to model count data. In our
paper, we are using it to predict the number of deaths, based on the
year and illness, for the pollutants present in the air. By analyzing
this relationship, we provide valuable insights that can inform
policymakers, healthcare professionals, and the public about the impact
of air pollution on respiratory health in Alberta. This research aims to
contribute to a better understanding of the health effects of air
pollution and to support the development of targeted strategies for air
quality improvement and public health protection in Alberta. Our
analytical framework explores whether the variables assessed seem to
correlate, if at all, to the total deaths in respiratory-related
illnesses through 2012-2022.

This paper is organized as follows: In the Data section, we outline the
sources of three different datasets, detail the data-cleaning processes
applied to each dataset, and describe any data merging procedures used
to prepare the data for input into various models. The Results section
focuses on analyzing trends and correlations between AQHI, PM2.5, and
respiratory and cardiac illnesses. Additionally, we discuss the trends
and patterns identified by our model, along with the correlation
analysis between these variables. In the Discussion section, we present
our overall findings, discuss any biases and weaknesses in the data that
may have influenced these findings, and explain our approach to
analyzing these limitations.

Overall, this research has the potential to inform public health
strategies and interventions aimed at reducing respiratory illnesses and
improving overall health in Alberta and beyond.

\section{Data}\label{sec-data}

\subsection{Data Source and
Collection:}\label{data-source-and-collection}

The study relies on datasets obtained from the provincial open databases
of Alberta, accessible through the official website government (2024).
Three key datasets were utilized to extract relevant variables for
analysis, aiming to uncover the relationship between air quality and
mortality rates in Alberta. The analysis begins with the leading causes
of death dataset for Alberta, sourced from the provincial open data
portal government (2023b). This dataset provides insights into mortality
rates associated with various illnesses, facilitating the examination of
trends related to respiratory and heart-related illnesses. To explore
potential correlations between air quality and mortality rates, the
study incorporates the Air Quality Health Index (AQHI) dataset for
Alberta, sourced from the provincial open data portal government
(2023a). This dataset offers comprehensive information on the AQHI
across different municipalities in Alberta over multiple years.
Additionally, the study utilizes PM2.5 air pollutant concentration level
data sourced from Alberta's official resources Alberta Government
(2023). This dataset provides detailed information on the concentration
levels of PM2.5 pollutants over several years, offering valuable
insights into air quality trends. The following subsections outline the
sources, collection methodologies, and data-cleaning procedures
implemented to ensure the accuracy and reliability of the datasets used
in the analysis. This meticulous approach ensures that the data is
prepared for thorough analysis, facilitating the exploration of
correlations between air quality indicators and mortality rates in
Alberta.

Leading Causes of Death in Alberta Data: The disease data is found from
the government of Alberta's open data portal, and was last updated on
September 22, 2023 and continues to be updated annually. This dataset
encompasses mortality data related to the top 30 common causes of death.
It reports on types of diseases, causes of death, mortality denoted by
total death counts, and ranking for 2000-2022. Due to our focus on
respiratory illnesses, in the leading cause of death dataset, we grouped
diseases by categories. Our category of focus included filtering on
illnesses like acute myocardial infarction, malignant neoplasms of the
trachea, bronchus, and lung, other chronic obstructive pulmonary
disease, and all other forms of chronic ischemic heart disease. Leading
causes of death are measured and ranked by the top 30 most common death
causes each specific year. The causes of death are classified based on
the International Classification of Diseases 10th Edition.

AQHI Data: The second dataset we used is the air quality health index
(AQHI) dataset found at the government of Alberta's open data portal.
This dataset contains AQHI by municipality for the years 2012-2022 and
reports air quality health index, and health risk both quantitatively
and qualitatively. To use the AQHI dataset we employed simple
data-cleaning practice to maintain descriptive variable names and
readability. The data is measured by the percentage of hours for each
year at a given air quality level, by municipality. The Air Quality
Health Index is calculated based on the relative risks of a combination
of common air pollutants that is known to harm human health. These
pollutants are ozone (O3) at ground level, particulate matter (PM2.5),
and nitrogen dioxide (NO2). Risks are defined as follows: 1-3 High
Quality; 4-6 Moderate Quality; 7-9 Low Quality; 10+ Very Low Quality.

PM2.5 Data: We used the PM2.5 data set retrieved from Alberta.ca
(Government of Alberta) which was last updated in April 2023. It reports
on average PM2.5 concentration levels through the years 2000-2021 using
a provincial average, the 10th percentile quantities, and the 90th
percentile quantities, with a focus on 8 municipalities Edmonton, Fort
McMurray, Grande Prairie, Lethbridge, Medicine Hat, and Red Deer
respectively and lastly reports the Canadian Ambient Air Quality
Standard (CAAQS) value.

The Alberta Air Zone report Environment and Areas (2021), which is
linked to our dataset, provides a detailed explanation of the
measurement and processing of the PM2.5 quantity. Alberta Air Zones
divides Alberta into six air zones which are aligned with Alberta's
Land-use Framework regional boundaries. Ambient air quality in Alberta
is monitored at continuous air monitoring stations located within these
air zones. PM2.5 quantities are taken throughout these stations across
Alberta, and they measure the quantities in µg (micrograms per cubic
meter of air).

\subsection{Data Cleaning}\label{data-cleaning}

We used R (R Core Team 2023) and Wickham et al. (2019a) for data
cleaning and processing, utilizing packages like tidyverse (Wickham et
al. 2019b) for data manipulation and janitor (Firke 2023) for cleaning
column names. Other packages used includes \texttt{ggplot2} (Wickham
2016), \texttt{dplyr} (Wickham et al. 2023), \texttt{readr} (Wickham,
Hester, and Bryan 2024), \texttt{tibble} (Müller and Wickham 2023),
\texttt{janitor} (Firke 2023),\texttt{reshape2} (Wickham 2007),
\texttt{knitr} (Xie 2023), \texttt{ggbeeswarm} (Clarke, Sherrill-Mix,
and Dawson 2023), \texttt{ggrepel} (Slowikowski 2024),
\texttt{kableExtra}(Zhu 2024), \texttt{readxl}(Wickham and Bryan 2023),
\texttt{MASS}(Venables and Ripley 2002), \texttt{rstanarm}(Goodrich et
al. 2022), \texttt{modelsummary}(Arel-Bundock 2022) and \texttt{here}
(Müller 2020).

The raw air quality data were preprocessed to remove inconsistencies and
irrelevant information. Specifically, we filtered the dataset to include
observations from the years 2012 to 2021, which are relevant to our
analysis. Additionally, we merged this dataset with additional
information on peak pollution levels for comprehensive analysis. Similar
to the previous datasets, the raw mortality data underwent cleaning
procedures to focus on specific causes relevant to our analysis. We
filtered the dataset to include observations up to 2021 and merged it
with additional information on air quality for correlation analysis. The
raw data on AQHI were filtered to include observations from the years
2011 to 2021 for consistency with other datasets. Additionally, the data
were aggregated at the municipal level for further analysis.

\subsection{Data Modifications}\label{data-modifications}

In this study, we constructed unique datasets by thoughtfully selecting
and merging data from the Government of Alberta's open data portal,
Alberta.ca, spanning the years 2012 to 2022. Our process involved
merging variables from various datasets to create specific datasets
tailored for model building and analysis. One such dataset,
`cleaned\_chart\_data,' was created by merging variables such as causes
of death, total deaths, provincial average PM2.5 levels, and CAAQS. This
dataset was designed to facilitate our analysis of any significant
correlations between these variables. A snapshot of this data is
referenced in Table X. Additionally, we derived two other datasets,
`merged\_data' and `merged\_heart\_data,' by merging variables related
to heart disease numbers, lung disease numbers, and provincial average
PM2.5 values. These datasets were instrumental in examining the impact
of PM2.5 on each type of illness, as previously discussed. Overall, our
methodology ensured the creation of comprehensive datasets that allowed
for a detailed investigation into the relationships between PM2.5 levels
and various health outcomes in Alberta.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-all-1.pdf}

}

\caption{\label{fig-all}Provincial Average PM2.5 Quantity Vs Total
Deaths}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-lung-1.pdf}

}

\caption{\label{fig-lung}Lung Related Causes Mortality Rates vs Average
PM2.5 Quantity}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-heart-1.pdf}

}

\caption{\label{fig-heart}Heart Related Causes Mortality Rates vs
Average PM2.5 Quantity}

\end{figure}%

Talk more about it.

And also planes (\textbf{?@fig-planes}). (You can change the height and
width, but don't worry about doing that until you have finished every
other aspect of the paper - Quarto will try to make it look nice and the
defaults usually work well once you have enough text.)

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-annual-1.pdf}

}

\caption{\label{fig-annual}Annual Trends of PM2.5 Concentrations in
Alberta}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-four-1.pdf}

}

\caption{\label{fig-four}Mortality Rates vs.~Air Quality Health Index}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-cause-1.pdf}

}

\caption{\label{fig-cause}Total Deaths by Year for Each Cause}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-average_air-1.pdf}

}

\caption{\label{fig-average\_air}Average Air Quality Health Index by
Year}

\end{figure}%

Talk way more about it.

\section{Model}\label{model}

The goal of our modelling strategy is twofold. Firstly,\ldots{}

Here we briefly describe the Bayesian analysis model used to
investigate\ldots{} Background details and diagnostics are included in
Appendix~\ref{sec-model-details}.

\subsection{Model set-up}\label{model-set-up}

Define \(y_i\) as the number of seconds that the plane remained aloft.
Then \(\beta_i\) is the wing width and \(\gamma_i\) is the wing length,
both measured in millimeters.

\begin{align}
y_i | \mu_i, \phi &\sim \text{NegBin}(\mu_i, \phi) \\
\mu_i &= \exp(\alpha + \beta x_i) \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta &\sim \text{Normal}(0, 2.5) \\
\phi &\sim \text{Exponential}(1)
\end{align}

We run the model in R (R Core Team 2023) using the \texttt{rstanarm}
package of (\textbf{rstanarm?}). We use the default priors from
\texttt{rstanarm}.

\subsubsection{Model justification}\label{model-justification}

We expect a positive relationship between the size of the wings and time
spent aloft. In particular\ldots{}

We can use maths by including latex between dollar signs, for instance
\(\theta\).

\section{Results}\label{results}

Our results are summarized in \textbf{?@tbl-modelresults}.

\begin{table}

\caption{\label{tbl-modelcauses}Different causes of mortality and their
death counts}

\centering{

\centering
\begin{tabular}[t]{lcc}
\toprule
  & Poisson & Negative binomial\\
\midrule
Ischemic Heart Disease & \num{0.510} & \num{0.510}\\
 &  & \vphantom{2} (\num{0.002})\\
Trachea/Bronchus/Lung Cancer & \num{0.353} & \num{0.353}\\
 &  & \vphantom{1} (\num{0.002})\\
COPD & \num{0.004} & \num{0.004}\\
 &  & (\num{0.002})\\
\midrule
Num.Obs. & \num{9048} & \num{9048}\\
Log.Lik. & \num{-69064.136} & \num{-53402.269}\\
ELPD & \num{-69078.6} & \num{-53405.5}\\
ELPD s.e. & \num{468.1} & \num{70.9}\\
LOOIC & \num{138157.2} & \num{106811.0}\\
LOOIC s.e. & \num{936.3} & \num{141.9}\\
WAIC & \num{138157.2} & \num{106811.0}\\
RMSE & \num{97.30} & \num{97.30}\\
\bottomrule
\end{tabular}

}

\end{table}%

\begin{table}

\caption{\label{tbl-modelheart}Heart related causes death count and the
air quality}

\centering{

\centering
\begin{tabular}[t]{lc}
\toprule
  & heart causes model\\
\midrule
(Intercept) & \num{8.02}\\
 & (\num{0.18})\\
provincial\_average & \num{0.00}\\
 & (\num{0.01})\\
\midrule
Num.Obs. & \num{21}\\
Log.Lik. & \num{-163.733}\\
ELPD & \num{-164.5}\\
ELPD s.e. & \num{0.5}\\
LOOIC & \num{329.0}\\
LOOIC s.e. & \num{1.0}\\
WAIC & \num{328.9}\\
RMSE & \num{142.90}\\
\bottomrule
\end{tabular}

}

\end{table}%

\begin{table}

\caption{\label{tbl-modellung}Lung related causes death count and the
air quality}

\centering{

\centering
\begin{tabular}[t]{lc}
\toprule
  & Lung Causes model\\
\midrule
(Intercept) & \num{7.57}\\
 & (\num{0.20})\\
provincial\_average & \num{0.01}\\
 & (\num{0.01})\\
\midrule
Num.Obs. & \num{21}\\
Log.Lik. & \num{-160.858}\\
ELPD & \num{-161.6}\\
ELPD s.e. & \num{0.8}\\
LOOIC & \num{323.3}\\
LOOIC s.e. & \num{1.6}\\
WAIC & \num{323.2}\\
RMSE & \num{248.30}\\
\bottomrule
\end{tabular}

}

\end{table}%

\section{Discussion}\label{discussion}

\subsection{First discussion point}\label{sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

\subsection{Second discussion point}\label{second-discussion-point}

\subsection{Third discussion point}\label{third-discussion-point}

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

Weaknesses and next steps should also be included.

\newpage

\appendix

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}

\section{Additional data details}\label{additional-data-details}

\section{Model details}\label{sec-model-details}

\subsection{Posterior predictive
check}\label{posterior-predictive-check}

In \textbf{?@fig-ppcheckandposteriorvsprior-1} we implement a posterior
predictive check. This shows\ldots{}

In \textbf{?@fig-ppcheckandposteriorvsprior-2} we compare the posterior
with the prior. This shows\ldots{}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-ppcheckfirst-1.pdf}

}

\subcaption{\label{fig-ppcheckfirst-1}Posterior prediction check}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-ppcheckfirst-2.pdf}

}

\subcaption{\label{fig-ppcheckfirst-2}Comparing the posterior with the
prior}

\end{minipage}%

\caption{\label{fig-ppcheckfirst}Examining how the model fits, and is
affected by, the data}

\end{figure}%

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-ppcheckheartlung-1.pdf}

}

\subcaption{\label{fig-ppcheckheartlung-1}Posterior prediction check}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-ppcheckheartlung-2.pdf}

}

\subcaption{\label{fig-ppcheckheartlung-2}Comparing the posterior with
the prior}

\end{minipage}%

\caption{\label{fig-ppcheckheartlung}Examining how the model fits, and
is affected by, the data}

\end{figure}%

\subsection{Diagnostics}\label{diagnostics}

Figure~\ref{fig-stanareyouokay-1} is a trace plot. It shows\ldots{} This
suggests\ldots{}

Figure~\ref{fig-stanareyouokay-2} is a Rhat plot. It shows\ldots{} This
suggests\ldots{}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-1.pdf}

}

\subcaption{\label{fig-stanareyouokay-1}Trace plot}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-2.pdf}

}

\subcaption{\label{fig-stanareyouokay-2}Rhat plot}

\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-3.pdf}

}

\subcaption{\label{fig-stanareyouokay-3}Trace plot}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-4.pdf}

}

\subcaption{\label{fig-stanareyouokay-4}Rhat plot}

\end{minipage}%

\caption{\label{fig-stanareyouokay}Checking the convergence of the MCMC
algorithm}

\end{figure}%

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-pollution}
Alberta Government. 2023. {``Air Indicators -- Fine Particulate
Matter.''}
\url{https://www.alberta.ca/air-indicators-fine-particulate-matter}.

\bibitem[\citeproctext]{ref-model}
Arel-Bundock, Vincent. 2022. {``{modelsummary}: Data and Model Summaries
in {R}.''} \emph{Journal of Statistical Software} 103 (1): 1--23.
\url{https://doi.org/10.18637/jss.v103.i01}.

\bibitem[\citeproctext]{ref-pm}
Board, California Air Resources. 2024. {``Inhalable Particulate Matter
and Health (PM2.5 and PM10).''}
\url{https://ww2.arb.ca.gov/resources/inhalable-particulate-matter-and-health}.

\bibitem[\citeproctext]{ref-gg}
Clarke, Erik, Scott Sherrill-Mix, and Charlotte Dawson. 2023.
\emph{Ggbeeswarm: Categorical Scatter (Violin Point) Plots}.
\url{https://github.com/eclarke/ggbeeswarm}.

\bibitem[\citeproctext]{ref-report}
Environment, Ministry of, and Protected Areas. 2021. {``Status of Air
Quality in Alberta.''}
\url{https://open.alberta.ca/dataset/9b00aab3-c37d-4080-854e-5f329c621b92/resource/057c65ac-7837-49bb-9528-38c2611540c4/download/epa-alberta-air-zones-report-2019-2021.pdf}.

\bibitem[\citeproctext]{ref-jan}
Firke, Sam. 2023. \emph{Janitor: Simple Tools for Examining and Cleaning
Dirty Data}. \url{https://github.com/sfirke/janitor}.

\bibitem[\citeproctext]{ref-rstan}
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2022.
{``Rstanarm: {Bayesian} Applied Regression Modeling via {Stan}.''}
\url{https://mc-stan.org/rstanarm/}.

\bibitem[\citeproctext]{ref-air}
government, Alberta. 2023a. {``Air Quality Index by Municipality.''}
\url{https://open.alberta.ca/opendata/air-quality-index-by-municipality\#detailed}.

\bibitem[\citeproctext]{ref-deaths}
---------. 2023b. {``Leading Causes of Death.''}
\url{https://open.alberta.ca/opendata/leading-causes-of-death}.

\bibitem[\citeproctext]{ref-alberta}
---------. 2024. {``Alberta.''} \url{https://www.alberta.ca/}.

\bibitem[\citeproctext]{ref-ncbi}
Med, Lancet Respir. 2020. {``Prevalence and Attributable Health Burden
of Chronic Respiratory Diseases, 1990--2017: A Systematic Analysis for
the Global Burden of Disease Study 2017.''}
\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7284317/}.

\bibitem[\citeproctext]{ref-here}
Müller, Kirill. 2020. \emph{Here: A Simpler Way to Find Your Files}.
\url{https://here.r-lib.org/}.

\bibitem[\citeproctext]{ref-tib}
Müller, Kirill, and Hadley Wickham. 2023. \emph{Tibble: Simple Data
Frames}. \url{https://tibble.tidyverse.org/}.

\bibitem[\citeproctext]{ref-who}
Organization, World Health. 2024. {``Air Quality, Energy and Health.''}
\url{https://www.who.int/teams/environment-climate-change-and-health/air-quality-energy-and-health/health-impacts\#:~:text=The\%20main\%20pathway\%20of\%20exposure,and\%20ultimately\%20leading\%20to\%20disease}.

\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-repel}
Slowikowski, Kamil. 2024. \emph{Ggrepel: Automatically Position
Non-Overlapping Text Labels with 'Ggplot2'}.
\url{https://ggrepel.slowkow.com/}.

\bibitem[\citeproctext]{ref-mass}
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics
with s}. Fourth. New York: Springer.
\url{https://www.stats.ox.ac.uk/pub/MASS4/}.

\bibitem[\citeproctext]{ref-reshape}
Wickham, Hadley. 2007. {``Reshaping Data with the {reshape} Package.''}
\emph{Journal of Statistical Software} 21 (12): 1--20.
\url{http://www.jstatsoft.org/v21/i12/}.

\bibitem[\citeproctext]{ref-ggplot}
---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}.
Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\bibitem[\citeproctext]{ref-tidy}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019b.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-rohan}
---------, et al. 2019a. {``Welcome to the {tidyverse}.''} \emph{Journal
of Open Source Software} 4 (43): 1686.
\url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-readxl}
Wickham, Hadley, and Jennifer Bryan. 2023. \emph{Readxl: Read Excel
Files}. \url{https://CRAN.R-project.org/package=readxl}.

\bibitem[\citeproctext]{ref-dp}
Wickham, Hadley, Romain François, Lionel Henry, Kirill Müller, and Davis
Vaughan. 2023. \emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://dplyr.tidyverse.org}.

\bibitem[\citeproctext]{ref-read}
Wickham, Hadley, Jim Hester, and Jennifer Bryan. 2024. \emph{Readr: Read
Rectangular Text Data}. \url{https://readr.tidyverse.org}.

\bibitem[\citeproctext]{ref-knit}
Xie, Yihui. 2023. \emph{Knitr: A General-Purpose Package for Dynamic
Report Generation in r}. \url{https://yihui.org/knitr/}.

\bibitem[\citeproctext]{ref-kable}
Zhu, Hao. 2024. \emph{kableExtra: Construct Complex Table with 'Kable'
and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

\end{document}

### 2024-03-22 scripts/00-simulate_data.R

@@ -1,0 +1,62 @@
+#### Preamble ####
+# Purpose: Cleans the two data sets for the years 2018 to 2021
+# Author: Navya Hooda, Shivank Goel, Vanshika Vanshika
+# Date: 07 March 2024 
+# Contact: shivankg.goel@mail.utoronto.ca
+# License: MIT
+
+# Load the tidyverse package for data manipulation
+library(tidyverse)
+library(MASS)
+
+# Set seed for reproducibility
+set.seed(2024)
+
+#### Simulate death data ####
+
+# Since we have 4 diseases and 12 years, each disease should appear exactly 12 times to match the number of years
+death_data <- tibble(
+  year = rep(2011:2022, times = 4),  # Replicate each year 4 times to match the number of diseases
+  disease = rep(c("Other chronic obstructive pulmonary disease", 
+                  "All other forms of chronic ischemic heart disease", 
+                  "Acute myocardial infarction", 
+                  "Malignant neoplasms of trachea, bronchus and lung"), each = 12),  # Each disease repeated 12 times
+  deaths = rnbinom(n = 48, size = 100, prob = 0.02)  # We now need 48 random numbers to match the expanded dataset
+)
+
+#### Simulate air quality data ####
+
+# Assuming the PM2.5 data is the same for both locations, we replicate the PM2.5 data twice for each year
+pm25_data <- tibble(
+  year = rep(2011:2022, times = 2),  # Each year appears twice
+  pm25 = rnbinom(n = 24, size = 50, prob = 0.05)  # We now need 24 random numbers to match the number of years times two
+)
+
+#### Combine both datasets for modeling ####
+
+# Join datasets by 'year'
+combined_data <- left_join(death_data, pm25_data, by = "year")
+
+head(combined_data)
+
+# Fit a Negative Binomial Model
+model <- glm.nb(deaths ~ pm25 + factor(year) + factor(disease), data = combined_data)
+
+# Model Diagnostics
+par(mfrow=c(2,2))
+plot(model)
+
+# Plotting Observed vs. Predicted Deaths
+combined_data$predicted_deaths <- predict(model, type = "response")
+ggplot(combined_data, aes(x = deaths, y = predicted_deaths)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(x = "Observed Deaths", y = "Predicted Deaths") +
[…]

### 2024-03-22 scripts/02-data_cleaning.R

@@ -1,0 +1,249 @@
+#### Preamble ####
+# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Author: Navya h
+# Date: 6 April 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+#### Workspace setup ####
+library(tidyverse)
+library(haven)
+library(janitor)
+library(readxl)
+library(dplyr)
+library(readxl)
+
+#reading data
+
+#schooling mode data
+schoolmode = read_dta("inputs/raw data/schooling_mode_data.dta")
+#Colorado
+colorado = read_excel("inputs/raw data/learningmodel/Colorado_Districts_LearningModelData_Final.xlsx")
+#Connecticut
+connecticut = read_excel("inputs/raw data/learningmodel/Connecticut_Districts_LearningModelData_Final.xlsx")
+#Massachusets
+massachusets = read_excel("inputs/raw data/learningmodel/Massachusetts_Districts_LearningModelData_Final.xlsx")
+#Minnesota 
+minnesota = read_excel("inputs/raw data/learningmodel/Minnesota_Districts_LearningModelData_Final.xlsx")
+#Mississippi
+mississippi = read_excel("inputs/raw data/learningmodel/Mississippi_Schools_LearningModelData_Final.xlsx")
+#Ohio
+ohio = read_excel("inputs/raw data/learningmodel/Ohio_Districts_LearningModelData_Final.xlsx")
+#Rhode Island
+rhode = read_excel("inputs/raw data/learningmodel/RhodeIsland_Schools_LearningModelData_Final.xlsx")
+#Virginia
+virginia = read_excel("inputs/raw data/learningmodel/Virginia_Districts_LearningModelData_Final.xlsx")
+#West Virginia
+westvirginia = read_excel("inputs/raw data/learningmodel/WestVirginia_Districts_LearningModelData_Final.xlsx")
+#Winscosin
+winscosin = read_excel("inputs/raw data/learningmodel/Wisconsin_Schools_LearningModelData_Final.xlsx")
+#Wyoming
+wyoming = read_excel("inputs/raw data/learningmodel/Wyoming_Districts_LearningModelData_Final.xlsx")
+
+#cleaning
+
+# Simplified names
[…]

### 2024-03-22 scripts/03-test_data.R

@@ -1,0 +1,70 @@
+#### Preamble ####
+# Purpose: Data Validation and Testing for Mortality and Air Quality Data Analysis
+# Authors: Navya Hooda, Shivank Goel, Vanshika Vanshika
+# Date: 12th February 2024
+# Contact: shivankg.goel@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: tidyverse, testthat
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+
+#Unit Testing Reference:https://testthat.r-lib.org/reference/test_that.html 
+
+# Load the datasets
+cleaned_air_data <- read.csv("data/analysis_data/cleaned_air_data.csv")
+cleaned_avg_pm25_conc <- read.csv("data/analysis_data/cleaned_avg_pm25_conc.csv")
+cleaned_mortality_data <- read.csv("data/analysis_data/cleaned_file.csv")
+
+#### Test data ####
+
+# Test the structure and type of the air quality data
+test_that("Air quality data columns are of the correct type", {
+  expect_type(cleaned_air_data$year, "integer")
+  expect_type(cleaned_air_data$health_risk, "character")
+  expect_type(cleaned_air_data$original_value, "double")
+})
+
+# Test the structure and type of the PM2.5 concentration data
+test_that("PM2.5 concentration data columns are of the correct type", {
+  expect_type(cleaned_avg_pm25_conc$years, "integer")
+  expect_type(cleaned_avg_pm25_conc$provincial_average, "double")
+  expect_type(cleaned_avg_pm25_conc$x10th_percentile, "double")
+  expect_type(cleaned_avg_pm25_conc$x90th_percentile, "double")
+})
+
+# Test the structure and type of the mortality data
+test_that("Mortality data columns are of the correct type", {
+  expect_type(cleaned_mortality_data$year, "integer")
+  expect_type(cleaned_mortality_data$cause, "character")
+  expect_type(cleaned_mortality_data$ranking, "integer")
+  expect_type(cleaned_mortality_data$total_deaths, "integer")
+})
+
+# Test that the years are within the expected range
+test_that("Year values are within the expected range", {
+  expect_true(all(cleaned_air_data$year >= 2012 & cleaned_air_data$year <= 2022))
+  expect_true(all(cleaned_avg_pm25_conc$years >= 2012 & cleaned_avg_pm25_conc$years <= 2022))
[…]

### 2024-03-22 scripts/04-model.R

@@ -1,0 +1,85 @@
+#### Preamble ####
+# Purpose: Models different relations between variables such as mortality rates and air pollutants. 
+# Author: Vanshika Vanshika, Shivank Goel, Navya Hooda
+# Date: 13 March 2024
+# Contact: vanshika.vanshika@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(here)
+library(readr)
+library(MASS)
+
+#### Read data ####
+deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
+air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
+data_average <- read.csv(here("data/analysis_data/cleaned_chart_data.csv"))
+data_peak <- read.csv(here("data/analysis_data/cleaned_peak_data.csv"))
+alberta_cod <- merge(deaths_data, air_quality_data, by = "year")
+heart <- read.csv(here("data/analysis_data/merged_heart_data.csv"))
+lung <- read.csv(here("data/analysis_data/merged_data.csv"))
+
+# Fit Poisson regression model
+cause_of_death_alberta_poisson <-
+  stan_glm(
+    total_deaths ~ cause,
+    data = alberta_cod,
+    family = poisson(link = "log"),
+    seed = 853
+  )
+
+# Fit Negative Binomial regression model
+cause_of_death_alberta_neg_binomial <-
+  stan_glm(
+    total_deaths ~ cause,
+    data = alberta_cod,
+    family = neg_binomial_2(link = "log"),
+    seed = 853
+  )
+
+
+
+# Convert year and air_quality_health_index to numeric
+heart <- heart %>%
+  mutate(provincial_average = as.numeric(provincial_average),
+         total_deaths = as.numeric(total_deaths))
+lung <- lung %>%
+  mutate(provincial_average = as.numeric(provincial_average),
+         total_deaths = as.numeric(total_deaths))
+
+# Fit models
+
+# Bayesian Negative Binomial Model with stan_glm
+heart_model <- stan_glm(total_deaths ~ provincial_average, data = heart, 
+                        family = neg_binomial_2(link = "log"), seed = 853)
+
+# Poisson Regression Model
+lung_model <- stan_glm(total_deaths ~ provincial_average, data = lung, 
[…]

### 2024-03-26 paper/paper.qmd
updates in model

@@
 
-deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
-air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
-data <- read.csv(here("data/analysis_data/cleaned_chart_data.csv"))
-data_peak <- read.csv(here("data/analysis_data/cleaned_peak_data.csv"))
-data_lung <- read.csv(here("data/analysis_data/merged_data.csv"))
-data_heart <- read.csv(here("data/analysis_data/merged_heart_data.csv"))
+breach_data <- read.csv(here("data/analysis_data/breach_data.csv"))
 ```
@@
 
-# Plot the regression graph
-ggplot(data, aes(x = provincial_average, y = total_deaths)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
-  labs(x = "Provincial Average PM2.5", y = "Total Deaths") +
-  theme_minimal()
+breach_data %>%
+  count(year) %>%
+  ggplot(aes(x = year, y = n)) +
+  geom_line() +
+  labs(title = "Cyberattacks Over Time", x = "Year", y = "Number of Attacks")
+
+``` 
+```{r}
+breach_data %>%
+  count(sector) %>%
+  ggplot(aes(x = reorder(sector, n), y = n)) +
+  geom_bar(stat = "identity") +
+  coord_flip() +
+  labs(title = "Cyberattacks by Sector", x = "Sector", y = "Number of Attacks")
+```
+
+```{r}
+breach_data %>%
+  count(country) %>%
+  ggplot(aes(x = reorder(country, n), y = n)) +
+  geom_bar(stat = "identity") +
+  coord_flip() +
+  labs(title = "Cyberattacks by Country", x = "Country", y = "Number of Attacks")
 
@@
 ```{r}
-#| label: fig-lung
-#| fig-cap: Lung Related Causes Mortality Rates vs Average PM2.5 Quantity 
-#| echo: false
-#| warning: false
+breach_data %>%
+  count(country) %>%
+  ggplot(aes(x = reorder(country, n), y = n, fill = country)) +
+  geom_bar(stat = "identity") +
+  coord_flip() +
+  labs(title = "Cyberattacks by Country", x = "Country", y = "Number of Attacks") +
+  theme_minimal() +
+  theme(legend.position = "none") # Hides the legend
 
-# Plot the regression graph for all lungs causes
-ggplot(data_lung, aes(x = provincial_average, y = total_deaths)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
-  labs(x = "Provincial Average", y = "Total Deaths") +
-  theme_minimal()
 ```
@@
 ```{r}
-#| label: fig-heart
-#| fig-cap: Heart Related Causes Mortality Rates vs Average PM2.5 Quantity 
-#| echo: false
-#| warning: false
-# Plot the regression graph for all heart causes
-ggplot(data, aes(x = provincial_average, y = total_deaths)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
-  labs(x = "Provincial Average", y = "Total Deaths") +
-  theme_minimal()
+breach_data %>%
+  count(attack_type) %>%
+  ggplot(aes(x = "", y = n, fill = attack_type)) +
+  geom_bar(width = 1, stat = "identity") +
+  coord_polar("y") +
+  labs(title = "Distribution of Attack Types")
 
 ```
-
 Talk more about it.
@@
 ```{r}
-#| label: fig-annual
-#| fig-cap: Annual Trends of PM2.5 Concentrations in Alberta
-#| echo: false
-#| warning: false
-#| message: false
+breach_data %>%
+  ggplot(aes(x = as.factor(year), y = impact_on_data)) +
+  geom_boxplot() +
+  labs(title = "Impact on Data by Year", x = "Year", y = "Impact on Data")
 
-# Assuming your data frame is named 'cleaned_avg_PM25_conc' and is already filtered and renamed.
-# Convert factors to character to avoid issues in ggplot
-data[] <- lapply(data, function(x) if(is.factor(x)) as.character(x) else x)
-# Creating the plot with ggplot
-p <- ggplot(data, aes(x = year)) +
-  geom_line(aes(y = provincial_average, colour = "Provincial Average")) +
-  geom_line(aes(y = x10th_percentile, colour = "10th Percentile")) +
-  geom_line(aes(y = x90th_percentile, colour = "90th Percentile")) +
-  labs(x = "Years", y = "PM2.5 Concentration (μg/m³)") +
-  scale_colour_manual(values = c("Provincial Average" = "blue", "10th Percentile" = "green", "90th Percentile" = "red")) +
-  theme_minimal()
-
-# Display the plot
-p
 ```
-
 ```{r}
-#| label: fig-four
-#| fig-cap: Mortality Rates vs. Air Quality Health Index
-#| echo: false
-#| warning: false
-#| message: false
-
-library(MASS)
-
-air_quality_data <- air_quality_data[air_quality_data$air_quality_health_index != "10+",]
-air_quality_data$air_quality_health_index <- as.numeric(air_quality_data$air_quality_health_index)
-
-# Calculate the average Air Quality Health Index per year
-average_aqi_by_year <- air_quality_data %>%
-  group_by(year) %>%
-  summarize(average_aqi = mean(air_quality_health_index))
-
-combined_data <- merge(deaths_data, average_aqi_by_year, by = "year")
-
-# Plot scatter plot of mortality rates vs. air quality health index
-ggplot(combined_data, aes(x = average_aqi, y = total_deaths)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~cause) +
-  labs(
-       x = "Average Air Quality Health Index",
-       y = "Total Deaths")
-```
-
-```{r}
-#| label: fig-cause
-#| fig-cap: Total Deaths by Year for Each Cause
-#| echo: false
-#| warning: false
-#| message: false
-
-# Summarize total deaths for each disease by year
-total_deaths_by_year <- deaths_data %>%
-  group_by(year, cause) %>%
-  summarize(total_deaths = sum(total_deaths))
-
-# Plot for mortality data
-ggplot(total_deaths_by_year, aes(x = as.factor(year), y = total_deaths, group = cause, color = cause)) +
-  geom_line() +
-  labs(
-       x = "Year",
-       y = "Total Deaths") +
-  theme(legend.position = "top") +
-  theme(legend.text = element_text(size = 5)) +
-  guides(color = guide_legend(direction = "vertical"))
-```
-
-```{r}
-#| label: fig-average_air
-#| fig-cap: Average Air Quality Health Index by Year
-#| echo: false
-#| warning: false
-#| message: false
-# Ensure the 'year' column is in numeric format for proper line plotting
-average_aqi_by_year$year <- as.numeric(as.character(average_aqi_by_year$year))
-
-# Plotting with ggplot2
-ggplot(average_aqi_by_year, aes(x = year, y = average_aqi)) +
-  geom_line() +  # for lines
-  geom_point() + # for points
-  labs(
-       x = "Year",
-       y = "Average Air Quality Index") +
-  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) # Optional: Adjust x-axis breaks
-```
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align}
-y_i | \mu_i, \phi &\sim \text{NegBin}(\mu_i, \phi) \\
-\mu_i &= \exp(\alpha + \beta x_i) \\
-\alpha &\sim \text{Normal}(0, 2.5) \\
-\beta &\sim \text{Normal}(0, 2.5) \\
-\phi &\sim \text{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-library(here)
-library(ggplot2)
-library(modelsummary)
-
-# Assuming you've already read your models as before
-neg_binom_model <- readRDS(file = here::here("models/neg_binom_model.rds"))
-stan_neg_binom_model <- readRDS(file = here::here("models/stan_neg_binom_model.rds"))
-stan_poisson_model <-readRDS(file = here::here("models/stan_poisson_model.rds"))
-
-heart_model <- readRDS(file = here::here("models/heart_model.rds"))
-lung_model <-readRDS(file = here::here("models/lung_model.rds"))
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-
-poisson_model <-
-  readRDS(file = here::here("models/poisson_model.rds"))
-
-# # Make predictions for heart and lung models
-# heart_predictions <- predict(heart_model, newdata = data.frame(provincial_average = heart$provincial_average), se.fit = TRUE)
-# lung_predictions <- predict(lung_model, newdata = data.frame(provincial_average = lung$provincial_average), se.fit = TRUE)
-# # Create trend graphs
-# # For example, let's plot predicted values against provincial average for heart data
-# heart_pred_df <- data.frame(provincial_average = heart$provincial_average,
-#                             total_deaths_pred = heart_predictions$fit,
-#                             total_deaths_pred_upper = heart_predictions$fit + 1.96 * heart_predictions$se.fit,
-#                             total_deaths_pred_lower = heart_predictions$fit - 1.96 * heart_predictions$se.fit)
-# 
-# # Plot
-# ggplot(heart_pred_df, aes(x = provincial_average, y = total_deaths_pred)) +
-#   geom_line() +
-#   geom_ribbon(aes(ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper), alpha = 0.2) +
-#   labs(x = "Provincial Average", y = "Predicted Total Deaths", title = "Predicted Total Deaths vs. Provincial Average (Heart Data)") +
-#   theme_minimal()
-
-# class(heart_model$provincial_average)
-# class(lung_model$provincial_average)
-# 
-# # Create a range of values for the predictor variable
-# provincial_average_range <- seq(min(heart$provincial_average, lung$provincial_average),
-#                                 max(heart$provincial_average, lung$provincial_average),
-#                                 length.out = 100)
-# 
-# # Make predictions for heart and lung models
-# heart_predictions <- predict(heart_model, newdata = data.frame(provincial_average = provincial_average_range), se.fit = TRUE)
-# lung_predictions <- predict(lung_model, newdata = data.frame(provincial_average = provincial_average_range), se.fit = TRUE)
-# 
-# # Create data frames for plotting
-# heart_pred_df <- data.frame(provincial_average = provincial_average_range,
-#                             total_deaths_pred = heart_predictions$fit,
-#                             total_deaths_pred_upper = heart_predictions$fit + 1.96 * heart_predictions$se.fit,
-#                             total_deaths_pred_lower = heart_predictions$fit - 1.96 * heart_predictions$se.fit)
-# 
-# lung_pred_df <- data.frame(provincial_average = provincial_average_range,
-#                            total_deaths_pred = lung_predictions$fit,
-#                            total_deaths_pred_upper = lung_predictions$fit + 1.96 * lung_predictions$se.fit,
-#                            total_deaths_pred_lower = lung_predictions$fit - 1.96 * lung_predictions$se.fit)
-# 
-# # Plot predicted deaths for heart and lung models
-# ggplot() +
-#   geom_line(data = heart_pred_df, aes(x = provincial_average, y = total_deaths_pred, color = "Heart Data")) +
-#   geom_ribbon(data = heart_pred_df, aes(x = provincial_average, ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper, fill = "Heart Data"), alpha = 0.2) +
-#   geom_line(data = lung_pred_df, aes(x = provincial_average, y = total_deaths_pred, color = "Lung Data")) +
-#   geom_ribbon(data = lung_pred_df, aes(x = provincial_average, ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper, fill = "Lung Data"), alpha = 0.2) +
-#   labs(x = "Provincial Average", y = "Predicted Total Deaths", title = "Predicted Total Deaths due to Air Pollution") +
-#   scale_color_manual(values = c("Heart Data" = "blue", "Lung Data" = "red")) +
-#   scale_fill_manual(values = c("Heart Data" = "blue", "Lung Data" = "red"), guide = FALSE) +
-#   theme_minimal()
+breach_data %>%
+  count(year, attack_type) %>%
+  ggplot(aes(x = year, y = n, fill = attack_type)) +
+  geom_area(position = 'stack') +
+  labs(title = "Attack Types Over Years", x = "Year", y = "Count")
 
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelcauses
-#| tbl-cap: "Different causes of mortality and their death counts"
-#| warning: false
-library(modelsummary)
+library(ggplot2)
+library(dplyr)
 
-if (!requireNamespace("modelsummary", quietly = TRUE)) {
-  install.packages("modelsummary")
+# Assuming 'overall_nature_of_attack' is your categorical variable indicating the nature/type of attack
+# and 'number_of_users_affected' is a numeric variable indicating the number of users impacted
+
+# First, let's calculate summary statistics for each nature of attack
+
+breach_data %>%
+  count(year, sector) %>%
+  ggplot(aes(x = year, y = sector, fill = n)) +
+  geom_tile() +
+  scale_fill_gradient(low = "lightblue", high = "red") +
+  labs(title = "Heatmap of Cyberattacks per Year and Sector",
+       x = "Year",
+       y = "Sector")
+
+```
+```{r}
+breach_data %>%
+  count(year, overall_nature_of_attack) %>%
+  ggplot(aes(x = as.factor(year), y = n, fill = overall_nature_of_attack)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Stacked Bar Chart of Attack Types by Year",
+       x = "Year",
+       y = "Number of Attacks")
+
+```
+```{r}
+library(ggplot2)
+
+ggplot(breach_data, aes(x = as.factor(year), y = number_of_users_affected)) +
+  geom_boxplot() +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(title = "Boxplot of Number of Users Affected Across Years",
+       x = "Year",
+       y = "Number of Users Affected")
+```
+
+```{r}
+
+breach_data %>%
+  filter(year == 2013) %>%
+  arrange(desc(number_of_users_affected)) %>%
+  head()  # This shows the top entries for 2013
+
+breach_data %>%
+  filter(year == 2014) %>%
+  arrange(desc(number_of_users_affected)) %>%
+  head()  # This shows the top entries for 2013
+
+# Install and load the scales package
+if (!requireNamespace("scales", quietly = TRUE)) {
+  install.packages("scales")
 }
-library(modelsummary)
-
-if (!requireNamespace("broom.mixed", quietly = TRUE)) {
-  install.packages("broom.mixed")
-}
-
-library(broom.mixed)
-
-coef_short_names <- 
-  c("causeAll other forms of chronic ischemic heart disease" = "Ischemic Heart Disease",
-    "causeMalignant neoplasms of trachea, bronchus and lung" = "Trachea/Bronchus/Lung Cancer",
-    "Acute myocardial infarction" = "Heart Attack",
-    "causeOther chronic obstructive pulmonary disease" = "COPD"
-  )
-
-# Summarize the models
-modelsummary(
-  list(
-    "Poisson" = poisson_model,
-    "Negative binomial" = first_model
-  ),
-  coef_map = coef_short_names,
-  estimate_function = function(model) {
-    broom.mixed::tidy(model, parameters = "all", conf.int = TRUE)
-  }
-)
+library(scales)
 
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelheart
-#| tbl-cap: "Heart related causes death count and the air quality"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "heart causes model" = heart_model
-  ),
-  #statistic = "mad",
-  fmt = 2
-)
+breach_data %>%
+  count(country) %>%
+  mutate(percentage = n / sum(n) * 100) %>%
+  ggplot(aes(x = reorder(country, percentage), y = percentage)) +
+  geom_bar(stat = "identity") +
+  coord_flip() +
+  labs(title = "Percentage of Total Cyberattacks by Country",
+       x = "Country",
+       y = "Percentage of Attacks")
 
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modellung
-#| tbl-cap: "Lung related causes death count and the air quality"
-#| warning: false
-modelsummary::modelsummary(
-  list(
-    "Lung Causes model" = lung_model
-  ),
-  #statistic = "mad",
-  fmt = 2
-)
 
-```
-# Discussion
+breach_data %>%
+  mutate(impact_score = case_when(
+    impact_on_data == "Low" ~ 1,
+    impact_on_data == "Medium" ~ 2,
+    impact_on_data == "High" ~ 3,
+    TRUE ~ NA_real_
+  )) %>%
+  group_by(organisation_size) %>%
+  summarize(average_impact = mean(impact_score, na.rm = TRUE)) %>%
+  ggplot(aes(x = organisation_size, y = average_impact, fill = organisation_size)) +
+  geom_col() +
+  labs(title = "Average Impact Severity by Organization Size",
+       x = "Organisation Size",
+       y = "Average Impact Score") +
+  theme_minimal()
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckfirst
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme(legend.position = "bottom")
-
-pp_check(poisson_model) +
-  theme(legend.position = "bottom")
 ```
@@
 ```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckheartlung
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(heart_model) +
-  theme(legend.position = "bottom")
-
-pp_check(lung_model) +
-  theme(legend.position = "bottom")
+ggplot(breach_data, aes(x = number_of_users_affected)) +
+  geom_density(fill = "blue", alpha = 0.5) +
+  labs(title = "Density Plot of Number of Users Affected by Cyberattacks",
+       x = "Number of Users Affected",
+       y = "Density")
 
@@
 
-## Diagnostics
+```{r}
+ggplot(breach_data, aes(x = organisation_size, fill = organisation_size)) +
+  geom_bar() +
+  labs(title = "Number of Cyberattacks by Organization Size",
+       x = "Organization Size",
+       y = "Number of Cyberattacks") +
+  theme_minimal()
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+```
+```{r}
+library(modelsummary)
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+logistic_model <- readRDS(file = here::here("models/restructuring_model.rds"))
 
+modelsummary(list("Logistic Regression" = logistic_model))
+
+```
 ```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+breach_data <- breach_data %>% mutate(row_id = row_number())
 
-plot(first_model, "trace")
+# Adjust factors in your data to match the model's training data
+breach_data <- breach_data %>%
+  mutate(country = factor(country, levels = levels(logistic_model$model$country)))
 
-plot(first_model, "rhat")
+# Generate predictions
+breach_predictions <- predict(logistic_model, newdata = breach_data, type = "response")
 
-plot(stan_neg_binom_model, "trace")
+# Combine the predictions with the original data
+breach_data <- breach_data %>% mutate(predicted_prob = breach_predictions)
 
-plot(stan_neg_binom_model, "rhat")
+# Scatter plot with jitter
+ggplot(breach_data, aes(x = organisation_size, y = predicted_prob)) +
+  geom_jitter(alpha = 0.3) +
+  labs(x = "Organisation Size", y = "Predicted Probability of Restructuring")
+
 ```

### 2024-03-26 scripts/02-data_cleaning.R

@@ -1,249 +1,48 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Navya h
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Clean the cyberattack dataset for analysis on breaches from 2010 to 2020
+# Author: Shivank Goel
+# Date: 22 March 2024 
+# Contact: shivankg.goel@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(tidyverse)
-library(haven)
 library(janitor)
+library(lubridate)
 library(readxl)
 library(dplyr)
-library(readxl)
 
-#reading data
+#### Clean data ####
 
-#schooling mode data
-schoolmode = read_dta("inputs/raw data/schooling_mode_data.dta")
-#Colorado
-colorado = read_excel("inputs/raw data/learningmodel/Colorado_Districts_LearningModelData_Final.xlsx")
-#Connecticut
-connecticut = read_excel("inputs/raw data/learningmodel/Connecticut_Districts_LearningModelData_Final.xlsx")
-#Massachusets
-massachusets = read_excel("inputs/raw data/learningmodel/Massachusetts_Districts_LearningModelData_Final.xlsx")
-#Minnesota 
-minnesota = read_excel("inputs/raw data/learningmodel/Minnesota_Districts_LearningModelData_Final.xlsx")
-#Mississippi
-mississippi = read_excel("inputs/raw data/learningmodel/Mississippi_Schools_LearningModelData_Final.xlsx")
-#Ohio
-ohio = read_excel("inputs/raw data/learningmodel/Ohio_Districts_LearningModelData_Final.xlsx")
-#Rhode Island
-rhode = read_excel("inputs/raw data/learningmodel/RhodeIsland_Schools_LearningModelData_Final.xlsx")
-#Virginia
-virginia = read_excel("inputs/raw data/learningmodel/Virginia_Districts_LearningModelData_Final.xlsx")
-#West Virginia
-westvirginia = read_excel("inputs/raw data/learningmodel/WestVirginia_Districts_LearningModelData_Final.xlsx")
-#Winscosin
[…]

### 2024-03-26 scripts/04-model.R

@@ -1,85 +1,53 @@
 #### Preamble ####
-# Purpose: Models different relations between variables such as mortality rates and air pollutants. 
-# Author: Vanshika Vanshika, Shivank Goel, Navya Hooda
-# Date: 13 March 2024
-# Contact: vanshika.vanshika@mail.utoronto.ca
+# Purpose: Models different relations between variables such as cyberattack outcomes and organizational attributes. 
+# Author: Shivank Goel  
+# Date: [Today's Date]
+# Contact: [Your Contact Information]
 # License: MIT
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
-library(here)
 library(readr)
-library(MASS)
 
 #### Read data ####
-deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
-air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
-data_average <- read.csv(here("data/analysis_data/cleaned_chart_data.csv"))
-data_peak <- read.csv(here("data/analysis_data/cleaned_peak_data.csv"))
-alberta_cod <- merge(deaths_data, air_quality_data, by = "year")
-heart <- read.csv(here("data/analysis_data/merged_heart_data.csv"))
-lung <- read.csv(here("data/analysis_data/merged_data.csv"))
+# Assuming the data file is directly in the current working directory
+breach_data <- read_csv("data/analysis_data/breach_data.csv")
 
-# Fit Poisson regression model
-cause_of_death_alberta_poisson <-
+# Check the structure of the dataset
+str(breach_data)
+
+# Convert categorical variables to factors
+breach_data$organisation_size <- as.factor(breach_data$organisation_size)
+breach_data$country <- as.factor(breach_data$country)
+breach_data$restructuring_after_attack <- as.numeric(breach_data$restructuring_after_attack == "Yes")
+
+# Fit logistic regression model
+restructuring_model <-
   stan_glm(
-    total_deaths ~ cause,
-    data = alberta_cod,
-    family = poisson(link = "log"),
+    restructuring_after_attack ~ organisation_size + country,
+    data = breach_data,
+    family = binomial(link = "logit"),
     seed = 853
   )
 
-# Fit Negative Binomial regression model
-cause_of_death_alberta_neg_binomial <-
-  stan_glm(
-    total_deaths ~ cause,
-    data = alberta_cod,
[…]

### 2024-03-26 scripts/00-simulate_data.R
data simulation done

@@ -1,62 +1,54 @@
 #### Preamble ####
-# Purpose: Cleans the two data sets for the years 2018 to 2021
-# Author: Navya Hooda, Shivank Goel, Vanshika Vanshika
-# Date: 07 March 2024 
+# Purpose: Simulates the cyberattack dataset for analysis on breaches from 2010 to 2020
+# Author:  Shivank Goel
+# Date: 27 March 2024 
 # Contact: shivankg.goel@mail.utoronto.ca
 # License: MIT
 
 # Load the tidyverse package for data manipulation
 library(tidyverse)
 library(MASS)
-
 # Set seed for reproducibility
 set.seed(2024)
 
-#### Simulate death data ####
+#### Simulate cyberattack data ####
 
-# Since we have 4 diseases and 12 years, each disease should appear exactly 12 times to match the number of years
-death_data <- tibble(
-  year = rep(2011:2022, times = 4),  # Replicate each year 4 times to match the number of diseases
-  disease = rep(c("Other chronic obstructive pulmonary disease", 
-                  "All other forms of chronic ischemic heart disease", 
-                  "Acute myocardial infarction", 
-                  "Malignant neoplasms of trachea, bronchus and lung"), each = 12),  # Each disease repeated 12 times
-  deaths = rnbinom(n = 48, size = 100, prob = 0.02)  # We now need 48 random numbers to match the expanded dataset
-)
+# Simulating the number of organizations affected by cyberattacks for 5 years
+# We'll assume the number of affected organizations follows a negative binomial distribution
 
-#### Simulate air quality data ####
+# Defining variables
+years <- 2016:2020
+organization_sizes <- c("Small", "Medium", "Large", "NA")
+levels_of_digital_intensity <- c("Low", "Medium", "High", "NA")
+sectors <- c("Healthcare", "Finance", "Education", "Retail", "NA")
 
-# Assuming the PM2.5 data is the same for both locations, we replicate the PM2.5 data twice for each year
-pm25_data <- tibble(
-  year = rep(2011:2022, times = 2),  # Each year appears twice
-  pm25 = rnbinom(n = 24, size = 50, prob = 0.05)  # We now need 24 random numbers to match the number of years times two
-)
+# Simulating the data
+cyberattack_data <- expand.grid(year = years,
+                                organization_size = organization_sizes,
[…]

### 2024-04-02 paper/paper.qmd
testing graphs and models

@@
 ```{r}
-#| label: fig-all
-#| fig-cap: Provincial Average PM2.5 Quantity Vs Total Deaths
-#| echo: false
-#| warning: false
-
 breach_data %>%
   count(year) %>%
-  ggplot(aes(x = year, y = n)) +
-  geom_line() +
-  labs(title = "Cyberattacks Over Time", x = "Year", y = "Number of Attacks")
+  ggplot(aes(x = as.factor(year), y = n)) +  # Convert year to factor to treat it as discrete
+  geom_line(group=1) +  # Ensure geom_line treats the data as connected points
+  scale_x_discrete(breaks = levels(as.factor(breach_data$year))) +  # Specify breaks at every level of the factor
+  labs(title = "Cyberattacks Over Time", x = "Year", y = "Number of Attacks") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x labels to fit them all
 
@@
 ```
+```{r}
+ggplot(breach_data, aes(x = number_of_users_affected)) +
+  stat_ecdf(geom = "step") +
+  labs(title = "Cumulative Distribution of Number of Users Affected",
+       x = "Number of Users Affected",
+       y = "Cumulative Probability")
+```
+```{r}
+breach_data %>%
+  group_by(year) %>%
+  summarize(total_users_affected = sum(number_of_users_affected, na.rm = TRUE)) %>%
+  ggplot(aes(x = year, y = total_users_affected)) +
+  geom_line() +
+  labs(title = "Number of Users Affected Over Years",
+       x = "Year",
+       y = "Total Number of Users Affected") +
+  theme_minimal()
+
+```
+
+```{r}
+breach_data %>%
+  ggplot(aes(x = sector, y = number_of_users_affected)) +
+  geom_boxplot() +
+  labs(title = "Spread of Number of Users Affected Across Sectors",
+       x = "Sector",
+       y = "Number of Users Affected") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+breach_data %>%
+  ggplot(aes(x = number_of_users_affected)) +
+  geom_histogram(bins = 30, fill = "blue", color = "black") +
+  labs(title = "Distribution of Number of Users Affected by Cyberattacks",
+       x = "Number of Users Affected",
+       y = "Frequency") +
+  theme_minimal()
+
+```
+```{r}
+breach_data %>%
+  filter(year == 2013) %>%
+  summarise(
+    median_users = median(number_of_users_affected, na.rm = TRUE),
+    iqr_users = IQR(number_of_users_affected, na.rm = TRUE),
+    upper_bound = median_users + 1.5 * iqr_users
+  )
+
+breach_data %>% filter(year == 2019)
+
+breach_data %>%
+  filter(year == 2019)%>%
+  summarise(
+    median_users = median(number_of_users_affected, na.rm = TRUE),
+    iqr_users = IQR(number_of_users_affected, na.rm = TRUE),
+    upper_bound = median_users + 1.5 * iqr_users
+  )
+
+```
+
+```{r}
+ggplot(breach_data, aes(x = log1p(number_of_users_affected))) +
+  geom_density() +
+  labs(x = "Log(Number of Users Affected + 1)", y = "Density") +
+  theme_minimal()
+```
+
+```{r}
+library(modelsummary)
+lrmodel <- readRDS(file = here::here("models/lrmodel.rds"))
+
+modelsummary(list("Linear Regression" = lrmodel))
+
+ggplot(yearly_attacks, aes(x = year, y = number_of_attacks)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE, color = "blue") +
+  theme_minimal() +
+  labs(title = "Trend of Cyberattacks Over Years",
+       x = "Year", y = "Number of Attacks")
+
+library(broom)
+# Augment the data with model diagnostics
+augmented_data <- augment(lrmodel)
+
+# Plot 1: Histogram of Residuals
+ggplot(augmented_data, aes(x = .resid)) +
+  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
+  theme_minimal() +
+  labs(title = "Histogram of Residuals", y = "Number of Occurrences", x = "Residuals")
+
+# Plot 2: Residuals vs. Predictor Variable (year)
+ggplot(augmented_data, aes(x = year, y = .resid)) +
+  geom_point(color = "red") +
+  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
+  theme_minimal() +
+  labs(title = "Residuals vs. Year", y = "Residuals", x = "Year")
+
+# Plot 3: Residuals vs. Fitted Values
+ggplot(augmented_data, aes(x = .fitted, y = .resid)) +
+  geom_point(color = "green") +
+  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
+  theme_minimal() +
+  labs(title = "Residuals vs. Fitted Values", y = "Residuals", x = "Fitted Values")
+
+# Plot 4: Actual vs. Predicted Values
+ggplot(augmented_data, aes(x = number_of_users_affected, y = .fitted)) +
+  geom_point(color = "purple") +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  theme_minimal() +
+  labs(title = "Actual vs. Predicted Values", y = "Predicted", x = "Actual")
+
+```
+
+```{r}
+liner_model <- readRDS(file = here::here("models/linear_model.rds"))
+
+# Compare the actual outcomes with simulations from the posterior distribution
+pp_check(liner_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+augmented_data <- augment(sensitivity_model)
+
+# Residuals vs. Fitted Values
+ggplot(augmented_data, aes(x = .fitted, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dotted") +
+  theme_minimal() +
+  labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals")
+
+# Actual vs. Predicted Values
+ggplot(augmented_data, aes(x = number_of_users_affected, y = .fitted)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
+  theme_minimal() +
+  labs(title = "Actual vs. Predicted Values", x = "Actual Values", y = "Predicted Values")
+```
+
+```{r}
+cyberattack_neg_binomial <- readRDS(file = here::here("models/cyberattack_neg_binomial.rds"))
+cyberattack_poisson <- readRDS(file = here::here("models/cyberattack_poisson.rds"))
+
+#### Posterior Predictive Checks ####
+# Poisson model
+pp_poisson <- pp_check(cyberattack_poisson) + 
+  theme(legend.position = "bottom") +
+  labs(title = "Posterior Predictive Check for Poisson Model")
+
+# Negative Binomial model
+pp_neg_binomial <- pp_check(cyberattack_neg_binomial) + 
+  theme(legend.position = "bottom") +
+  labs(title = "Posterior Predictive Check for Negative Binomial Model")
+
+# Display the plots
+print(pp_poisson)
+print(pp_neg_binomial)
+
+```
+
+```{r}
+library(ggplot2)
+library(dplyr)
+library(forcats) # For fct_collapse
+
+# Assuming breach_data is already loaded and has columns named 'year', 'country', and 'frequency'
+# and frequency is already calculated
+
+# Prepare the dataset
+breach_data_summary <- breach_data %>%
+  mutate(country = fct_collapse(country,
+                                "Other countries" = setdiff(unique(country), 
+                                                              c("USA", "Australia", "Canada", "Global", "Japan", "UK")))) %>%
+  group_by(year, country) %>%
+  summarize(frequency = n(), .groups = 'drop')
+
+# Separating the data by country for ease of plotting
+country_data <- breach_data_summary %>%
+  filter(country %in% c("Australia", "Canada", "Global", "Japan", "UK", "Other countries"))
+
+usa_data <- breach_data_summary %>%
+  filter(country == "USA")
+
+# Find the max frequency for USA to scale the secondary axis accordingly
+max_usa_frequency <- max(usa_data$frequency)
+
+# Define a color palette
+color_palette <- c("Australia" = "#1b9e77", "Canada" = "#d95f02", "Global" = "#7570b3", "Japan" = "#e7298a", "UK" = "#66a61e", "Other countries" = "#e6ab02", "USA" = "#666666")
+
+# Plot
+gg <- ggplot() +
+  # Add bar plots for all countries except USA
+  geom_bar(data = country_data, aes(x = as.factor(year), y = frequency, fill = country), stat = "identity", position = "dodge") +
+  # Add line for USA with points
+  geom_line(data = usa_data, aes(x = as.factor(year), y = frequency), color = color_palette["USA"]) +
+  geom_point(data = usa_data, aes(x = as.factor(year), y = frequency), color = color_palette["USA"]) +
+  # Add primary y-axis for "Other countries"
+  scale_y_continuous(name = "Frequency of Cyber Attacks (Other Countries)", labels = scales::comma) +
+  # Add secondary y-axis for USA
+  scale_y_continuous(name = "Frequency of Cyber Attacks (USA)",
+                     sec.axis = sec_axis(~ . * max_usa_frequency / max(breach_data_summary$frequency), name = "USA", labels = scales::comma)) +
+  # Set the breaks for x-axis to every year level
+  scale_x_discrete(breaks = levels(as.factor(breach_data$year))) +
+  scale_fill_manual(values = color_palette) +
+  labs(title = "Overview of Cyber Attacks by Year and Country",
+       subtitle = "Bar plots for countries; line plot for USA scaled on secondary axis") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), # Rotate x-axis labels
+        legend.position = "bottom") # Adjust the position of the legend
+
+# Print the plot
+print(gg)
+
+# Save the plot
+ggsave("cyber_attacks_overview.png", plot = gg, width = 12, height = 8, dpi = 300)
+
+```
+
+```{r}
+
+library(ggplot2)
+library(dplyr)
+library(forcats) # For fct_collapse to group levels into "Other countries"
+
+# Assuming breach_data is already loaded and has columns named 'year', 'country', and 'frequency'
+# and frequency is already calculated (e.g., it is not just a raw count of rows)
+
+# Prepare the dataset
+breach_data_summary <- breach_data %>%
+  mutate(country = fct_collapse(country,
+                                "Other countries" = setdiff(unique(country), 
+                                                              c("USA", "Australia", "Canada", "Global", "Japan", "UK")))) %>%
+  group_by(year, country) %>%
+  summarize(frequency = n(), .groups = 'drop')
+
+# Separating the data by country for ease of plotting
+country_data <- breach_data_summary %>%
+  filter(country %in% c("Australia", "Canada", "Global", "Japan", "UK", "Other countries"))
+
+usa_data <- breach_data_summary %>%
+  filter(country == "USA")
+
+# Find the max frequency for USA to scale the secondary axis accordingly
+max_usa_frequency <- max(usa_data$frequency)
+
+# Plot
+gg <- ggplot() +
+  # Add bar plots for all countries except USA
+  geom_bar(data = country_data, aes(x = as.factor(year), y = frequency, fill = country), stat = "identity", position = "dodge") +
+  # Add line for USA with points
+  geom_line(data = usa_data, aes(x = as.factor(year), y = frequency), color = "blue") +
+  geom_point(data = usa_data, aes(x = as.factor(year), y = frequency), color = "blue") +
+  # Add primary y-axis for "Other countries"
+  scale_y_continuous(name = "Frequency of Cyber Attacks (Other Countries)") +
+  # Add secondary y-axis for USA
+  scale_y_continuous(name = "Frequency of Cyber Attacks (USA)",
+                     sec.axis = sec_axis(~ . * max_usa_frequency / max(breach_data_summary$frequency), name = "USA")) +
+  # Set the breaks for x-axis to every year level
+  scale_x_discrete(breaks = levels(as.factor(breach_data$year))) +
+  scale_fill_manual(values = c("Australia" = "color1", "Canada" = "color2", "Global" = "color3", "Japan" = "color4", "UK" = "color5", "Other countries" = "color6")) +
+  labs(title = "Overview of Cyber Attacks by Year and Country",
+       subtitle = "Bar plots for countries; line plot for USA scaled on secondary axis") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), # Rotate x-axis labels
+        legend.position = "bottom") # Adjust the position of the legend
+
+# Print the plot
+print(gg)
+
+```
+
+Make sure to replace "color1", "color2", etc., with the actual colors you want for the bars.
+Adjust the names of the columns if needed, based on your dataset.
+The `scale_y_continuous` for the USA might need to be adjusted based on the actual frequencies in your dataset.
+The `sec.axis` function is used to create the secondary
+
+```
 

### 2024-04-02 scripts/02-data_cleaning.R

@@ -20,13 +20,6 @@ breach_data <- read_excel("data/raw_data/Dataset-cyberattacks-06102020.xlsx", sh
 # Simplifying names
 breach_data <- clean_names(breach_data)
 
-# Filtering data from years 2010 to 2020
-breach_data <- breach_data %>%
-  filter(year >= 2010 & year <= 2020)
-
-# Dropping 'settlement_paid', 'effect_on_share_price', 'summary' columns
-breach_data <- breach_data %>%
-  select(-c(settlement_paid, effect_on_share_price, summary))
 
 # Ensure 'attack_type' is treated as a character
 breach_data$attack_type <- as.character(breach_data$attack_type)
@@ -40,7 +33,10 @@ breach_data <- breach_data %>%
   mutate(organisation_size = ifelse(is.na(organisation_size) | organisation_size == "NA", "Unknown", organisation_size))
 
 
+breach_data <- breach_data %>%
+  filter(number_of_users_affected != "Missing")
 
+breach_data$number_of_users_affected <- as.numeric(as.character(breach_data$number_of_users_affected))
 
 
 # Saving cleaned data

### 2024-04-02 scripts/04-model.R

@@ -8,7 +8,9 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(bayesplot)
 library(readr)
+library(MASS)
 
 #### Read data ####
 # Assuming the data file is directly in the current working directory
@@ -18,9 +20,13 @@ breach_data <- read_csv("data/analysis_data/breach_data.csv")
 str(breach_data)
 
 # Convert categorical variables to factors
-breach_data$organisation_size <- as.factor(breach_data$organisation_size)
 breach_data$country <- as.factor(breach_data$country)
 breach_data$restructuring_after_attack <- as.numeric(breach_data$restructuring_after_attack == "Yes")
+breach_data$organisation_size <- as.factor(breach_data$organisation_size)
+breach_data$sector <- as.factor(breach_data$sector)
+breach_data$critical_industry <- as.factor(breach_data$critical_industry)
+breach_data$level_of_digital_intensity <- as.factor(breach_data$level_of_digital_intensity)
+
 
 # Fit logistic regression model
 restructuring_model <-
@@ -34,20 +40,67 @@ restructuring_model <-
 # Summary of the model
 summary(restructuring_model)
 
-# Poisson regression model
-poisson_model <- glm(
-  number_of_users_affected ~ organisation_size + country,
-  data = breach_data,
-  family = poisson()
-)
 
-# Summary of the model
-summary(poisson_model)
+
+yearly_attacks <- breach_data %>%
+  group_by(year) %>%
+  summarise(number_of_attacks = n())
+
+# Linear regression model
+lrmodel <- lm(number_of_attacks ~ year, data = yearly_attacks)
 
 # Save the model
-saveRDS(poisson_model, "models/poisson_model.rds")
+saveRDS(lrmodel, "models/lrmodel.rds")
+
+# Fit a Bayesian linear regression model
+linear_model <- stan_glm(
+  number_of_users_affected ~ year, 
+  data = breach_data, 
+  family = gaussian(),
+  prior = normal(0, 2.5, autoscale = TRUE),
+  prior_intercept = normal(0, 2.5, autoscale = TRUE),
+  seed = 853
+)
+
+
+# Aggregate data to count number of attacks by year, sector, and organisation size
+aggregated_data <- breach_data %>%
+  group_by(sector, organisation_size) %>%
+  summarise(number_of_attacks = n(), .groups = 'drop')
[…]

### 2024-04-02 paper/paper.qmd
rough draft

--- 2024-03-26 paper/paper.qmd
@@
 ---
-title: "Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air pollution"
-subtitle: "Based on Data Collected From The Province Of Alberta"
+title: "TODO"
+subtitle: "TODO"
 author: 
-  - Vanshika Vanshika
   - Shivank Goel
-  - Navya Hooda
-thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
+thanks: "Code and data are available at: https://github.com/shivankgoel003/DataBreach_Ransomware_Stats"
 date: today
@@
 
-The impact of air pollution on human health has increasingly become a global concern, with respiratory illnesses being a significant consequence of poor air quality.
-The National Library of Medicine reports that approximately 4 million people die prematurely each year from chronic respiratory diseases linked to air pollution @ncbi.
-The World Health Organization also suggests that air pollution poses a risk for all-cause mortality and specific diseases, especially the exposure to air pollution is strongly linked with outcomes such as stroke, ischemic heart disease, chronic obstructive pulmonary disease, lung cancer, pneumonia, and cataracts @who.
+*TODO* today's world, where everything is connected online, cybersecurity is super important.
+It's all about keeping our digital stuff safe from bad guys who try to steal or mess with it.
+Cyber threats, like hackers breaking into computer systems or stealing personal information, are becoming more common and tricky to deal with.
 
-The pollutants most significantly concerning for public health, are particulate matter (PM), carbon monoxide (CO), ozone (O3), nitrogen dioxide (NO2), and sulfur dioxide (SO2).
-Of these, fine particulate matter is particularly worrisome due to its ability to penetrate the lungs deeply, enter the bloodstream, and reach organs, leading to systemic damage to tissues and cells.
+This paper is all about diving into the world of cybersecurity to learn about the problems we face, the trends we're seeing, and how we can protect ourselves better.
+We'll start by looking at some big cyber incidents that have happened recently.
+By understanding what happened in these incidents, we can figure out how to stop similar attacks in the future.
 
-Specifically, PM2.5 is of greater concern due to its harmful properties and dangers.
-PM2.5 is airborne particulate matter (PM2.5) and is not a single pollutant but rather a mixture of many chemical species and aerosols.
-PM2.5 is associated with the greatest proportion of adverse health effects related to air pollution, both in the United States and worldwide.
-Long-term (months to years) exposure to PM2.5 has been linked to premature death, particularly in people who have chronic heart or lung diseases, and reduced lung function growth in children @pm.
+Then, we'll talk about the new kinds of cyber threats that are popping up all the time.
+From sneaky tricks like ransomware to people tricking us into giving away our info, we'll explore the different ways bad guys try to break into our digital stuff.
 
-In this paper, we analyze data from Alberta, Canada, and aim to explore the estimand of the number of deaths that can be correlated to the Air Quality Health Index (AQHI) and PM2.5 quantities.
-Specifically, we assess how AQHI relates to the prevalence of respiratory and cardiac illnesses in this specific region, focusing on four major types: Chronic Obstructive Pulmonary Disease, Ischemic Heart Disease, Acute Myocardial Infarction, and Lung Cancer.
-We use the mortality rate in Alberta data to select respiratory-related illnesses and air quality health index data for Alberta through their provincial open data portal.
-To analyze respiratory illnesses, we use air pollution measured through the AQHI and specifically the particulate matter 2.5 (PM2.5) air pollutant as a predictor of mortality related to respiratory illnesses.
-In total, we will be assessing the overall relationship between AQHI and respiratory illnesses, the relationship between PM2.5 and respiratory illnesses, as well as the relevance of PM2.5 levels in lung and heart disease values as a means to draw any notable significance of PM2.5 in different types of illnesses.
-
-Using negative binomial regression, this study seeks to uncover trends and correlations between AQHI and the prevalence of respiratory illnesses.
-Negative binomial regression is a type of generalized linear model (GLM) that is specifically designed to model count data.
-In our paper, we are using it to predict the number of deaths, based on the year and illness, for the pollutants present in the air.
-By analyzing this relationship, we provide valuable insights that can inform policymakers, healthcare professionals, and the public about the impact of air pollution on respiratory health in Alberta.
-This research aims to contribute to a better understanding of the health effects of air pollution and to support the development of targeted strategies for air quality improvement and public health protection in Alberta.
-Our analytical framework explores whether the variables assessed seem to correlate, if at all, to the total deaths in respiratory-related illnesses through 2012-2022.
-
-This paper is organized as follows: In the Data section, we outline the sources of three different datasets, detail the data-cleaning processes applied to each dataset, and describe any data merging procedures used to prepare the data for input into various models.
-The Results section focuses on analyzing trends and correlations between AQHI, PM2.5, and respiratory and cardiac illnesses.
-Additionally, we discuss the trends and patterns identified by our model, along with the correlation analysis between these variables.
-In the Discussion section, we present our overall findings, discuss any biases and weaknesses in the data that may have influenced these findings, and explain our approach to analyzing these limitations.
-
-Overall, this research has the potential to inform public health strategies and interventions aimed at reducing respiratory illnesses and improving overall health in Alberta and beyond.
+Finally, we'll talk about what we can do to fight back against cyber threats.
+This includes using smart technology, making rules and policies to keep things safe, and teaching people how to be careful online.
+By learning more about cyber risks and working together to stay safe, we can make sure our digital world is a lot safer for everyone.
 
@@
 ## Data Source and Collection:
+
+**TODO**
 
@@
 ```
+```{r}
+ggplot(breach_data, aes(x = number_of_users_affected)) +
+  stat_ecdf(geom = "step") +
+  labs(title = "Cumulative Distribution of Number of Users Affected",
+       x = "Number of Users Affected",
+       y = "Cumulative Probability")
+```
+```{r}
+breach_data %>%
+  group_by(year) %>%
+  summarize(total_users_affected = sum(number_of_users_affected, na.rm = TRUE)) %>%
+  ggplot(aes(x = year, y = total_users_affected)) +
+  geom_line() +
+  labs(title = "Number of Users Affected Over Years",
+       x = "Year",
+       y = "Total Number of Users Affected") +
+  theme_minimal()
+
+```
+
+```{r}
+breach_data %>%
+  ggplot(aes(x = sector, y = number_of_users_affected)) +
+  geom_boxplot() +
+  labs(title = "Spread of Number of Users Affected Across Sectors",
+       x = "Sector",
+       y = "Number of Users Affected") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+breach_data %>%
+  ggplot(aes(x = number_of_users_affected)) +
+  geom_histogram(bins = 30, fill = "blue", color = "black") +
+  labs(title = "Distribution of Number of Users Affected by Cyberattacks",
+       x = "Number of Users Affected",
+       y = "Frequency") +
+  theme_minimal()
+
+```
+```{r}
+breach_data %>%
+  filter(year == 2013) %>%
+  summarise(
+    median_users = median(number_of_users_affected, na.rm = TRUE),
+    iqr_users = IQR(number_of_users_affected, na.rm = TRUE),
+    upper_bound = median_users + 1.5 * iqr_users
+  )
+
+breach_data %>% filter(year == 2019)
+
+breach_data %>%
+  filter(year == 2019)%>%
+  summarise(
+    median_users = median(number_of_users_affected, na.rm = TRUE),
+    iqr_users = IQR(number_of_users_affected, na.rm = TRUE),
+    upper_bound = median_users + 1.5 * iqr_users
+  )
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+breach_data$year <- as.numeric(as.character(breach_data$year))
+
+# Now proceed with your data manipulation
+breach_data_summary <- breach_data %>%
+  mutate(country = fct_collapse(country,
+                                "Other" = setdiff(unique(country), 
+                                                   c("USA", "Australia", "Canada", "Global", "Japan", "UK")))) %>%
+  group_by(year, country) %>%
+  summarize(frequency = n(), .groups = 'drop')
+
+# Separating the data by country for ease of plotting
+country_data <- breach_data_summary %>%
+  filter(country != "USA")
+
+usa_data <- breach_data_summary %>%
+  filter(country == "USA") %>%
+  arrange(year)
+
+# Ensure year is numeric for the line plot
+breach_data_summary$year <- as.numeric(as.character(breach_data_summary$year))
+
+# Recalculate max frequencies if necessary
+max_usa_freq <- max(usa_data$frequency, na.rm = TRUE)
+max_other_freq <- max(country_data$frequency, na.rm = TRUE)
+
+# Determine scale factor for secondary axis
+scale_factor <- max_other_freq / max_usa_freq
+breach_data_summary$year <- as.numeric(as.character(breach_data_summary$year))
+breach_data_summary <- breach_data_summary %>% filter(!is.na(year))
+
+# Plot
+# Define a color palette
+formal_palette <- c("Australia" = "#4878D0", "Canada" = "#6ACC65", "Global" = "#D65F5F", 
+                    "Japan" = "#B47CC7", "UK" = "#C4AD66", "Other" = "#77BEDB")
+
+# Plot
+gg <- ggplot() +
+  # Add bars for all countries except USA
+  geom_bar(data = country_data, aes(x = year, y = frequency, fill = country), stat = "identity", position = position_stack()) +
+  # Add line for USA
+  geom_line(data = usa_data, aes(x = year, y = frequency * scale_factor, group = 1), color = "#4D4D4D", size = 1) +
+  # Define the primary y-axis with the secondary axis for the USA
+  scale_y_continuous(
+    name = "Frequency of Cyber Attacks (Other Countries)", 
+    limits = c(0, max_other_freq * 1.1),  # Set limits for better control, slightly above max to ensure space for the line
+    sec.axis = sec_axis(~ . / scale_factor, name = "Frequency of USA Cyber Attacks", labels = scales::comma)
+  ) +
+  # Set breaks for the x-axis to unique years
+  scale_x_continuous(breaks = sort(unique(breach_data_summary$year))) +
+  # Apply the formal color palette
+  scale_fill_manual(values = formal_palette) +
+  labs(title = "Overview of Cyber Attacks by Year and Country",
+       subtitle = "Bar plots for countries; line plot for USA") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), # Rotate x-axis labels
+        legend.position = "bottom", # Adjust the position of the legend
+        plot.title = element_text(hjust = 0.5), # Center the plot title
+        plot.subtitle = element_text(hjust = 0.5)) # Center the plot subtitle
+
+# Print the plot
+print(gg)
+
+```
+
+```{r}
+
+```
 

### 2024-04-02 scripts/02-data_cleaning.R

@@ -31,13 +31,10 @@ breach_data <- breach_data %>%
 # Replace both NA values and "NA" strings with "Unknown" in 'organisation_size' column
 breach_data <- breach_data %>%
   mutate(organisation_size = ifelse(is.na(organisation_size) | organisation_size == "NA", "Unknown", organisation_size))
-
-
 breach_data <- breach_data %>%
   filter(number_of_users_affected != "Missing")
 
 breach_data$number_of_users_affected <- as.numeric(as.character(breach_data$number_of_users_affected))
-
 
 # Saving cleaned data
 write.csv(breach_data, file = "data/analysis_data/breach_data.csv", row.names = FALSE)

### 2024-04-02 scripts/04-model.R

@@ -46,6 +46,50 @@ yearly_attacks <- breach_data %>%
   group_by(year) %>%
   summarise(number_of_attacks = n())
 
+yearly_data <- breach_data %>%
+  group_by(year) %>%
+  summarise(
+    number_of_attacks = n(), sector
+    number_of_users_affected = sum(number_of_users_affected, na.rm = TRUE)
+    # You can add more summary stats if needed
+  )
+
+library(forecast)
+
+# Assuming 'number_of_attacks' is aggregated by year
+ts_data <- ts(yearly_data$number_of_attacks, start=min(yearly_data$year), frequency=1)
+
+# Fit ARIMA model
+arima_model <- auto.arima(ts_data)
+
+# Future predictions
+future_values <- forecast(arima_model, h=5)  # Forecast next 5 years
+
+# Plot
+autoplot(future_values)
+
+
+trend_model <- lm(number_of_users_affected ~ year, data = yearly_data)
+
+
+# For linear regression with interaction terms
+interaction_model <- lm(number_of_attacks ~ year * sector, data = yearly_data)
+
+# For separate time series in each sector
+sector_ts_models <- breach_data %>%
+  group_by(sector) %>%
+  do(trend_model = auto.arima(.$number_of_users_affected))
+
+
+# Plotting the trend of total number of users affected over the years
+ggplot(yearly_data, aes(x = year, y = number_of_users_affected)) +
+  geom_line() +
+  geom_point() +
+  labs(title = "Yearly Trend of Number of Users Affected by Cyberattacks",
+       x = "Year",
+       y = "Total Number of Users Affected")
+
+
 # Linear regression model
 lrmodel <- lm(number_of_attacks ~ year, data = yearly_attacks)
 
@@ -78,7 +122,7 @@ cyberattack_poisson <- stan_glm(
 
 #### Fit Negative Binomial Model with stan_glm ####
 cyberattack_neg_binomial <- stan_glm(
-  number_of_attacks ~ year + sector + organisation_size, 
+  number_of_attacks ~  sector + organisation_size, 
   data = aggregated_data, 
   family = neg_binomial_2(link = "log"), 
   seed = 853

### 2024-04-07 paper/paper.qmd
data section done, introduction done

@@
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+header-includes:
+   - \usepackage{tikz}
+   - \usetikzlibrary{shapes.geometric, arrows.meta, positioning, fit, calc}
+   - \usepackage{pgfplots} 
+   - \usepackage{float}
 format: pdf
@@
 
-*TODO* today's world, where everything is connected online, cybersecurity is super important.
-It's all about keeping our digital stuff safe from bad guys who try to steal or mess with it.
-Cyber threats, like hackers breaking into computer systems or stealing personal information, are becoming more common and tricky to deal with.
+In today's digital world, every click, every online transaction, and every shared piece of data is a potential entry point for cyber threats.
+Cybercrime is not just evolving, rather it is expanding at an alarming rate, with both the frequency and severity of attacks rising every year.
+The purpose of these attacks is to harm companies and organizations financially, however, in some cases these attacks can have military or political purposes.
+“According to a report published by the Identity Theft Resource Center (ITRC), a record number of 1862 data breaches occurred in 2021 in the US.
+Sectors like healthcare, finance, business, and retail are the most commonly attacked, impacting millions of Americans every year” (https://www.upguard.com/blog/biggest-data-breaches-us)
 
-This paper is all about diving into the world of cybersecurity to learn about the problems we face, the trends we're seeing, and how we can protect ourselves better.
-We'll start by looking at some big cyber incidents that have happened recently.
-By understanding what happened in these incidents, we can figure out how to stop similar attacks in the future.
+Despite such severe threats and impacts of cyberattacks, still certain companies tend to oversee this concern.
+As per PWC 2024 Global Digital Trust Insights report, “about one-third of organisations have no risk management plan to address cloud service provider challenges.
+Half are ‘very satisfied’ with their technology capabilities in key cybersecurity areas.
+More than 30% of companies don’t consistently follow what should be standard practices of cyber defence.” (https://www.pwc.com/us/en/services/consulting/cybersecurity-risk-regulatory/library/global-digital-trust-insights.html)
 
-Then, we'll talk about the new kinds of cyber threats that are popping up all the time.
-From sneaky tricks like ransomware to people tricking us into giving away our info, we'll explore the different ways bad guys try to break into our digital stuff.
+Therefore, in response to such attacks, there is a need for a plan, that not just keep the intruder or hackers out but also quickly alert if an attack does happen.
+Our study looks at cyber resilience, which is “ the ability to anticipate, withstand, recover from, and adapt to adverse conditions, stresses, attacks, or compromises on systems that use or are enabled by cyber resources.” as defined by National Institute of Standards and Technology (https://csrc.nist.gov/glossary/term/cyber_resiliency).
+To thoroughly analyze our study, we break it into three major research questions :
 
-Finally, we'll talk about what we can do to fight back against cyber threats.
-This includes using smart technology, making rules and policies to keep things safe, and teaching people how to be careful online.
-By learning more about cyber risks and working together to stay safe, we can make sure our digital world is a lot safer for everyone.
+**RQ1**: How do things like the size of the company and the type of business it does affect its ability to handle cyber attacks?
+
+**RQ2**: Which methods or strategies used by companies work best to reduce the damage from cyber attacks?
+
+**RQ3**: How does the business’s specific situation, like its industry or how much it relies on digital tools, change the impact of cyber attacks on it?
+
+The estimand of our study is the measurable effect of specific characteristics of an organization including size, sector and digital intensity on their cyber resilience.
+As a key finding, our regression models reveal factors such as organizational size, sector, and digital intensity significantly influence an organization's cyber resilience posture.For example, larger companies often have stronger defenses against cyber attacks, and on the other hand, companies that use a lot of digital technology in their work have different levels of protection.
+
+We aim to study and answer these questions by performing an analysis on a dataset of data breaches and ransomware attacks over 14 years from 2004, published by the University of Queensland.
+
+The remainder of this paper is structured as follows: @sec-data provides an overview of our methodology, including the data collection process and the analytical techniques used to explore the dataset of cyber attacks.
+We provide the background and overview of the study in @sec-background.
+@sec-model presents the regression models, discussing how we applied these models to understand the impact of various factors like organizational size, sector, and digital intensity on cyber resilience, @sec-results displays the interpretations of the model alongside other findings from analyzing the data, and @sec-discussion provides a discussion on the implications of the findings as well as the weaknesses of this paper and its next steps for further study on this subject.
+
+# Background {#sec-background}
+
+As discussed earlier, cyber resilience is about an organization's ability to keep its operations running smoothly in the face of cyber threats.
+It is not just about preventing cyber attacks, but also being prepared to deal with them effectively when they do happen.
+It is about recovery and adaptation, and extends beyond traditional cyber security measures.
+Cyber resilience surrounds various elements:
+
+1.
+Governance: This is the structure and processes that define the organization's approach to cyber threats.
+It is about leadership, accountability, and ensuring that the policies are in place and followed as desired.An effective governance is characterized by use of well defined frameworks, and presence of dedicated cybersecurity roles.
+
+- The use of well-defined frameworks that guide the organization's cybersecurity protocols.
+- The presence of dedicated cybersecurity roles such as a Chief Information Security Officer can prevent damage to IT systems and network.
+
+2. Prevention, Detection, and Recovery: These are the specific controls and strategies used to prevent attacks, detect them promptly, and recover from any damage caused. This approach involves:
+
+- Setting up appropriate remote access controls to secure unauthorized access.
+- Implementing proper network segmentation to control traffic flow and prevent the spread of threats within networks.
+- Adding an encryption to protect confidential data.
+- Utilizing detection systems to identify potential threats.
+- Developing restructuring plans as part of recovery measures to restore systems after the attack.
+
+3.
+Learning and Adapting: An organization needs to continuously learn from past incidents and attacks.
+It must adapt its strategies accordingly.
+This could involve updating its policies, training employees, and revising its approach to security.
+
+4.External Factors: Factors like the industry the organization is in, its size, and its digital intensity (how much it relies on digital technology) can also impact its cyber resilience.
+
+\begin{figure}[H]
+\centering
+\begin{tikzpicture}[
+    node distance=2.5 cm and 1.5 cm, % Distance between nodes
+    mynode/.style={ % Style for the nodes
+        ellipse, 
+        draw, 
+        align=center, 
+        minimum width=3cm, 
+        minimum height=1cm
+    },
+    myarrow/.style={ % Style for the arrows
+        ->, 
+        >=stealth, 
+        thick,
+        shorten <=2pt,
+        shorten >=2pt
+    },
+    mydoublearrow/.style={ % Style for double-sided arrows
+        <->, 
+        >=stealth, 
+        thick,
+        shorten <=2pt,
+        shorten >=2pt
+    }
+]
+
+% Nodes
+\node[mynode] (cyberres) {Cyber resilience};
+\node[mynode, above left=of cyberres] (gov) {Governance};
+\node[mynode, above right=of cyberres] (prev) {Prevention,\\ detection\\ and recovery};
+\node[mynode, below=of cyberres] (response) {Organisation's\\ responses};
+\node[mynode, left=of cyberres] (nature) {Nature of attack};
+\node[mynode, right=of response] (outcomes) {Outcomes};
+\node[mynode, below=of response] (exposure) {Exposure\\ factors};
+\node[mynode, below=of outcomes] (severity) {Breach\\ severity};
+\node[mynode, below=of exposure] (stakeholder) {Stakeholder\\ response};
+
+% Arrows
+\draw[myarrow] (gov) to[out=270,in=135] (cyberres);
+\draw[myarrow] (prev) to[out=270,in=45] (cyberres);
+\draw[myarrow] (nature) to[out=270,in=180] (cyberres);
+\draw[myarrow] (cyberres) -- (response);
+\draw[mydoublearrow] (cyberres) -- (outcomes);
+\draw[mydoublearrow] (cyberres) -- (exposure);
+
+% Reverse the direction of these arrows to point from 'outcomes' to its subcategories
+\draw[myarrow] (outcomes) -- (severity);
+\draw[myarrow] (outcomes) -- (stakeholder);
+
+\end{tikzpicture}
+\caption{Conceptual model of organizational cyber resilience}
+\label{fig:cyber_res_model}
+\end{figure}
 
@@
 ## Data Source and Collection: 
+Our analysis is based on sampling of 514 data breaches and ransomware attacks spanning over 14 years from 2004 to 2019.
+The dataset was obtained from the website of University of Queensland, and was prepared and compiled by researchers Tsen, Elinor, Ko, Ryan, and Slapnicar, Sergeja https://espace.library.uq.edu.au/view/UQ:dfe5027  at the University of Queensland.
+The data is thorough and encompasses a wide range of cyber attack incidents.
+It offers insights into various aspects of these incidents, including the types of breaches, affected organizations, and the extent of impact.
 
-**TODO**
-
-The study relies on datasets obtained from the provincial open databases of Alberta, accessible through the official website @alberta.
-Three key datasets were utilized to extract relevant variables for analysis, aiming to uncover the relationship between air quality and mortality rates in Alberta.
-The analysis begins with the leading causes of death dataset for Alberta, sourced from the provincial open data portal @deaths.
-This dataset provides insights into mortality rates associated with various illnesses, facilitating the examination of trends related to respiratory and heart-related illnesses.
-To explore potential correlations between air quality and mortality rates, the study incorporates the Air Quality Health Index (AQHI) dataset for Alberta, sourced from the provincial open data portal @air.
-This dataset offers comprehensive information on the AQHI across different municipalities in Alberta over multiple years.
-Additionally, the study utilizes PM2.5 air pollutant concentration level data sourced from Alberta's official resources @pollution.
-This dataset provides detailed information on the concentration levels of PM2.5 pollutants over several years, offering valuable insights into air quality trends.
-The following subsections outline the sources, collection methodologies, and data-cleaning procedures implemented to ensure the accuracy and reliability of the datasets used in the analysis.
-This meticulous approach ensures that the data is prepared for thorough analysis, facilitating the exploration of correlations between air quality indicators and mortality rates in Alberta.
-
-Leading Causes of Death in Alberta Data: The disease data is found from the government of Alberta’s open data portal, and was last updated on September 22, 2023 and continues to be updated annually.
-This dataset encompasses mortality data related to the top 30 common causes of death.
-It reports on types of diseases, causes of death, mortality denoted by total death counts, and ranking for 2000-2022.
-Due to our focus on respiratory illnesses, in the leading cause of death dataset, we grouped diseases by categories.
-Our category of focus included filtering on illnesses like acute myocardial infarction, malignant neoplasms of the trachea, bronchus, and lung, other chronic obstructive pulmonary disease, and all other forms of chronic ischemic heart disease.
-Leading causes of death are measured and ranked by the top 30 most common death causes each specific year.
-The causes of death are classified based on the International Classification of Diseases 10th Edition.
-
-AQHI Data: The second dataset we used is the air quality health index (AQHI) dataset found at the government of Alberta’s open data portal.
-This dataset contains AQHI by municipality for the years 2012-2022 and reports air quality health index, and health risk both quantitatively and qualitatively.
-To use the AQHI dataset we employed simple data-cleaning practice to maintain descriptive variable names and readability.
-The data is measured by the percentage of hours for each year at a given air quality level, by municipality.
-The Air Quality Health Index is calculated based on the relative risks of a combination of common air pollutants that is known to harm human health.
-These pollutants are ozone (O3) at ground level, particulate matter (PM2.5), and nitrogen dioxide (NO2).
-Risks are defined as follows: 1-3 High Quality; 4-6 Moderate Quality; 7-9 Low Quality; 10+ Very Low Quality.
-
-PM2.5 Data: We used the PM2.5 data set retrieved from Alberta.ca (Government of Alberta) which was last updated in April 2023.
-It reports on average PM2.5 concentration levels through the years 2000-2021 using a provincial average, the 10th percentile quantities, and the 90th percentile quantities, with a focus on 8 municipalities Edmonton, Fort McMurray, Grande Prairie, Lethbridge, Medicine Hat, and Red Deer respectively and lastly reports the Canadian Ambient Air Quality Standard (CAAQS) value.
-
-The Alberta Air Zone report @report, which is linked to our dataset, provides a detailed explanation of the measurement and processing of the PM2.5 quantity.
-Alberta Air Zones divides Alberta into six air zones which are aligned with Alberta’s Land-use Framework regional boundaries.
-Ambient air quality in Alberta is monitored at continuous air monitoring stations located within these air zones.
-PM2.5 quantities are taken throughout these stations across Alberta, and they measure the quantities in µg (micrograms per cubic meter of air).
+The dataset represents a detailed aggregation of data breaches and ransomware attacks, and as per authors, it was originally sourced from publicly disclosed media reports.
+The data integrates information from multiple public databases such as Privacy Rights Clearinghouse, Information is Beautiful, the Repository of Industrial Security Incidents, and Carnegie Mellon’s list of Banking Cyber Incidents.
+These sources were chosen for their public accessibility and frequent citation in academic and industry literature.
+The approach to data collection was guided by the PRISMA methodology which ensured a systematic and thorough compilation process.
 
@@
 
-The raw air quality data were preprocessed to remove inconsistencies and irrelevant information.
-Specifically, we filtered the dataset to include observations from the years 2012 to 2021, which are relevant to our analysis.
-Additionally, we merged this dataset with additional information on peak pollution levels for comprehensive analysis.
-Similar to the previous datasets, the raw mortality data underwent cleaning procedures to focus on specific causes relevant to our analysis.
-We filtered the dataset to include observations up to 2021 and merged it with additional information on air quality for correlation analysis.
-The raw data on AQHI were filtered to include observations from the years 2011 to 2021 for consistency with other datasets.
-Additionally, the data were aggregated at the municipal level for further analysis.
+The cyber breach data was preprocessed to remove inconsistencies and irrelevant information. Firstly,
+variable names were simplified and standardized for consistency and ease of analysis.
+A key challenge  faced was the significant number of missing values in the 'number of users affected' column.
+This variable was central to our study as we aimed to study trends related to the scale of impact using linear regression analysis.
+To address this issue, a choice was made to exclude records with missing or uncertain values in this column.
+While this decision resulted in some data loss, it was a necessary measure to maintain the integrity and accuracy of our trend analysis.
+Also, in columns like 'attack_type' and 'organisation_size', missing values were replaced with "Unknown" to maintain data integrity.
 
-## Data Modifications
+## Measurement and Exploratory Data Analysis
 
-In this study, we constructed unique datasets by thoughtfully selecting and merging data from the Government of Alberta's open data portal, Alberta.ca, spanning the years 2012 to 2022.
-Our process involved merging variables from various datasets to create specific datasets tailored for model building and analysis.
-One such dataset, 'cleaned_chart_data,' was created by merging variables such as causes of death, total deaths, provincial average PM2.5 levels, and CAAQS.
-This dataset was designed to facilitate our analysis of any significant correlations between these variables.
-A snapshot of this data is referenced in Table X.
-Additionally, we derived two other datasets, 'merged_data' and 'merged_heart_data,' by merging variables related to heart disease numbers, lung disease numbers, and provincial average PM2.5 values.
-These datasets were instrumental in examining the impact of PM2.5 on each type of illness, as previously discussed.
-Overall, our methodology ensured the creation of comprehensive datasets that allowed for a detailed investigation into the relationships between PM2.5 levels and various health outcomes in Alberta.
+As part of the measurement, we converted  real-world cyber incidents into quantifiable data within our dataset.
+The dataset variables were defined and measured based on the nature of the cyber incidents they represent.
+Each entry in the dataset corresponds to a distinct cyber incident, with variables relating to the incident.
+Here is how we defined and measured key variables:
+
+- **`organisation_size`**: This categorical variable categorizes the size of the affected organization into 'Small', 'Medium', 'Large', or 'Unknown', based on the number of employees or annual revenue as per commonly accepted business standards.
+
+- **`sector`**: The sector to which the affected organization belongs is classified according to standard industry classifications. This ensures each entry aligns with the appropriate economic sector.
+
+- **`cyber_security_role`**: This binary variable indicates the presence (`Yes`) or absence (`No`) of a dedicated cybersecurity role within the organization, at the time of the incident.
+
+- **`number_of_users_affected`**: Represented as a numerical variable, this measures the estimated number of individuals whose data was compromised during the breach
+
+- **`undertook_investigation`**: It captures whether an investigation was initiated following the cyber incident (1 for Yes, 0 for No).
+
+- **`breach_severity`**: To highlight the complexity and impact of cyber incidents, we introduced a custom variable, `breach_severity`.
+This variable was constructed to study the nature of cyber breaches, combining several key aspects of an incident:
+
+- **`impact_on_data`**: This reflects the nature of data compromise during the breach (categorized as 'High', 'Medium', or 'Low').
+
+- **`subsequent_fraudulent_use_of_data`**: Considers if the breached data was later used for fraudulent activities.
+
+The `breach_severity` variable was formulated through a custom function in our data processing script, which combined these elements to classify each incident into 'High', 'Medium', or 'Low' severity categories.
+This classification was based on the overall impact, the nature of data compromised, and the extent of misuse of data.
+This measure provides an understanding of the impact of each breach, beyond the simple binary or categorical measures commonly used.
+
+To achieve a clear understanding of the data, we include a variety of graphs and tables that represent the characteristics of each variable within our dataset.
+These visualizations illustrate the distribution and relationships among key variables, offering a broad picture of the patterns and trends among our data.
+@tbl-summary
+shows the summary statistics for organization size and sector.
+For each category within these variables, we present the count and the relative frequency, expressed as a percentage of the total sample.
+The frequency distribution of variables such as organisation_size and sector indicates the diversity of the dataset showing various sizes of organizations and a range of sectors.
+Table \ref{cyber_security_stats} shows  the decripitive statistics for certain variables.
+
+*CS Role Yes (27.27%)*: This shows that approximately 27% of the organizations in the dataset have a designated Cyber Security (CS) role.
+
+*CS Role No (72.73%)*: Conversely, nearly 73% of organizations do not have a designated CS role.
+
+*Framework Yes (36.36%)*: About 36% of the organizations adhere to a cyber security framework. Such frameworks provide structured guidelines and best practices for managing cyber security risks.
+
+*Framework No (63.64%)*: The majority, approximately 64%, do not follow a specific cyber security framework.
+
+*Prevention Low (45%)*:  45% of the organizations were categorized as having Low prevention measures, indicating basic or minimal preventive security measures.
+
+*Prevention Medium (36.36%)*: 36.36% fell into the Medium prevention category, suggesting more substantial but not so strong security measures.
+
+*Prevention High (18.18%)*: Only 18.18% were classified under High prevention, reflecting strong preventive strategies against cyber threats.
 
 ```{r}
+#| label: tbl-summary
+#| fig-cap: Summary Statistics for Organization Size and Sector
+#| echo: false
+#| warning: false
+#| message: false
+# Load the necessary libraries
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Assuming 'organisation_size' and 'sector' are some of the variables you want summarized
+# Create a function that will return the summary statistics you need
+get_summary_stats <- function(data, variable_name) {
+  summary_stats <- data %>%
+    group_by(!!sym(variable_name)) %>%
+    summarise(
+      Count = n(),
+      Frequency = n() / nrow(data) * 100
+    ) %>%
+    ungroup() %>%
+    arrange(desc(Frequency))
+
+  return(summary_stats)
+}
+
+# Call the function for 'organisation_size'
+summary_organisation_size <- get_summary_stats(breach_data, "organisation_size")
+
+# Call the function for 'sector'
+summary_sector <- get_summary_stats(breach_data, "sector")
+
+summary_combined <- rbind(
+  data.frame(
+    Variable = rep("a. Organization Size", nrow(summary_organisation_size)),
+    Category = summary_organisation_size$organisation_size,
+    Count = summary_organisation_size$Count,
+    `Frequency (%)` = round(summary_organisation_size$Frequency, 2)
+  ),
+  data.frame(
+    Variable = rep("b. Sector", nrow(summary_sector)),
+    Category = summary_sector$sector,
+    Count = summary_sector$Count,
+    `Frequency (%)` = round(summary_sector$Frequency, 2)
+  )
+)
+
+# Create the table using kable and add LaTeX formatting
+kable(summary_combined, format = "latex", booktabs = TRUE, escape = FALSE) %>%
+  kable_styling(latex_options = c("striped", "scale_down")) %>%
+  add_header_above(c(" " = 1, "Summary Statistics" = 3)) %>%
+  pack_rows("Organization Size", 1, nrow(summary_organisation_size), bold = TRUE) %>%
+  pack_rows("Sector", nrow(summary_organisation_size) + 1, nrow(summary_combined), bold = TRUE)
+
+```
+
+\begin{table}[H]
+\centering
+\caption{Descriptive Statistics for Cyber Security Variables}
+\label{cyber_security_stats} 
+\begin{tabular}{@{}ll@{}}
+\toprule
+\textbf{Variable} & \textbf{Frequency (\%)} \\ \midrule
+\multicolumn{2}{@{}l}{\textit{a. Governance (N = 514)}} \\
+CS Role Yes & 27.27 \\
+CS Role No & 72.73 \\ \addlinespace
+\multicolumn{2}{@{}l}{\textit{b. Cyber Security Frameworks (N = 514)}} \\
+Framework Yes & 36.36 \\
+Framework No & 63.64 \\ \addlinespace
+\multicolumn{2}{@{}l}{\textit{c. Prevention, Detection and Recovery}} \\
+Prevention Low & 45.45 \\
+Prevention Medium & 36.36 \\
+Prevention High & 18.18 \\ \addlinespace
+\end{tabular}
+\end{table}
+
+In order to observe the trend of number of cyberattacks over a span of years, from 2004 to 2019,
+we plotted a line graph [@fig-attacks].
+It is evident from the plot that the frequency of attacks has fluctuated over the years, with a peak in 2017.
+However, the decline following this peak may indicate the impact of improved cybersecurity measures, or a possible transition to different types of cyber threats not captured in this dataset.
+This visualization provides an overview of the nature of cyber threats and the ongoing battle between cybersecurity efforts and threat actors.
+
+```{r}
+#| label: fig-attacks
+#| fig-cap: Cyberattacks Over Time
+#| echo: false
+#| warning: false
+
 breach_data %>%
@@
   scale_x_discrete(breaks = levels(as.factor(breach_data$year))) +  # Specify breaks at every level of the factor
-  labs(title = "Cyberattacks Over Time", x = "Year", y = "Number of Attacks") +
+  labs(x = "Year", y = "Number of Attacks") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x labels to fit them all
@@
 ``` 
-```{r}
-breach_data %>%
-  count(sector) %>%
-  ggplot(aes(x = reorder(sector, n), y = n)) +
-  geom_bar(stat = "identity") +
-  coord_flip() +
-  labs(title = "Cyberattacks by Sector", x = "Sector", y = "Number of Attacks")
-```
+We also plotted a bar graph @fig-sector to count the number of incidents across various sectors.
+The bar chart clearly indicates that the 'Human Health Activities' sector has the highest count of incidents, standing out significantly from the other sectors.
+This might suggest that health sector is a more frequent target for cyber incidents or probably it is more diligent in reporting such events.
+The other sectors show a range of incident counts, with most appearing to have far fewer incidents in comparison.
+This could point to different levels of risk exposure, varying security measures, or reporting practices across these sectors.
 
 ```{r}
-breach_data %>%
-  count(country) %>%
-  ggplot(aes(x = reorder(country, n), y = n)) +
-  geom_bar(stat = "identity") +
-  coord_flip() +
-  labs(title = "Cyberattacks by Country", x = "Country", y = "Number of Attacks")
+#| label: fig-sector
+#| fig-cap: Bar Plot of Sector
+#| echo: false
+#| warning: false
+
+ggplot(breach_data, aes(x = sector)) +
+  geom_bar(fill = "coral", color = "black") +
+  labs(x = "Sector", y = "Count") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = rel(0.8)),
+        plot.margin = unit(c(1,1,1,4), "lines"))
 
@@
 
-```{r}
-breach_data %>%
-  count(country) %>%
-  ggplot(aes(x = reorder(country, n), y = n, fill = country)) +
-  geom_bar(stat = "identity") +
-  coord_flip() +
-  labs(title = "Cyberattacks by Country", x = "Country", y = "Number of Attacks") +
-  theme_minimal() +
-  theme(legend.position = "none") # Hides the legend
+[@fig-overviewofattack] is a creative visualization that effectively depicts the distribution and comparison of cyber attacks across various countries, with a specific emphasis on the United States.
+It combines a stacked bar chart for multiple countries and a line plot for the USA allowing for a dual-axis comparison due to the disproportionate number of attacks in the USA compared to other countries.
 
-```
+The bar segments represent the frequency of attacks in countries such as Australia, Canada, Japan, the UK, and others, with each color corresponding to a different country.
+The stacked nature of the bars shows how the total number of attacks is divided among these countries within each year.
+
+The line plot, on the other hand, tracks the frequency of cyber attacks in the USA across the same timeframe, adjusted by a scale factor for direct comparison on a secondary y-axis.
+This representation highlights the stark contrast in the volume of attacks between the USA and other countries while providing a clear year-by-year trend analysis.
+
+The choice to categorize all countries with fewer attacks under a consolidated "Other" category is a practical approach to maintain clarity in the visualization, avoiding overcrowding the chart with too many individual country representations.
 
 ```{r}
-breach_data %>%
-  count(attack_type) %>%
-  ggplot(aes(x = "", y = n, fill = attack_type)) +
-  geom_bar(width = 1, stat = "identity") +
-  coord_polar("y") +
-  labs(title = "Distribution of Attack Types")
+#| label: fig-overviewofattack
+#| fig-cap: Overview of Cyber Attacks by Year and Country
+#| echo: false
+#| warning: false
 
-```
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-breach_data %>%
-  ggplot(aes(x = as.factor(year), y = impact_on_data)) +
-  geom_boxplot() +
-  labs(title = "Impact on Data by Year", x = "Year", y = "Impact on Data")
-
-```
-```{r}
-breach_data %>%
-  count(year, attack_type) %>%
-  ggplot(aes(x = year, y = n, fill = attack_type)) +
-  geom_area(position = 'stack') +
-  labs(title = "Attack Types Over Years", x = "Year", y = "Count")
-
-```
-
-```{r}
-library(ggplot2)
-library(dplyr)
-
-# Assuming 'overall_nature_of_attack' is your categorical variable indicating the nature/type of attack
-# and 'number_of_users_affected' is a numeric variable indicating the number of users impacted
-
-# First, let's calculate summary statistics for each nature of attack
-
-breach_data %>%
-  count(year, sector) %>%
-  ggplot(aes(x = year, y = sector, fill = n)) +
-  geom_tile() +
-  scale_fill_gradient(low = "lightblue", high = "red") +
-  labs(title = "Heatmap of Cyberattacks per Year and Sector",
-       x = "Year",
-       y = "Sector")
-
-```
-```{r}
-breach_data %>%
-  count(year, overall_nature_of_attack) %>%
-  ggplot(aes(x = as.factor(year), y = n, fill = overall_nature_of_attack)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Stacked Bar Chart of Attack Types by Year",
-       x = "Year",
-       y = "Number of Attacks")
-
-```
-```{r}
-library(ggplot2)
-
-ggplot(breach_data, aes(x = as.factor(year), y = number_of_users_affected)) +
-  geom_boxplot() +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(title = "Boxplot of Number of Users Affected Across Years",
-       x = "Year",
-       y = "Number of Users Affected")
-```
-
-```{r}
-
-breach_data %>%
-  filter(year == 2013) %>%
-  arrange(desc(number_of_users_affected)) %>%
-  head()  # This shows the top entries for 2013
-
-breach_data %>%
-  filter(year == 2014) %>%
-  arrange(desc(number_of_users_affected)) %>%
-  head()  # This shows the top entries for 2013
-
-# Install and load the scales package
-if (!requireNamespace("scales", quietly = TRUE)) {
-  install.packages("scales")
-}
-library(scales)
-
-```
-
-```{r}
-breach_data %>%
-  count(country) %>%
-  mutate(percentage = n / sum(n) * 100) %>%
-  ggplot(aes(x = reorder(country, percentage), y = percentage)) +
-  geom_bar(stat = "identity") +
-  coord_flip() +
-  labs(title = "Percentage of Total Cyberattacks by Country",
-       x = "Country",
-       y = "Percentage of Attacks")
-
-```
-
-```{r}
-
-breach_data %>%
-  mutate(impact_score = case_when(
-    impact_on_data == "Low" ~ 1,
-    impact_on_data == "Medium" ~ 2,
-    impact_on_data == "High" ~ 3,
-    TRUE ~ NA_real_
-  )) %>%
-  group_by(organisation_size) %>%
-  summarize(average_impact = mean(impact_score, na.rm = TRUE)) %>%
-  ggplot(aes(x = organisation_size, y = average_impact, fill = organisation_size)) +
-  geom_col() +
-  labs(title = "Average Impact Severity by Organization Size",
-       x = "Organisation Size",
-       y = "Average Impact Score") +
-  theme_minimal()
-
-```
-
-```{r}
-ggplot(breach_data, aes(x = number_of_users_affected)) +
-  geom_density(fill = "blue", alpha = 0.5) +
-  labs(title = "Density Plot of Number of Users Affected by Cyberattacks",
-       x = "Number of Users Affected",
-       y = "Density")
-
-```
-
-```{r}
-ggplot(breach_data, aes(x = organisation_size, fill = organisation_size)) +
-  geom_bar() +
-  labs(title = "Number of Cyberattacks by Organization Size",
-       x = "Organization Size",
-       y = "Number of Cyberattacks") +
-  theme_minimal()
-
-```
-```{r}
-library(modelsummary)
-
-logistic_model <- readRDS(file = here::here("models/restructuring_model.rds"))
-
-modelsummary(list("Logistic Regression" = logistic_model))
-
-```
-```{r}
-breach_data <- breach_data %>% mutate(row_id = row_number())
-
-# Adjust factors in your data to match the model's training data
-breach_data <- breach_data %>%
-  mutate(country = factor(country, levels = levels(logistic_model$model$country)))
-
-# Generate predictions
-breach_predictions <- predict(logistic_model, newdata = breach_data, type = "response")
-
-# Combine the predictions with the original data
-breach_data <- breach_data %>% mutate(predicted_prob = breach_predictions)
-
-# Scatter plot with jitter
-ggplot(breach_data, aes(x = organisation_size, y = predicted_prob)) +
-  geom_jitter(alpha = 0.3) +
-  labs(x = "Organisation Size", y = "Predicted Probability of Restructuring")
-
-```
-```{r}
-ggplot(breach_data, aes(x = number_of_users_affected)) +
-  stat_ecdf(geom = "step") +
-  labs(title = "Cumulative Distribution of Number of Users Affected",
-       x = "Number of Users Affected",
-       y = "Cumulative Probability")
-```
-```{r}
-breach_data %>%
-  group_by(year) %>%
-  summarize(total_users_affected = sum(number_of_users_affected, na.rm = TRUE)) %>%
-  ggplot(aes(x = year, y = total_users_affected)) +
-  geom_line() +
-  labs(title = "Number of Users Affected Over Years",
-       x = "Year",
-       y = "Total Number of Users Affected") +
-  theme_minimal()
-
-```
-
-```{r}
-breach_data %>%
-  ggplot(aes(x = sector, y = number_of_users_affected)) +
-  geom_boxplot() +
-  labs(title = "Spread of Number of Users Affected Across Sectors",
-       x = "Sector",
-       y = "Number of Users Affected") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-```{r}
-breach_data %>%
-  ggplot(aes(x = number_of_users_affected)) +
-  geom_histogram(bins = 30, fill = "blue", color = "black") +
-  labs(title = "Distribution of Number of Users Affected by Cyberattacks",
-       x = "Number of Users Affected",
-       y = "Frequency") +
-  theme_minimal()
-
-```
-```{r}
-breach_data %>%
-  filter(year == 2013) %>%
-  summarise(
-    median_users = median(number_of_users_affected, na.rm = TRUE),
-    iqr_users = IQR(number_of_users_affected, na.rm = TRUE),
-    upper_bound = median_users + 1.5 * iqr_users
-  )
-
-breach_data %>% filter(year == 2019)
-
-breach_data %>%
-  filter(year == 2019)%>%
-  summarise(
-    median_users = median(number_of_users_affected, na.rm = TRUE),
-    iqr_users = IQR(number_of_users_affected, na.rm = TRUE),
-    upper_bound = median_users + 1.5 * iqr_users
-  )
-
-```
-
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
 breach_data$year <- as.numeric(as.character(breach_data$year))
@@
   scale_y_continuous(
-    name = "Frequency of Cyber Attacks (Other Countries)", 
+    name = "Frequency of Cyber Attacks (Other Countries except USA)", 
     limits = c(0, max_other_freq * 1.1),  # Set limits for better control, slightly above max to ensure space for the line
@@
   scale_fill_manual(values = formal_palette) +
-  labs(title = "Overview of Cyber Attacks by Year and Country",
-       subtitle = "Bar plots for countries; line plot for USA") +
+  labs(subtitle = "Bar plots for Other countries; line plot for USA") +
   theme_minimal() +
@@
 ```
+[@fig-attacktype] represents a stacked area chart, with each colored layer representing a different type of attack, allowing for an easy comparison of their occurrences over time.
+It is clear that some attack types, like installed malware, show peaks and troughs, possibly depicting the nature of cyber threats and security measures.
+These trends can be helpful for understanding the changing landscape of cyber risks and preparing for future security strategies.
 
 ```{r}
+#| label: fig-attacktype
+#| fig-cap: Attack Types Over Years
+#| echo: false
+#| warning: false
+
+attack_type_palette <- c(
+  "Installed malware" = "#4878D0", 
+  "Misuse of resources" = "#6ACC65", 
+  "Physical Theft" = "#D65F5F", 
+  "Unknown" = "#B47CC7", 
+  "Web compromise" = "#C4AD66"
+)
+
+breach_data %>%
+  count(year, attack_type) %>%
+  ggplot(aes(x = year, y = n, fill = attack_type)) +
+  geom_area(position = 'stack') +
+  labs(x = "Year", 
+       y = "Count", 
+       fill = "Attack Type") +
+  scale_fill_manual(values = attack_type_palette)
+```
+# Model {# sec-model}
+
+```{r}
+library(modelsummary)
+
+logistic_model <- readRDS(file = here::here("models/restructuring_model.rds"))
+
+modelsummary(list("Logistic Regression" = logistic_model))
 
 ```
+```{r}
+breach_data <- breach_data %>% mutate(row_id = row_number())
+
+# Adjust factors in your data to match the model's training data
+breach_data <- breach_data %>%
+  mutate(country = factor(country, levels = levels(logistic_model$model$country)))
+
+# Generate predictions
+breach_predictions <- predict(logistic_model, newdata = breach_data, type = "response")
+
+# Combine the predictions with the original data
+breach_data <- breach_data %>% mutate(predicted_prob = breach_predictions)
+
+# Scatter plot with jitter
+ggplot(breach_data, aes(x = organisation_size, y = predicted_prob)) +
+  geom_jitter(alpha = 0.3) +
+  labs(x = "Organisation Size", y = "Predicted Probability of Restructuring")
+
+```
+
+```{r}
+breach_data %>%
+  group_by(year) %>%
+  summarize(total_users_affected = sum(number_of_users_affected, na.rm = TRUE)) %>%
+  ggplot(aes(x = year, y = total_users_affected)) +
+  geom_line() +
+  labs(title = "Number of Users Affected Over Years",
+       x = "Year",
+       y = "Total Number of Users Affected") +
+  theme_minimal()
+
+```
+
+```{r}
+# Summary statistics for numerical variables
+summary(breach_data$number_of_users_affected)
+summary(breach_data$year)
+
+# Frequency tables for categorical variables
+table(breach_data$sector)
+table(breach_data$organisation_size)
+table(breach_data$critical_industry)
+table(breach_data$level_of_digital_intensity)
+table(breach_data$country)
+table(breach_data$cyber_security_role)
+table(breach_data$cyber_security_frameworks)
+table(breach_data$education_and_awareness_policy)
+table(breach_data$policy)
+table(breach_data$prevention_detection_and_recovery)
+table(breach_data$detector)
+table(breach_data$restructuring_after_attack)
+table(breach_data$bribe_ransom_paid)
+table(breach_data$free_identity_or_credit_theft_monitoring)
+table(breach_data$additional_disclosure_of_information)
+table(breach_data$overall_nature_of_attack)
+table(breach_data$attack_type)
+table(breach_data$attacker)
+table(breach_data$attack_vector)
+table(breach_data$impact_on_data)
+# ... continue for other categorical variables as needed
+
+# Histogram for a continuous variable (e.g., number_of_users_affected)
+ggplot(breach_data, aes(x = number_of_users_affected)) +
+  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
+  labs(title = "Histogram of Number of Users Affected", x = "Number of Users Affected", y = "Frequency")
+
+```
+```{r}
+linear_model_RQ2 <- readRDS(file = here::here("models/linear_model_RQ2.rds"))
+# Plotting diagnostics for the linear regression model (Example: linear_model_RQ2)
+ggplot(breach_data, aes(x = year, y = number_of_users_affected)) +
+  geom_point() +
+  geom_smooth(method = "lm", color = "blue") +
+  labs(title = "Linear Regression of Number of Users Affected Over Years", 
+       x = "Year", 
+       y = "Number of Users Affected")
+
+# Assuming breach_data is your dataset and you have already created models named linear_model_severity and linear_model_investigations
+
+# Partial regression plot for breach severity with a specific predictor (e.g., 'organisation_size')
+# install.packages("car") # Uncomment if the car package is not installed
+
+```
+
+```{r regression-table, results='asis', echo=FALSE}
+
+library(broom)
+library(stargazer)
+model_investigation <- readRDS(file = here::here("models/model_investigation.rds"))
+
+model_investigation_frame <- augment(model_investigation)
+
+ggplot(model_investigation_frame, aes(x = organisation_size, y = .resid)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
+  theme_classic() +
+  labs(y = "Residuals", x = "Organisation Size")
+
+ggplot(model_investigation_frame, aes(x = .resid)) +
+  geom_histogram(binwidth = 1) +
+  theme_classic() +
+  labs(y = "Number of occurrences", x = "Residuals")
+
+```
+$\begin{table}[!htbp] \centering 
+  \caption{Linear regression analysis with dependent variable} 
+  \label{} 
+\begin{tabular}{@{\extracolsep{5pt}}lc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-2} 
+\\[-1.8ex] & undertook\_investigation \\ 
+\hline \\[-1.8ex] 
+ critical\_industry & $-$0.03 (0.12) \\ 
+  organisation\_sizeMedium & $-$0.04 (0.06) \\ 
+  organisation\_sizeSmall & 0.07 (0.08) \\ 
+  organisation\_sizeUnknown & $-$0.03 (0.06) \\ 
+  level\_of\_digital\_intensityLow & $-$0.03 (0.62) \\ 
+  level\_of\_digital\_intensityLow-Medium & $-$0.29 (0.67) \\ 
+  level\_of\_digital\_intensityMedium-High & $-$0.31 (0.66) \\ 
+  sectorAdministrative and support service & $-$0.62 (0.72) \\ 
+  sectorAdvertising and other business services & $-$0.48 (0.66) \\ 
+  sectorArts, entertainment and recreation & $-$0.14 (0.12) \\ 
+  sectorChemicals and chemical products & 0.20 (0.37) \\ 
+  sectorComputer, electronic and optical products & 0.19 (0.23) \\ 
+  sectorConstruction & $-$0.61 (0.45) \\ 
+  sectorEducation & $-$0.11 (0.18) \\ 
+  sectorElectrical equipment & $-$0.23 (0.51) \\ 
+  sectorElectricity, gas, steam and air conditioning & $-$0.56 (0.34) \\ 
+  sectorFinance and insurance & $-$0.09 (0.63) \\ 
+  sectorFood products, beverages and tobacco & $-$0.06 (0.26) \\ 
+  sectorHuman health activities & $-$0.14 (0.22) \\ 
+  sectorIT and other information services & $-$0.29 (0.63) \\ 
+  sectorLegal and accounting activities & $-$0.63 (0.69) \\ 
+  sectorMachine equipment & $-$0.08 (0.24) \\ 
+  sectorPharmaceutical products & $-$0.27 (0.38) \\ 
+  sectorPublishing, audiovisual and broadcasting & $-$0.12 (0.22) \\ 
+  sectorResidential care and social work activities & $-$0.18 (0.24) \\ 
+  sectorScientific research and development & $-$0.37 (0.68) \\ 
+  sectorTelecommunications & $-$0.40 (0.65) \\ 
+  sectorTextiles, wearing apparel and leather &  \\ 
+  sectorTransportation storage & 0.06 (0.31) \\ 
+  sectorWholesale, retail trade and repair &  \\ 
+  countryChina & 0.88 (0.51) \\ 
+  countryFrance & $-$0.17 (0.51) \\ 
+  countryGermany & $-$0.01 (0.40) \\ 
+  countryGlobal & 0.28 (0.29) \\ 
+  countryHong Kong & 0.56 (0.57) \\ 
+  countryIndia & $-$0.33 (0.51) \\ 
+  countryJapan & 0.13 (0.35) \\ 
+  countryNorway & 0.04 (0.50) \\ 
+  countryPhilippines & $-$0.15 (0.51) \\ 
+  countryQatar & $-$0.33 (0.51) \\ 
+  countryRussia & $-$0.12 (0.41) \\ 
+  countrySingapore & 0.83$^{*}$ (0.36) \\ 
+  countrySouth Africa & 0.85 (0.51) \\ 
+  countrySouth Korea & 0.53 (0.36) \\ 
+  countryTurkey & 0.85 (0.51) \\ 
+  countryUAE & 0.60 (0.57) \\ 
+  countryUK & 0.35 (0.31) \\ 
+  countryUSA & 0.19 (0.26) \\ 
+  Constant & 0.45 (0.70) \\ 
+ \hline \\[-1.8ex] 
+Observations & 513 \\ 
+R$^{2}$ & 0.15 \\ 
+Adjusted R$^{2}$ & 0.07 \\ 
+Residual Std. Error & 0.43 (df = 465) \\ 
+F Statistic & 1.80$^{**}$ (df = 47; 465) \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.05; $^{**}$p$<$0.01; $^{***}$p$<$0.001} \\ 
+\end{tabular} 
+\end{table} $
 

### 2024-04-07 scripts/02-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Clean the cyberattack dataset for analysis on breaches from 2010 to 2020
+# Purpose: Clean the cyberattack dataset for analysis on cyber breaches
 # Author: Shivank Goel
 # Date: 22 March 2024 
 # Contact: shivankg.goel@mail.utoronto.ca
@@ -10,6 +10,7 @@ library(tidyverse)
 library(janitor)
 library(lubridate)
 library(readxl)
+library(arrow)
 library(dplyr)
 
 #### Clean data ####
@@ -34,8 +35,14 @@ breach_data <- breach_data %>%
 breach_data <- breach_data %>%
   filter(number_of_users_affected != "Missing")
 
+# Convert 'number_of_users_affected' to numeric, handling non-numeric cases
 breach_data$number_of_users_affected <- as.numeric(as.character(breach_data$number_of_users_affected))
+
+# Remove unwanted columns
+# Remove unwanted columns using dplyr::select to avoid any masking issues
+breach_data <- dplyr::select(breach_data, -c(summary, effect_on_share_price, inappropriate_remote_access))
+
 
 # Saving cleaned data
 write.csv(breach_data, file = "data/analysis_data/breach_data.csv", row.names = FALSE)
-
+write_parquet(breach_data, "data/analysis_data/breach_data.parquet")

### 2024-04-07 scripts/04-model.R

@@ -9,12 +9,150 @@
 library(tidyverse)
 library(rstanarm)
 library(bayesplot)
+library(forcats)
 library(readr)
 library(MASS)
+library(broom)
+library(vcd)
+library(caret)
 
 #### Read data ####
 # Assuming the data file is directly in the current working directory
 breach_data <- read_csv("data/analysis_data/breach_data.csv")
+
+
+#########
+calculate_breach_severity <- function(impact, credit_card_leak_1, credit_card_leak_2, ssn_leak, fraudulent_use) {
+  if (is.na(impact) || is.na(credit_card_leak_1) || is.na(credit_card_leak_2) || is.na(ssn_leak) || is.na(fraudulent_use)) {
+    return(NA)  # Return NA if any of the inputs is NA
+  } else if (impact == 'High' || fraudulent_use == 'Yes') {
+    return('High')
+  } else if ((credit_card_leak_1 == 'Yes' || credit_card_leak_2 == 'Yes' || ssn_leak == 'Yes') && impact != 'Low') {
+    # If any credit card or SSN data is leaked, and impact is not 'Low', consider it 'Medium' severity
+    return('Medium')
+  } else {
+    return('Low')
+  }
+}
+
+# Apply the function to each row in the dataframe
+breach_data$breach_severity <- mapply(calculate_breach_severity,
+                                      breach_data$aspect_of_confidentiality_integrity_availability_triad_affected,
+                                      breach_data$track_1_credit_card_details_leaked_exposed,
+                                      breach_data$track_2_credit_card_details_leaked_exposed,
+                                      breach_data$social_security_number_tax_number_leaked_exposed,
+                                      breach_data$subsequent_fraudulent_use_of_data)
+
+
+# Fit the linear regression model using the dataframe without NAs
+breach_data$breach_severity_numeric <- as.numeric(factor(breach_data$breach_severity, levels = c("Low", "Medium", "High")))
+
+# Now, let's select the independent variables and convert categorical variables to factors
+# (make sure to replace these variable names with the actual column names in your dataset)
+breach_data$organisation_size <- as.factor(breach_data$organisation_size)
+breach_data$level_of_digital_intensity <- as.factor(breach_data$level_of_digital_intensity)
+breach_data$sector <- as.factor(breach_data$sector)
[…]

### 2024-04-07 paper/paper.qmd
model section in progress

@@
 ```
-# Model {# sec-model}
+# Model {#sec-model}
+
+## Model Set-up
+
+After performing exploratory data analysis, we got a broad overview of the general trend between different variables of our dataset.
+To strengthen answers to our research question claims made earlier, we implement  regression models and take a close look at what makes an organization good at dealing with cyber threats.
+We use these models to understand how certain factors, like the size of a company or the industry, affect its ability to handle a cyberattack.
+
+**Simple Models for Numbers (Linear Regression):**
+
+We were interested in knowing how often cyberattacks happen and how bad they are based on the company's size or the kind of work it does. We use linear regression for this. 
+The basic form of a linear regression equation is
+
+$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + ... + \beta_n X_n + \varepsilon$
+
+Where:
+- $Y$ is the outcome we're interested in.
+
+- $\beta_0$ is the intercept, the starting point of our equation when all our predictors (Xs) are zero.
+
+- $\beta_1, \beta_2, ..., \beta_n$ are the coefficients. These tell us how much change we expect in $Y$ for a one-unit change in each predictor $X$, assuming all other predictors are held constant.
+
+- $X_1, X_2, ..., X_n$ are the predictor variables.
+
+- $\varepsilon$ represents the error term. It accounts for the difference between our predicted value of $Y$ and its actual value.
+
+We expected to study if bigger companies, for example, have more cyberattacks or not. In form of equation we can represent it as follows:
+
+$\text{Impact} = (\text{Base Impact}) + (\text{Effect of Company Size}) + (\text{Effect of Industry}) + \dots$
+
+**Complex Models for Yes/No Outcomes (Bayesian Logistic Regression)**:
+
+Now, what if we want to know if a company will take action after an attack, like starting an investigation?
+This is a yes/no type of question.
+We use a different method called logistic regression, which is good for when the outcome is not a number but a choice between two things.
+We make this model even better by using what we call a Bayesian approach.
+It means we use past knowledge to make better guesses.
+So, we can say:
 
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| 
+library(modelsummary)
+
+breach_severity_model <- readRDS(file = here::here("models/breach_severity_model.rds"))
+
+modelsummary(list("Linear Regression" = breach_severity_model))
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| 
 library(modelsummary)
@@
 ```
-```{r}
-breach_data <- breach_data %>% mutate(row_id = row_number())
-
-# Adjust factors in your data to match the model's training data
-breach_data <- breach_data %>%
-  mutate(country = factor(country, levels = levels(logistic_model$model$country)))
-
-# Generate predictions
-breach_predictions <- predict(logistic_model, newdata = breach_data, type = "response")
-
-# Combine the predictions with the original data
-breach_data <- breach_data %>% mutate(predicted_prob = breach_predictions)
-
-# Scatter plot with jitter
-ggplot(breach_data, aes(x = organisation_size, y = predicted_prob)) +
-  geom_jitter(alpha = 0.3) +
-  labs(x = "Organisation Size", y = "Predicted Probability of Restructuring")
-
-```
 
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
 breach_data %>%
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 # Summary statistics for numerical variables
@@
 ```
+
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
 linear_model_RQ2 <- readRDS(file = here::here("models/linear_model_RQ2.rds"))

### 2024-04-07 scripts/04-model.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Models different relations between variables such as cyberattack outcomes and organizational attributes. 
 # Author: Shivank Goel  
-# Date: [Today's Date]
+# Date: 7th April 2024
 # Contact: [Your Contact Information]
 # License: MIT
 
@@ -20,8 +20,24 @@ library(caret)
 # Assuming the data file is directly in the current working directory
 breach_data <- read_csv("data/analysis_data/breach_data.csv")
 
+# Convert categorical variables to factors
+breach_data$country <- as.factor(breach_data$country)
+breach_data$restructuring_after_attack <- as.numeric(breach_data$restructuring_after_attack == "Yes")
+breach_data$organisation_size <- as.factor(breach_data$organisation_size)
+breach_data$sector <- as.factor(breach_data$sector)
+breach_data$critical_industry <- as.factor(breach_data$critical_industry)
+breach_data$level_of_digital_intensity <- as.factor(breach_data$level_of_digital_intensity)
+breach_data$number_of_users_affected <- as.numeric(breach_data$number_of_users_affected)
 
-#########
+
+
+breach_data$critical_industry <- ifelse(breach_data$critical_industry == "Yes", 1, 0)
+breach_data$cyber_security_role <- ifelse(breach_data$cyber_security_role == "Yes", 1, 0)
+breach_data$undertook_investigation <- ifelse(breach_data$undertook_investigation == "Yes", 1, 0)
+
+
+######### Model 1
+
 calculate_breach_severity <- function(impact, credit_card_leak_1, credit_card_leak_2, ssn_leak, fraudulent_use) {
   if (is.na(impact) || is.na(credit_card_leak_1) || is.na(credit_card_leak_2) || is.na(ssn_leak) || is.na(fraudulent_use)) {
     return(NA)  # Return NA if any of the inputs is NA
@@ -47,14 +63,11 @@ breach_data$breach_severity <- mapply(calculate_breach_severity,
 # Fit the linear regression model using the dataframe without NAs
 breach_data$breach_severity_numeric <- as.numeric(factor(breach_data$breach_severity, levels = c("Low", "Medium", "High")))
 
-# Now, let's select the independent variables and convert categorical variables to factors
-# (make sure to replace these variable names with the actual column names in your dataset)
[…]

### 2024-04-07 paper/paper.qmd
added model for rq2, completed model rq1 writeup

@@
 
-4.External Factors: Factors like the industry the organization is in, its size, and its digital intensity (how much it relies on digital technology) can also impact its cyber resilience.
+4. External Factors: Factors like the industry the organization is in, its size, and its digital intensity (how much it relies on digital technology) can also impact its cyber resilience.
 
@@
 We use a different method called logistic regression, which is good for when the outcome is not a number but a choice between two things.
-We make this model even better by using what we call a Bayesian approach.
-It means we use past knowledge to make better guesses.
-So, we can say:
+
+## Specific models for Research Question
+
+**RQ1: Organizational attributes and cyberattack handling**
+
+To address the first research question about how an organization's size and business type affect its ability to handle cyber attacks, we developed a specific linear regression model.
+This model examines the relationships between various organizational attributes and the handling of cyberattacks.
+It predicts the severity of cyber breaches based on:
+
+1. Organizational Size (Medium, Small, Unknown)
+2. Digital Intensity (Low, Low-Medium, Medium-High)
+3. Sector (e.g., Public Administration and Defence, Education, Health)
+4. Country
+
+- Dependent Variable: Severity of cyber breaches.
+
+- Independent Variables: Organizational size, sector, digital intensity, and geographical location.
+
+[@tbl-modelbreachseverity] summarises the results from model.
+
+1.
+Organisation Size: The coefficients for different organization sizes (Medium, Small, Unknown) indicate how the severity of breaches changes with these categories compared to the baseline category.
+Medium (0.015) and Small (0.103) have positive coefficients, which suggests potentially higher severity compared to the baseline.
+Unknown (-0.072) indicates a lower severity than the baseline, which might reflect missing data issues or other factors.
+
+2.
+Level of Digital Intensity & Sector: These coefficients show how levels of digital intensity and different sectors affect breach severity.
+Negative coefficients, like for the public administration and defence sector (-0.581), suggest a lower severity compared to the baseline sector.
+
+3. Country: Similar to sectors, the coefficients for different countries compared to the baseline indicate variations in breach severity across geographical locations.
+
+R-squared (0.137): Indicates the proportion of variance in the dependent variable that's predictable from the independent variables.
+A value of 0.137 suggests the model explains about 13.7% of the variability in breach severity.
+
+Adjusted R-squared (0.099): It adjusts the R-squared for the number of predictors in the model. It is lower than the R-squared, which is common in models with many predictors.
+
+AIC (665.5) and BIC (746.7): These are measures for model comparison.
+
+**RQ2: Effectiveness of cybersecurity methods**
+
+**RQ3: Impact of specific business situations**
 
@@
 #| warning: false
+#| label: tbl-modelbreachseverity
+#| tbl-cap: "Linear Regression Model Summary for RQ1"
 #| message: false

### 2024-04-07 scripts/04-model.R

@@ -2,7 +2,7 @@
 # Purpose: Models different relations between variables such as cyberattack outcomes and organizational attributes. 
 # Author: Shivank Goel  
 # Date: 7th April 2024
-# Contact: [Your Contact Information]
+# Contact: shivankg.goel@mail.utoronto.ca
 # License: MIT
 
 #### Workspace setup ####
@@ -34,6 +34,20 @@ breach_data$number_of_users_affected <- as.numeric(breach_data$number_of_users_a
 breach_data$critical_industry <- ifelse(breach_data$critical_industry == "Yes", 1, 0)
 breach_data$cyber_security_role <- ifelse(breach_data$cyber_security_role == "Yes", 1, 0)
 breach_data$undertook_investigation <- ifelse(breach_data$undertook_investigation == "Yes", 1, 0)
+
+
+
+binary_columns <- c("cyber_security_role", "cyber_security_frameworks", "education_and_awareness_policy", "prevention_detection_and_recovery")
+breach_data[binary_columns] <- lapply(breach_data[binary_columns], function(x) as.numeric(x == "Yes"))
+
+# Building the Linear Regression Model
+rq2model <- lm(number_of_users_affected ~ cyber_security_role + cyber_security_frameworks + education_and_awareness_policy + prevention_detection_and_recovery + as.factor(organisation_size) + as.factor(sector) + as.numeric(level_of_digital_intensity), data = breach_data)
+
+# Summary of the Model
+summary(rq2model)
+
+
+
 
 
 ######### Model 1
@@ -240,7 +254,7 @@ saveRDS(restructuring_model, "models/restructuring_model.rds")
 saveRDS(linear_model_RQ2, "models/linear_model_RQ2.rds")
 saveRDS(linear_model_RQ3, "models/linear_model_RQ3.rds")
 saveRDS(model_investigation, "models/model_investigation.rds")
-
+saveRDS(rq2model, "models/rq2model.rds")
 
 
 

### 2024-04-08 paper/paper.qmd
made model tables

@@
    - \usepackage{float}
+   - \usepackage{booktabs}
+   - \usepackage{longtable}
+
 format: pdf
@@
 
-[@tbl-modelbreachseverity] summarises the results from model.
+[@breach_severity_model] summaries the results from model.
 
@@
 R-squared (0.137): Indicates the proportion of variance in the dependent variable that's predictable from the independent variables.
-A value of 0.137 suggests the model explains about 13.7% of the variability in breach severity.
+A value of 0.137 suggests that approximately 13.71% of the variation in breach severity is explained by the included predictors.
 
-Adjusted R-squared (0.099): It adjusts the R-squared for the number of predictors in the model. It is lower than the R-squared, which is common in models with many predictors.
+Adjusted R-squared (0.099): This value is slightly lower than the R-squared, and accounts for the number of predictors,.
+It suggests a good fit of the model considering the complexity introduced by multiple variables.
 
-AIC (665.5) and BIC (746.7): These are measures for model comparison.
+F-statistic of 3.619 and a p-value of 1.12e-06, suggests that the model demonstrates the statistical significance overall, indicating that at least some of the predictors are likely to be genuinely associated with variations in breach severity.
+
+\begin{longtable}{lcccc}
+\caption{Linear regression analysis with breach severity numeric as dependent variable} \label{tab:breach_severity_model} \\
+\toprule
+Predictor & Estimate & Std. Error & t value & Pr(>|t|) \\
+\midrule
+\endfirsthead
+
+\multicolumn{5}{c}%
+{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
+\toprule
+Predictor & Estimate & Std. Error & t value & Pr(>|t|) \\
+\midrule
+\endhead
+
+\midrule \multicolumn{5}{r}{\textit{Continued on next page}} \\
+\endfoot
+
+\endlastfoot
+
+Intercept & 2.08715 & 0.09139 & 22.837 & $<$ 2e-16 *** \\
+Organisation Size - Medium & 0.01467 & 0.08055 & 0.182 & 0.855 \\
+Organisation Size - Small & 0.10336 & 0.10093 & 1.024 & 0.306 \\
+Organisation Size - Unknown & -0.07151 & 0.07166 & -0.998 & 0.319 \\
+Level of Digital Intensity - Low & -0.29052 & 0.22750 & -1.277 & 0.202 \\
+Level of Digital Intensity - Low-Medium & 0.11865 & 0.16894 & 0.702 & 0.483 \\
+Level of Digital Intensity - Medium-High & 0.45764 & 0.15186 & 3.013 & 0.0027 ** \\
+Sector - Public Administration and Defence & -0.58112 & 0.20077 & -2.894 & 0.0040 ** \\
+Sector - Other & -0.44159 & 0.12026 & -3.672 & 0.0003 *** \\
+Sector - Arts, Entertainment and Recreation & -0.80423 & 0.20009 & -4.019 & 6.94e-05 *** \\
+Sector - Education & -0.27055 & 0.20130 & -1.344 & 0.180 \\
+Sector - Human Health Activities & -0.20620 & 0.19467 & -1.059 & 0.290 \\
+Country - Canada & -0.55895 & 0.30172 & -1.853 & 0.0646 . \\
+Country - Global & -0.41952 & 0.16428 & -2.554 & 0.0110 * \\
+Country - Japan & -0.21109 & 0.26805 & -0.788 & 0.431 \\
+Country - UK & 0.07899 & 0.30459 & 0.259 & 0.795 \\
+Country - Other & 0.13767 & 0.15593 & 0.883 & 0.378 \\
+Country - Singapore & -0.68348 & 0.30035 & -2.276 & 0.0234 * \\
+Country - South Korea & -0.74056 & 0.52226 & -1.418 & 0.157 \\
+\midrule
+Residual standard error & \multicolumn{4}{c}{0.5131 on 410 DF} \\
+Multiple R-squared & \multicolumn{4}{c}{0.1371} \\
+Adjusted R-squared & \multicolumn{4}{c}{0.09921} \\
+F-statistic & \multicolumn{4}{c}{3.619 on 18 and 410 DF} \\
+P-value & \multicolumn{4}{c}{1.12e-06} \\
+\bottomrule
+\end{longtable}
+
+**RQ3: Impact of specific business situations**
 
@@
 
-**RQ3: Impact of specific business situations**
+\begin{longtable}{lrrrr}
+\caption{Linear regression analysis with number of users affected as dependent variable} \label{tab:my_label} \\
+\toprule
+Model & Coefficients & Standard errors & T value & Sig \\
+\midrule
+\endfirsthead
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| label: tbl-modelbreachseverity
-#| tbl-cap: "Linear Regression Model Summary for RQ1"
-#| message: false
-#| 
-library(modelsummary)
+\multicolumn{5}{c}%
+{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
+\toprule
+Model & Coefficients & Standard errors & T value & Sig \\
+\midrule
+\endhead
 
-breach_severity_model <- readRDS(file = here::here("models/breach_severity_model.rds"))
+\midrule \multicolumn{5}{r}{\textit{Continued on next page}} \\
+\endfoot
 
-modelsummary(list("Linear Regression" = breach_severity_model))
+\endlastfoot
 
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-#| 
-library(modelsummary)
-
-logistic_model <- readRDS(file = here::here("models/restructuring_model.rds"))
-
-modelsummary(list("Logistic Regression" = logistic_model))
-
+Intercept & 36,687,660 & 32,842,060 & 1.117 & 0.264 \\
+Cyber Security Role & NA & NA & NA & NA \\
+Cyber Security Frameworks & -13,994,198 & 139,505,210 & -0.100 & 0.920 \\
+Education and Awareness Policy & 38,460,000 & 194,290,149 & 0.198 & 0.843 \\
+Prevention, Detection, and Recovery & NA & NA & NA & NA \\
+Organization Size - Medium & -7,347,889 & 29,571,997 & -0.248 & 0.804 \\
+Organization Size - Large & 8,946,092 & 25,705,975 & 0.348 & 0.728 \\
+Organization Size - Unknown & -14,379,354 & 28,602,186 & -0.503 & 0.615 \\
+Sector - Public Administration and Defence & 81,232,193 & 41,730,689 & 1.947 & 0.052 \\
+Sector - Other & 73,186,958 & 25,457,748 & 2.875 & 0.004 \\
+Sector - Arts, Entertainment and Recreation & 93,277,310 & 41,274,076 & 2.260 & 0.024 \\
+Sector - Education & 40,468,643 & 31,559,658 & 1.282 & 0.200 \\
+Sector - Human Health Activities & 47,359,281 & 28,602,617 & 1.656 & 0.098 \\
+Level of Digital Intensity & -28,344,216 & 9,431,846 & -3.005 & 0.003 \\
+\midrule
+R Squared & \multicolumn{4}{c}{0.093} \\
+Adjusted R Squared & \multicolumn{4}{c}{0.037} \\
+F Statistic & \multicolumn{4}{c}{1.648} \\
+Significance of F & \multicolumn{4}{c}{0.018} \\
+\bottomrule
+\end{longtable}
 ```
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-# Summary statistics for numerical variables
-summary(breach_data$number_of_users_affected)
-summary(breach_data$year)
-
-# Frequency tables for categorical variables
-table(breach_data$sector)
-table(breach_data$organisation_size)
-table(breach_data$critical_industry)
-table(breach_data$level_of_digital_intensity)
-table(breach_data$country)
-table(breach_data$cyber_security_role)
-table(breach_data$cyber_security_frameworks)
-table(breach_data$education_and_awareness_policy)
-table(breach_data$policy)
-table(breach_data$prevention_detection_and_recovery)
-table(breach_data$detector)
-table(breach_data$restructuring_after_attack)
-table(breach_data$bribe_ransom_paid)
-table(breach_data$free_identity_or_credit_theft_monitoring)
-table(breach_data$additional_disclosure_of_information)
-table(breach_data$overall_nature_of_attack)
-table(breach_data$attack_type)
-table(breach_data$attacker)
-table(breach_data$attack_vector)
-table(breach_data$impact_on_data)
-# ... continue for other categorical variables as needed
-
 # Histogram for a continuous variable (e.g., number_of_users_affected)

### 2024-04-08 scripts/04-model.R

@@ -35,21 +35,6 @@ breach_data$critical_industry <- ifelse(breach_data$critical_industry == "Yes", 
 breach_data$cyber_security_role <- ifelse(breach_data$cyber_security_role == "Yes", 1, 0)
 breach_data$undertook_investigation <- ifelse(breach_data$undertook_investigation == "Yes", 1, 0)
 
-
-
-binary_columns <- c("cyber_security_role", "cyber_security_frameworks", "education_and_awareness_policy", "prevention_detection_and_recovery")
-breach_data[binary_columns] <- lapply(breach_data[binary_columns], function(x) as.numeric(x == "Yes"))
-
-# Building the Linear Regression Model
-rq2model <- lm(number_of_users_affected ~ cyber_security_role + cyber_security_frameworks + education_and_awareness_policy + prevention_detection_and_recovery + as.factor(organisation_size) + as.factor(sector) + as.numeric(level_of_digital_intensity), data = breach_data)
-
-# Summary of the Model
-summary(rq2model)
-
-
-
-
-
 ######### Model 1
 
 calculate_breach_severity <- function(impact, credit_card_leak_1, credit_card_leak_2, ssn_leak, fraudulent_use) {
@@ -109,7 +94,22 @@ summary(breach_severity_model)
 
 ############
 
+#####
 
+binary_columns <- c("cyber_security_role", "cyber_security_frameworks", "education_and_awareness_policy", "prevention_detection_and_recovery")
+breach_data[binary_columns] <- lapply(breach_data[binary_columns], function(x) as.numeric(x == "Yes"))
+
+# Building the Linear Regression Model
+breach_data$organisation_size <- factor(breach_data$organisation_size, 
+                                        levels = c("Small", "Medium", "Large", "Unknown"),
+                                        labels = c("Small", "Medium", "Large", "Unknown"))
+
+rq2model <- lm(number_of_users_affected ~ cyber_security_role + cyber_security_frameworks + education_and_awareness_policy + prevention_detection_and_recovery + as.factor(organisation_size) + as.factor(sector_simplified) + as.numeric(level_of_digital_intensity), data = breach_data)
+
+# Summary of the Model
+summary(rq2model)
+
+####
 
 ######### Model 2
 

### 2024-04-10 paper/paper.qmd
paper write up done

@@
 bibliography: references.bib
+toc: true
 ---
@@
 
-In today's digital world, every click, every online transaction, and every shared piece of data is a potential entry point for cyber threats.
-Cybercrime is not just evolving, rather it is expanding at an alarming rate, with both the frequency and severity of attacks rising every year.
+In today's digital world, every click, every online transaction, and every shared piece of data poses an entry point for cyber threats.
+Cybercrime is not just growing quickly, rather it is turning out to be more of a serious concern with both the frequency and severity of attacks rising every year.
 The purpose of these attacks is to harm companies and organizations financially, however, in some cases these attacks can have military or political purposes.
 “According to a report published by the Identity Theft Resource Center (ITRC), a record number of 1862 data breaches occurred in 2021 in the US.
-Sectors like healthcare, finance, business, and retail are the most commonly attacked, impacting millions of Americans every year” (https://www.upguard.com/blog/biggest-data-breaches-us)
+Sectors like healthcare, finance, business, and retail are the most commonly attacked, impacting millions of Americans every year” [@UpGuard2024]
 
@@
 Half are ‘very satisfied’ with their technology capabilities in key cybersecurity areas.
-More than 30% of companies don’t consistently follow what should be standard practices of cyber defence.” (https://www.pwc.com/us/en/services/consulting/cybersecurity-risk-regulatory/library/global-digital-trust-insights.html)
+More than 30% of companies don’t consistently follow what should be standard practices of cyber defence.” [@PwC2023]
 
 Therefore, in response to such attacks, there is a need for a plan, that not just keep the intruder or hackers out but also quickly alert if an attack does happen.
-Our study looks at cyber resilience, which is “ the ability to anticipate, withstand, recover from, and adapt to adverse conditions, stresses, attacks, or compromises on systems that use or are enabled by cyber resources.” as defined by National Institute of Standards and Technology (https://csrc.nist.gov/glossary/term/cyber_resiliency).
-To thoroughly analyze our study, we break it into three major research questions :
+Our study looks at cyber resilience, which is “ the ability to anticipate, withstand, recover from, and adapt to adverse conditions, stresses, attacks, or compromises on systems that use or are enabled by cyber resources.” as defined by National Institute of Standards and Technology [@NISTGlossary2023].
+We break our study into three major research questions, which forms the thesis of our study:
 
-**RQ1**: How do things like the size of the company and the type of business it does affect its ability to handle cyber attacks?
+**RQ1**: How do organizational factors such as size and sector, influence the severity of cyber breaches experienced by companies?
 
-**RQ2**: Which methods or strategies used by companies work best to reduce the damage from cyber attacks?
+**RQ2**: Which cybersecurity strategies, including frameworks, policies, and preventive measures, are the most effective one, in reducing the damage caused by cyber attacks?
 
-**RQ3**: How does the business’s specific situation, like its industry or how much it relies on digital tools, change the impact of cyber attacks on it?
+**RQ3**: In what ways do the industry type and digital dependence of a business affect the overall impact of a cyber attack, in terms of preserving confidentiality, integrity and availability of data?
 
 The estimand of our study is the measurable effect of specific characteristics of an organization including size, sector and digital intensity on their cyber resilience.
-As a key finding, our regression models reveal factors such as organizational size, sector, and digital intensity significantly influence an organization's cyber resilience posture.For example, larger companies often have stronger defenses against cyber attacks, and on the other hand, companies that use a lot of digital technology in their work have different levels of protection.
+As a key finding, our regression models reveal factors such as organizational size, sector, and digital intensity significantly influence an organization's cyber resilience.
+For example, larger companies often have stronger defenses against cyber attacks, and on the other hand, companies that use a lot of digital technology in their work have different levels of protection.
 
@@
 
-The remainder of this paper is structured as follows: @sec-data provides an overview of our methodology, including the data collection process and the analytical techniques used to explore the dataset of cyber attacks.
-We provide the background and overview of the study in @sec-background.
+The remainder of this paper is structured as follows:  We provide the background and overview of the study in @sec-background.
+@sec-data provides an overview of our methodology, including the data collection process and the analytical techniques used to explore the dataset of cyber attacks.
 @sec-model presents the regression models, discussing how we applied these models to understand the impact of various factors like organizational size, sector, and digital intensity on cyber resilience, @sec-results displays the interpretations of the model alongside other findings from analyzing the data, and @sec-discussion provides a discussion on the implications of the findings as well as the weaknesses of this paper and its next steps for further study on this subject.
@@
 Our analysis is based on sampling of 514 data breaches and ransomware attacks spanning over 14 years from 2004 to 2019.
-The dataset was obtained from the website of University of Queensland, and was prepared and compiled by researchers Tsen, Elinor, Ko, Ryan, and Slapnicar, Sergeja https://espace.library.uq.edu.au/view/UQ:dfe5027  at the University of Queensland.
+The dataset was obtained from the website of University of Queensland, and was prepared and compiled by researchers Tsen, Elinor, Ko, Ryan, and Slapnicar, Sergeja [@tsen_dataset_2020] at the University of Queensland.
 The data is thorough and encompasses a wide range of cyber attack incidents.
@@
 In order to observe the trend of number of cyberattacks over a span of years, from 2004 to 2019,
-we plotted a line graph [@fig-attacks].
+we plotted a line graph @fig-attacks.
 It is evident from the plot that the frequency of attacks has fluctuated over the years, with a peak in 2017.
@@
 #| label: fig-sector
-#| fig-cap: Bar Plot of Sector
+#| fig-cap: Bar Plot of Count of Cyber Attacks by Sector
 #| echo: false
@@
 
-[@fig-overviewofattack] is a creative visualization that effectively depicts the distribution and comparison of cyber attacks across various countries, with a specific emphasis on the United States.
+@fig-overviewofattack is a creative visualization that effectively depicts the distribution and comparison of cyber attacks across various countries, with a specific emphasis on the United States.
 It combines a stacked bar chart for multiple countries and a line plot for the USA allowing for a dual-axis comparison due to the disproportionate number of attacks in the USA compared to other countries.
@@
 ```
-[@fig-attacktype] represents a stacked area chart, with each colored layer representing a different type of attack, allowing for an easy comparison of their occurrences over time.
+@fig-attacktype represents a stacked area chart, with each colored layer representing a different type of attack, allowing for an easy comparison of their occurrences over time.
 It is clear that some attack types, like installed malware, show peaks and troughs, possibly depicting the nature of cyber threats and security measures.
@@
 
-## Specific models for Research Question
+## Specific models for Research Questions
 
@@
 
-[@breach_severity_model] summaries the results from model.
+Table \ref{breach_severity_model} summaries the results from model.
 
 1.
-Organisation Size: The coefficients for different organization sizes (Medium, Small, Unknown) indicate how the severity of breaches changes with these categories compared to the baseline category.
-Medium (0.015) and Small (0.103) have positive coefficients, which suggests potentially higher severity compared to the baseline.
-Unknown (-0.072) indicates a lower severity than the baseline, which might reflect missing data issues or other factors.
+Organisation Size: The coefficients for different organization sizes (Medium, Small, Unknown) indicate how the severity of breaches changes with these categories compared to the baseline category.As compared to a baseline (Large), Medium (0.015) and Small (0.103) sizes have positive coefficients, suggesting potentially higher breach severity.
+Unknown (-0.072) indicates lower severity.
 
 2.
-Level of Digital Intensity & Sector: These coefficients show how levels of digital intensity and different sectors affect breach severity.
-Negative coefficients, like for the public administration and defence sector (-0.581), suggest a lower severity compared to the baseline sector.
+Level of Digital Intensity & Sector: These coefficients compare each level and sector against baselines, "Medium-High" for digital intensity and "Finance and Insurance" for the sector.
+The public administration and defence sector (-0.581) shows a lower severity.
 
-3. Country: Similar to sectors, the coefficients for different countries compared to the baseline indicate variations in breach severity across geographical locations.
+3. Country: Coefficients for countries are relative to the USA (reference level). They indicate variations in breach severity across geographical locations.
 
-R-squared (0.137): Indicates the proportion of variance in the dependent variable that's predictable from the independent variables.
-A value of 0.137 suggests that approximately 13.71% of the variation in breach severity is explained by the included predictors.
+R-squared (0.137): It indicates the proportion of variance in the dependent variable that's predictable from the independent variables.
+A value of 0.137 suggests that approximately 13.71% of the variation in breach severity is explained by the model.
 
-Adjusted R-squared (0.099): This value is slightly lower than the R-squared, and accounts for the number of predictors,.
+Adjusted R-squared (0.099): This value is slightly lower than the R-squared, and accounts for the number of predictors.
 It suggests a good fit of the model considering the complexity introduced by multiple variables.
@@
 \begin{longtable}{lcccc}
-\caption{Linear regression analysis with breach severity numeric as dependent variable} \label{tab:breach_severity_model} \\
+\caption{Linear regression analysis with breach severity numeric as dependent variable} \label{breach_severity_model} \\
 \toprule
@@
 
-**RQ3: Impact of specific business situations**
+**RQ2: Effectiveness of cybersecurity methods**
 
-**RQ2: Effectiveness of cybersecurity methods**
+To answer our second research question regarding the most effective strategies and methodsused by organizations to reduce the damage from cyberattacks, we constructed a linear regression model.
+This model shows the number of users impacted by cyber incidents, considering a relationship between cybersecurity strategies and organizational attributes :
+
+1. Implementation of Cybersecurity Frameworks
+2. Adoption of Education and Awareness Policies
+3. Execution of Prevention, Detection, and Recovery Measures
+4. Organizational Size (Medium, Large, Unknown)
+5. Sector Specificity
+6. Level of Digital Intensity
+
+- Dependent Variable: Number of users affected by cyberattacks.
+
+- Independent Variables: Cybersecurity measures, organizational size, sector, and digital intensity level.
+
+Table \ref{tab:my_label} depicts the results derived from the model.
+
+1. Cybersecurity Frameworks(-13,994,198) - The negative coefficient implies that cybersecurity frameworks might reduce the number of users affected.
+
+2.
+Education and Awareness Policies (38,460,000): A positive coefficient suggests a correlation between the implementation of these policies and an increase in the number of users affected.
+However, the insignificance of this result (p = 0.843) implies caution in interpreting this finding.
+
+3.
+Organizational size - Medium size (-7,347,889, p = 0.804), Large size (8,946,092, p = 0.728), and Unknown size (-14,379,354, p = 0.615): The model suggests no significant effect of organization size on the number of users affected, as all the p-values greater than 0.05.
+
+4.
+Sector Specifics: The model reveals significant positive relationships for 'Other' sectors (73,186,958, p = 0.004) and 'Arts, Entertainment, and Recreation' (93,277,310, p = 0.024), indicating these areas are likely to witness a higher number of users affected by cyberattacks.
+
+5. Digital Intensity Level (-28,344,216): A negative coefficient with p = 0.003 shows that increased digital intensity is associated with a greater number of users affected.
+
+The model's R-squared (0.093) signifies that about 9.3% of the variance in the number of users affected is explained by the model, while the adjusted R-squared (0.037) factors in the number of predictors.
+
+An F-statistic of 1.648 along with a p-value of 0.018 shows the model's overall significance, indicating that specific factors, particularly sector and digital intensity level, are helpful in understanding the scale of cyberattack impacts.
 
@@
 \end{longtable}
-```
+Note: NA indicates that the variable was not applicable or not included in the model.
+
+Also, the trend of our dependent variable `number_of_users_affcted` in RQ2 model, can be seen in @fig-numofusersaffected
 
 ```{r}
+#| label: fig-numofusersaffected 
+#| fig-cap: Number of Users Affected Over Years16
 #| echo: false
-#| eval: true
 #| warning: false
-#| message: false
 
@@
   geom_line() +
-  labs(title = "Number of Users Affected Over Years",
-       x = "Year",
+  labs(x = "Year",
        y = "Total Number of Users Affected") +
@@
 ```
+**RQ3: Impact of Organizational Attributes on Cyberattack Consequences**
 
-```{r}
-# Histogram for a continuous variable (e.g., number_of_users_affected)
-ggplot(breach_data, aes(x = number_of_users_affected)) +
-  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
-  labs(title = "Histogram of Number of Users Affected", x = "Number of Users Affected", y = "Frequency")
+To analyze  how specific attributes of a business, like its industry or dependency on digital tools, impact the consequences of cyberattacks, we used a multinomial logistic regression model.
+This model helps to determine the relationship between organizational characteristics and the level of impact on data, classified as 'Low', 'Medium', or 'High'.
 
-```
+1. Organizational Size (Medium, Small, Unknown) – with 'Large' as the reference category.
+2. Digital Intensity (Low, Low-Medium, Medium-High) – with 'High' as the reference category.
+3. Sector – 'Public Administration and Defence', 'Other', 'Arts, Entertainment and Recreation', etc., with 'Finance and Insurance' as the reference category.
+4. Country – including Canada, Global, Japan, etc., with 'USA' as the reference category.
+5. Dependent Variable: The level of impact on data (Low, Medium, High).
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+- Independent Variables: Organizational size, digital intensity, sector, country.
 
-linear_model_RQ2 <- readRDS(file = here::here("models/linear_model_RQ2.rds"))
-# Plotting diagnostics for the linear regression model (Example: linear_model_RQ2)
-ggplot(breach_data, aes(x = year, y = number_of_users_affected)) +
-  geom_point() +
-  geom_smooth(method = "lm", color = "blue") +
-  labs(title = "Linear Regression of Number of Users Affected Over Years", 
-       x = "Year", 
-       y = "Number of Users Affected")
+Table \ref{tab:impact_multinom_model} displays the model outcomes.
 
-# Assuming breach_data is your dataset and you have already created models named linear_model_severity and linear_model_investigations
+1.
+Organizational Size: Medium (-0.793, p = Medium) and Small (0.308, p = Small) sizes show different degrees of impact compared to the baseline 'Large' size, suggesting the size of the organization plays a role in the impact level of cyberattacks.
 
-# Partial regression plot for breach severity with a specific predictor (e.g., 'organisation_size')
-# install.packages("car") # Uncomment if the car package is not installed
+2.
+Level of Digital Intensity: A higher digital intensity has a correlation with a greater likelihood of 'High' impact breaches, as indicated by the positive coefficients for 'Low-Medium' (2.037) and 'Medium-High' (3.073) intensity levels.
 
-```
+3.
+Sector Specifics: Significant variations are observed across sectors.
+For example, 'Public Administration and Defence' (-1.907, p = Public Administration) and 'Arts, Entertainment and Recreation' (-4.116, p = Arts) sectors show a specific pattern in their vulnerability to cyberattacks.
 
-```{r regression-table, results='asis', echo=FALSE}
+4.
+Geographical Location: The impact of cyberattacks also differs based on geographical location.
+Canada (-14.948, p = Canada) and Singapore (-15.594, p = Singapore) shows a significant differences in the 'High' impact category, depiciting regional variations in cyberattack severity.
 
-library(broom)
-library(stargazer)
-model_investigation <- readRDS(file = here::here("models/model_investigation.rds"))
+The coefficients represent log odds ratios, where a positive value implies a higher likelihood of the respective impact level compared to the baseline ('High' impact).
 
-model_investigation_frame <- augment(model_investigation)
+\footnotesize
+\begin{longtable}{| p{4cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
+\caption{Multinomial logistic regression analysis with impact on data as dependent variable} \label{tab:impact_multinom_model} \\
+\toprule
+Predictor & Coefficients (Medium) & Coefficients (High) & Std. Errors (Medium) & Std. Errors (High) \\
+\midrule
+\endfirsthead
 
-ggplot(model_investigation_frame, aes(x = organisation_size, y = .resid)) +
-  geom_point() +
-  geom_hline(yintercept = 0, linetype = "dotted", color = "grey") +
-  theme_classic() +
-  labs(y = "Residuals", x = "Organisation Size")
+\multicolumn{5}{c}%
+{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
+\toprule
+Predictor & Coefficients (Medium) & Coefficients (High) & Std. Errors (Medium) & Std. Errors (High) \\
+\midrule
+\endhead
 
-ggplot(model_investigation_frame, aes(x = .resid)) +
-  geom_histogram(binwidth = 1) +
-  theme_classic() +
-  labs(y = "Number of occurrences", x = "Residuals")
+\midrule \multicolumn{5}{r}{\textit{Continued on next page}} \\
+\endfoot
 
-```
-$\begin{table}[!htbp] \centering 
-  \caption{Linear regression analysis with dependent variable} 
-  \label{} 
-\begin{tabular}{@{\extracolsep{5pt}}lc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-2} 
-\\[-1.8ex] & undertook\_investigation \\ 
-\hline \\[-1.8ex] 
- critical\_industry & $-$0.03 (0.12) \\ 
-  organisation\_sizeMedium & $-$0.04 (0.06) \\ 
-  organisation\_sizeSmall & 0.07 (0.08) \\ 
-  organisation\_sizeUnknown & $-$0.03 (0.06) \\ 
-  level\_of\_digital\_intensityLow & $-$0.03 (0.62) \\ 
-  level\_of\_digital\_intensityLow-Medium & $-$0.29 (0.67) \\ 
-  level\_of\_digital\_intensityMedium-High & $-$0.31 (0.66) \\ 
-  sectorAdministrative and support service & $-$0.62 (0.72) \\ 
-  sectorAdvertising and other business services & $-$0.48 (0.66) \\ 
-  sectorArts, entertainment and recreation & $-$0.14 (0.12) \\ 
-  sectorChemicals and chemical products & 0.20 (0.37) \\ 
-  sectorComputer, electronic and optical products & 0.19 (0.23) \\ 
-  sectorConstruction & $-$0.61 (0.45) \\ 
-  sectorEducation & $-$0.11 (0.18) \\ 
-  sectorElectrical equipment & $-$0.23 (0.51) \\ 
-  sectorElectricity, gas, steam and air conditioning & $-$0.56 (0.34) \\ 
-  sectorFinance and insurance & $-$0.09 (0.63) \\ 
-  sectorFood products, beverages and tobacco & $-$0.06 (0.26) \\ 
-  sectorHuman health activities & $-$0.14 (0.22) \\ 
-  sectorIT and other information services & $-$0.29 (0.63) \\ 
-  sectorLegal and accounting activities & $-$0.63 (0.69) \\ 
-  sectorMachine equipment & $-$0.08 (0.24) \\ 
-  sectorPharmaceutical products & $-$0.27 (0.38) \\ 
-  sectorPublishing, audiovisual and broadcasting & $-$0.12 (0.22) \\ 
-  sectorResidential care and social work activities & $-$0.18 (0.24) \\ 
-  sectorScientific research and development & $-$0.37 (0.68) \\ 
-  sectorTelecommunications & $-$0.40 (0.65) \\ 
-  sectorTextiles, wearing apparel and leather &  \\ 
-  sectorTransportation storage & 0.06 (0.31) \\ 
-  sectorWholesale, retail trade and repair &  \\ 
-  countryChina & 0.88 (0.51) \\ 
-  countryFrance & $-$0.17 (0.51) \\ 
-  countryGermany & $-$0.01 (0.40) \\ 
-  countryGlobal & 0.28 (0.29) \\ 
-  countryHong Kong & 0.56 (0.57) \\ 
-  countryIndia & $-$0.33 (0.51) \\ 
-  countryJapan & 0.13 (0.35) \\ 
-  countryNorway & 0.04 (0.50) \\ 
-  countryPhilippines & $-$0.15 (0.51) \\ 
-  countryQatar & $-$0.33 (0.51) \\ 
-  countryRussia & $-$0.12 (0.41) \\ 
-  countrySingapore & 0.83$^{*}$ (0.36) \\ 
-  countrySouth Africa & 0.85 (0.51) \\ 
-  countrySouth Korea & 0.53 (0.36) \\ 
-  countryTurkey & 0.85 (0.51) \\ 
-  countryUAE & 0.60 (0.57) \\ 
-  countryUK & 0.35 (0.31) \\ 
-  countryUSA & 0.19 (0.26) \\ 
-  Constant & 0.45 (0.70) \\ 
- \hline \\[-1.8ex] 
-Observations & 513 \\ 
-R$^{2}$ & 0.15 \\ 
-Adjusted R$^{2}$ & 0.07 \\ 
-Residual Std. Error & 0.43 (df = 465) \\ 
-F Statistic & 1.80$^{**}$ (df = 47; 465) \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.05; $^{**}$p$<$0.01; $^{***}$p$<$0.001} \\ 
-\end{tabular} 
-\end{table} $
+\endlastfoot
+
+Intercept & 1.305595 & 1.442978 & 0.4671905 & 0.4506995 \\
+Organisation Size - Medium & -0.7934138 & 0.3656129 & 0.3944130 & 0.3923776 \\
+Organisation Size - Small & 0.308093 & 2.674162 & 0.8033055 & 0.7916091 \\
+Organisation Size - Unknown & -0.8242923 & 0.1584841 & 0.3454811 & 0.3649247 \\
+Level of Digital Intensity - Low & 0.6370277 & 1.6435254 & 0.7712827 & 0.6704278 \\
+Level of Digital Intensity - Low-Medium & 2.037751 & 1.534643 & 0.8804760 & 0.8738375 \\
+Level of Digital Intensity - Medium-High & 2.085403 & 3.073733 & 0.8548181 & 0.8186515 \\
+Sector - Public Administration and Defence & -1.907019 & -5.743329 & 1.076650 & 1.186423 \\
+Sector - Other & -1.499581 & -1.889906 & 0.5814518 & 0.5479577 \\
+Sector - Arts, Entertainment and Recreation & -4.116223 & -5.118463 & 1.085765 & 1.012404 \\
+Sector - Education & -1.795158 & -4.288260 & 1.047509 & 1.174420 \\
+Sector - Human Health Activities & -1.850178 & -3.236125 & 1.007041 & 1.004086 \\
+Country - Canada & 0.03118258 & -14.94830280 & 1.335552e+00 & 1.259128e-05 \\
+Country - Global & -0.7303029 & -0.9285532 & 0.7006000 & 0.6770341 \\
+Country - Japan & 0.8169932 & -0.8046899 & 1.265776 & 1.554310 \\
+Country - UK & -0.6114617 & 0.2831945 & 1.500414 & 1.192295 \\
+Country - Other & -0.9633399 & 0.2562872 & 0.7914678 & 0.6999920 \\
+Country - Singapore & -0.7333256 & -15.5943691 & 1.251740e+00 & 3.353692e-06 \\
+Country - South Korea & -12.880809 & -1.007852 & 488.290745 & 1.410714 \\
+
+\bottomrule
+\end{longtable}
+
+# Results {#sec-results}
+
+\normalsize
+
+## Statistical Results
+
+In this section, we discuss the outcomes from our statistical analysis about cyberattack impacts on organizations and model results.
+The findings are discussed across three research questions (RQs), where each question aimed at understanding different aspects of cyberattack consequences and responses.
+
+**Summary Statistics** : Most of the organizations we looked at are large ones as they make up about 64% of our data.
+This could mean that big companies are more often the targets of cyberattacks, or maybe the big institutes just report these incidents more because they are more noticeable and rely a lot on computers and the internet.
+We also see that hospitals and schools get attacked a lot.
+Since they have a lot of sensitive information, hackers target them more.
+
+Our result matches with earlier discussed PWC 2024 Global Digital Trust Insights report that more than 30% of companies don’t consistently follow what should be standard practices of cyber defence.
+Based on our data we also found that only about 27% of these organizations have someone in charge of keeping their computers safe, and only about 36% follow a set of rules to protect themselves online.
+This makes us think that maybe not all companies are ready to defend themselves against cyberattacks.
+
+When we look at how they try to stop attacks, almost half of them are only doing the very basic things. This tells us that many organizations might need to do more to keep their data safe.
+
+**Cyberattacks Over Time** : The surge in cyberattacks observed in 2017 @fig-attacks might be linked to various factors, including an increase in attack surfaces, the prevalence of more sophisticated attack methods, or a better reporting system that has led to higher incidents.
+This upward trend could signal the need for enhanced security measures and more proactive threat detection systems.
+
+**Sector and Country Analysis**: The disparity in the frequency of attacks among sectors @fig-sector suggests different levels of cyber hygiene or different attacker motivations, such as financial gain or data theft, which are particularly high in sectors dealing with personal data.
+The dominance of attacks in the USA, alongside high frequencies in other countries too @fig-overviewofattack, might reflect not only the attractive target that the USA presents but also the global disparity of cyber threats.
+
+**Attack Types**: The attack types distribution @fig-attacktype indicates a strong nature of cyber threats, with installed malware and web compromises becoming increasingly common.
+This pattern could be might indicate that attackers have increased the use of more sophisticated tools and thus, there is a need for businesses to focus on securing web interfaces and being cautious against malwares.
+
+## Model Results
+
+**RQ1: Organizational Attributes and Cyberattack Handling**
+
+The model about how organizational size and sector influence cyberattack handling shows significant sector-specific effects.
+Public administration and defence (-0.581, p < 0.005), Other (-0.441, p < 0.001), and Arts, Entertainment and Recreation (-0.804, p < 0.001) sectors show a lower breach severity score, which indicates a better handling or lower impact of cyberattacks compared to the reference sector, Finance and Insurance.
+Conversely, the effect of organizational size is not statistically significant, further telling us that size alone does not reflect an organization's capacity to handle cyberattacks effectively.
+
+**RQ2: Cybersecurity Methods Effectiveness**
+
+Regarding the effectiveness of cybersecurity methods, the model reveals a complex picture.
+The negative coefficient for Cyber Security Frameworks (-13,994,198) suggests that such frameworks might reduce the number of users affected.
+However, the positive coefficient for Education and Awareness Policy (38,460,000), though not statistically significant (p = 0.843), may require a cautious interpretation, as it suggests a possible correlation with an increase in the number of users affected.
+
+**RQ3: Specific Organizational Contexts**
+
+For RQ3, the multinomial logistic regression model depicts the variable impact levels of cyberattacks across different organization sizes, sectors, and countries.
+Precisely, sectors such as Public Administration and Defence (coefficients ranging from -1.907 to -5.743) and countries like Canada and Singapore show varying degrees of vulnerability to high-impact breaches.
+
+# Discussion {#sec-discussion}
+
+## Findings
+Our study finds that larger organizations report cyberattacks more often, which might be due to their larger online operations and valuable information.
+Health and educational institutions are also common targets, probably because they hold sensitive personal data.
+The rising concern is that only about one in three organizations has a dedicated cyber security role or follows a cyber security framework.
+This suggests that many are not as prepared as they could be.
+Our models reveal that the presence of cybersecurity frameworks might be linked to a lower number of affected users, suggesting their potential protective effect.
+However, the implementation of education and awareness policies did not show a significant impact on reducing the number of users affected by cyberattacks.
+
+The models also indicate that sectors such as arts, entertainment, and recreation, alongside 'Other' sectors, tend to see a higher number of users affected, showing possible sector-specific vulnerabilities.
+Increased digital intensity correlates with a greater number of affected users, which implies that as organizations become more digitally dependent, the potential impact of cyber incidents also rises.
+
+## Limitations and Bias
+
+The analysis we performed hold many limitations.
+A major challenge  we encountered was the high number of missing values within the dataset, which can introduce bias and limit the accuracy of our findings.
+For example, missing information may lead to under-representation of certain attack types or organizational responses, resulting in skewness of our understanding regarding the cyber threats.
+
+The other obstacle was that the dataset contained just one numerical column (number of users affected), which provided a limited quantitative perspective.
+
+Apart from this, the dataset reflected a biasness as it was more centered on cyber attacks from U.S..
+Cybersecurity is a global issue, and the dominance of U.S.
+data could potentially overshadow the experiences of organizations in other regions.
+This geographical issues highlights the need for a more diverse compilation of data to study cyber risks more effectively.
+
+These limitations suggests that there is the necessity for a more extensive dataset to study trends in cybercrime.
+
+The weaknesses of the dataset are:
+
+- The dataset spans over 15 years and was from 2004 to 2020, however, cyber threats are growing rapidly. Data from earlier years may not accurately reflect current trends and vulnerabilities.
+
+- Not all organizations may report cyber incidents with the same transparency.
+This may result in a dataset that does not fully capture the extent of cyber threats across different sectors or sizes of organizations.
+
+- There might be some years when more number of cyber incidents may be reported than other, leading to biasness in the study.
+
+## Future Research
+
+Our study provides a foundational understanding of cyberattack patterns and their relationship with organizational factors and characteristics.
+However, it also open doors for future research in this area.
+For example, further investigation into the specific types of cybersecurity frameworks and their effectiveness across different sectors could help us to learn better hoe to deal with cyber attacks.
+It would also be beneficial to examine the role of cybersecurity education in different organizational cultures and its impact on reducing cyber risks.
+
+It would be valuable if future research is conducted considering more geographic data and more multiple years to understand the long-term effect of preventive measures undertook by companies.Research on the global variations in cyberattacks, specifically focusing on countries outside the USA can be more helpful in providing accurate findings.
 

### 2024-04-10 scripts/04-model.R

@@ -7,17 +7,13 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
-library(bayesplot)
 library(forcats)
 library(readr)
+library(nnet)
 library(MASS)
 library(broom)
-library(vcd)
-library(caret)
 
 #### Read data ####
-# Assuming the data file is directly in the current working directory
 breach_data <- read_csv("data/analysis_data/breach_data.csv")
 
 # Convert categorical variables to factors
@@ -34,6 +30,7 @@ breach_data$number_of_users_affected <- as.numeric(breach_data$number_of_users_a
 breach_data$critical_industry <- ifelse(breach_data$critical_industry == "Yes", 1, 0)
 breach_data$cyber_security_role <- ifelse(breach_data$cyber_security_role == "Yes", 1, 0)
 breach_data$undertook_investigation <- ifelse(breach_data$undertook_investigation == "Yes", 1, 0)
+
 
 ######### Model 1
 
@@ -79,37 +76,24 @@ breach_data$country_simplified <- breach_data$country %>%
 # Check unique values in the sector variable
 unique_sectors <- unique(breach_data$sector)
 
-# Simplify the sector variable based on significance (you will need domain knowledge or statistical tests to decide which ones to keep)
+# Simplify the sector variable based on significance 
 significant_sectors <- c("Finance and insurance", "IT and other information services", "Health activities", "Public administration and defence", "Other")
 breach_data$sector_simplified <- breach_data$sector %>% 
   fct_lump(length(significant_sectors)) %>% 
   fct_relevel(significant_sectors)
 
-# Next, we fit the linear regression model
-# (you might need to modify this formula based on the actual predictors you wish to include)
+# Fit the linear regression model
 breach_severity_model <- lm(breach_severity_numeric ~ organisation_size + level_of_digital_intensity + sector_simplified + country_simplified, data = breach_data)
 
-# Finally, let's look at the summary of the model to interpret the results
+# Summary of the model
 summary(breach_severity_model)
+
+levels(breach_data$country_simplified)
+
 
 ############
 
-#####
 
-binary_columns <- c("cyber_security_role", "cyber_security_frameworks", "education_and_awareness_policy", "prevention_detection_and_recovery")
[…]

### 2024-04-10 paper/paper.qmd
testing done, llm usage updates, unwanted files removed

@@
 ---
-title: "TODO"
-subtitle: "TODO"
+title: "Analyzing How Organization Size, Sector, and Cybersecurity Strategies Influence Impacts on Cyberattacks"
+subtitle: "Based on Data Collected From the University of Queensland"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Undoubtedly as technology is progressing and we are stepping more closer to the digital world, the digital threats are also increasingly rapidly.
+This studys discusses the influence of organizational size, sector, and cybersecurity strategies on cyberattack impacts.
+Utilizing a dataset from the University of Queensland about data breaches and ransomware attacks from 2004 to 2019, we analyze the trends and correlations and study how different factors contribute to the severity and frequency of these incidents.
+Our analysis reveals key insights: larger organizations report more cyber incidents, sectors like healthcare and education are frequent targets, and cybersecurity frameworks significantly help in reducing the user impact.
+These findings highlight critical areas for how organizations can improve their security and protect their data.
 header-includes:

### 2024-04-10 scripts/03-test_data.R

@@ -1,70 +1,0 @@
-#### Preamble ####
-# Purpose: Data Validation and Testing for Mortality and Air Quality Data Analysis
-# Authors: Navya Hooda, Shivank Goel, Vanshika Vanshika
-# Date: 12th February 2024
-# Contact: shivankg.goel@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: tidyverse, testthat
-
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
-
-
-#Unit Testing Reference:https://testthat.r-lib.org/reference/test_that.html 
-
-# Load the datasets
-cleaned_air_data <- read.csv("data/analysis_data/cleaned_air_data.csv")
-cleaned_avg_pm25_conc <- read.csv("data/analysis_data/cleaned_avg_pm25_conc.csv")
-cleaned_mortality_data <- read.csv("data/analysis_data/cleaned_file.csv")
-
-#### Test data ####
-
-# Test the structure and type of the air quality data
-test_that("Air quality data columns are of the correct type", {
-  expect_type(cleaned_air_data$year, "integer")
-  expect_type(cleaned_air_data$health_risk, "character")
-  expect_type(cleaned_air_data$original_value, "double")
-})
-
-# Test the structure and type of the PM2.5 concentration data
-test_that("PM2.5 concentration data columns are of the correct type", {
-  expect_type(cleaned_avg_pm25_conc$years, "integer")
-  expect_type(cleaned_avg_pm25_conc$provincial_average, "double")
-  expect_type(cleaned_avg_pm25_conc$x10th_percentile, "double")
-  expect_type(cleaned_avg_pm25_conc$x90th_percentile, "double")
-})
-
-# Test the structure and type of the mortality data
-test_that("Mortality data columns are of the correct type", {
-  expect_type(cleaned_mortality_data$year, "integer")
-  expect_type(cleaned_mortality_data$cause, "character")
-  expect_type(cleaned_mortality_data$ranking, "integer")
-  expect_type(cleaned_mortality_data$total_deaths, "integer")
-})
-
-# Test that the years are within the expected range
-test_that("Year values are within the expected range", {
-  expect_true(all(cleaned_air_data$year >= 2012 & cleaned_air_data$year <= 2022))
-  expect_true(all(cleaned_avg_pm25_conc$years >= 2012 & cleaned_avg_pm25_conc$years <= 2022))
[…]

### 2024-04-10 paper/datasheet_template.qmd
Adding datasheet

@@
 ---
-title: "Datasheet for Air quality index by municipality, Leading causes of death
-and Average PM2.5 concentrated data dataset"
+title: "Datasheet for Dataset of Data Breaches and Ransomware Attacks Over 15 Years"
 author: 
-  - Vanshika Vanshika
   - Shivank Goel
-  - Navya Hooda
-thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-abstract: "This datasheet documents the datasets used in a study analyzing the impact of air pollution on health outcomes in Alberta, Canada.
-The data for the study is sourced from the Alberta government's open data portal.
-It spans several years and includes variables like AQHI readings, PM2.5 concentrations, and mortality rates linked to specific health conditions.
-This dataset is helpful for its concise representation to study the environmental health in Alberta.
-It offers insights into public health implications of air pollution.
-The data cleaning and processing done using tools like R, ensures its reliability and accuracy.
-This datasheet is an essential guide to the dataset's structure, collection process, and potential applications."
+thanks: "Code and data are available at: https://github.com/shivankgoel003/DataBreach_Ransomware_Stat."
+date: today
+date-format: long
+abstract: "This datasheet documents the datasets used in a studying data breaches and ransomware attacks over 15 years, from 2004 to 2019.
+The dataset in the study  is compiled by researchers Tsen, Elinor, Ko, Ryan, and Slapnicar, Sergeja, and is based on public data.It serves as a great resource for analyzing trends and patterns in organizational cyber resilience.
+The focus is on the organizational size, industry sector, and impact of cybersecurity strategies, helping us to better understand how businesses deal with digital threats."
 
@@
 
-The dataset was created to investigate the correlation between air quality, particularly PM2.5 levels, and mortality rates due to respiratory and cardiac illnesses in Alberta.
-The aim was to understand how air pollution impacts public health in this region.
+The dataset was created to investigate the cybersecurity incidents impacting various organizations, considering their size and sector.
+It helps to understand the importance of cybersecurity and the need for a strong defense against such threats.
 
@@
 
-The datasets were created by the Alberta Government, specifically by Jobs, Economy and Northern Development, Service Alberta, and Environment and Parks departments.
+The dataset was compiled by Tsen, Elinor, Ko, Ryan, and Slapnicar, Sergeja, affiliated with The University of Queensland.
 
@@
 
-The creation of these datasets was likely funded by the Alberta government. No specific grant details were provided.
+The dataset compilation was part of a PhD project, presumably supported by funding allocated for academic research by The University of Queensland.
+No specific grant number is associated with the dataset creation, as per the available information.
 
@@
 
-These datasets provide critical insights into environmental and public health concerns in Alberta and are vital for ongoing research and policy development in these areas.
+The dataset is a resource for researchers, policymakers, and cybersecurity professionals to study cyber resilience in today's digital world.
 
@@
 
-The instances represent various types of environmental and public health data.
-The first dataset details the Air Quality Health Indices by municipality in Alberta.
-The second dataset provides information on the leading causes of death in Alberta.
-The third dataset offers data on PM2.5 concentrations in Alberta.
+The instances within the dataset represent cybersecurity incidents, especially data breaches and ransomware attacks on various organizations.
+Each instance shows details such as the nature of the attack, the sector and size of the organization affected, the geographical location, and the impact and response measures.
 
@@
 
-The exact number of instances is not specified, but each dataset covers multiple years of data across various locations in Alberta.
+The exact number of cyberattack instances are 1146 of each type withing 15 years of time frame from 2004 to 2020.
 
@@
 
-The datasets seem to be comprehensive for their respective scopes, covering broadly geographical areas within Alberta.
-It seems to be representative of the larger set of environmental and health data within the province.
+The dataset does not cover all possible instances but represents a broad selection from multiple resources.
+It is a sample of large number of cyber incidents, focusing on publicly reported and confirmed attacks.
+It includes a diverse range of incidents from various sectors and countries to provide a representative overview of cyber threats.
 
@@
 
-Each dataset has specific labels or targets.
-For the air quality index, the labels include AQHI values and municipal regions.
-The causes of death dataset labels include types of illnesses and mortality counts.
-The PM2.5 dataset's labels consist of concentration levels, locations, and time periods
+Each instance includes processed data with features such as the year of the attack, the name of the organization, critical industry classification, organization size, level of digital intensity, the sector, country, details of the cyber security role and frameworks, impact level, and actions taken.
+This provides a structured understanding of each incident.
 
@@
 
-Each dataset has specific labels or targets.
-For the air quality index, the labels include AQHI values and municipal regions.
-The causes of death dataset labels include types of illnesses and mortality counts.
-The PM2.5 dataset's labels consist of concentration levels, locations, and time periods.
+Each instance is labeled with the type of attack (e.g., ransomware, data breach) and classified by its impact level (e.g., low, medium, high) based on the response taken by the organization.
 
@@
 
-There is no specific mention of missing information. However, as with most large datasets, there may be gaps due to data unavailability or limitations in data collection.
+There is many missing information in the dataset.This might be due to provacy concerns, or some orgnizations did not report the exact information.
+For example, instances such as number of users affected has lot of gaps in the dataset.
 
@@
 
-The datasets do not explicitly mention relationships between individual instances, but they can be inferred.
-For example, higher AQHI and PM2.5 values may correlate with increased mortality rates from certain diseases.
+The datasets do not explicitly mention relationships between individual instances, however they can be drawn while doing statistical analysis.
 
@@
 
-Specific errors or sources of noise are not mentioned. However, as with any environmental data, there may be variabilities and inconsistencies due to natural fluctuations and measurement limitations.
+Specific errors or sources of noise are not mentioned.
+However, as with any databreach report data, there may be variabilities and inconsistencies due to natural fluctuations and measurement limitations.
 
@@
 
-The datasets are self-contained and available through the Alberta open data portal. They are subject to the Open Government Licence - Alberta, which allows for reuse and distribution.
+The datasets are self-contained and available through the website of University of Queensland.
 
@@
 
-There is no content in these datasets that would be considered offensive or insulting. The data is factual, focusing on environmental and health statistics.
+There is no content in these datasets that would be considered offensive or insulting. The data is factual, focusing on data breaches and cyber incidents.
 
@@
 
-The datasets do not specify sub-populations like age or gender. They are more focused on geographical and temporal aspects of air quality and health outcomes.
+The datasets do not specify sub-populations like age or gender.
 
@@
 
-Individual identification is not possible with these datasets as they deal with aggregated data at the municipal level or higher.
+Individual identification is not possible with these datasets as they deal with aggregated data covering various countries.
 
@@
 
-The datasets do not contain sensitive personal data. They are focused on environmental measures and aggregated health outcomes.
+The datasets do not contain sensitive personal data.
 
@@
 
-The data were directly observed and collected through environmental monitoring stations and health records.
-They are likely verified through standard government data validation processes, though specific details are not provided.
+The data were directly observed and collected through various resources such as : Carnegie Mellon’s List of Banking Cyber Incidents (2020), Repository of Industrial Security Incidents (2015), Privacy Rights ClearingHouse (2020) and Information is Beautiful (2020).
 
@@
 
-The air quality and PM2.5 data were likely collected using environmental monitoring stations equipped with sensors. The causes of death data were probably compiled from health records.
+Not applicable since the dataset was compiled from various resources.
 
@@
 
-The datasets seem to be comprehensive collections rather than samples.
-They likely  represent the entire population of interest (i.e., all municipalities in Alberta for air quality and all recorded causes of death).
+The datasets seem to be collections rather than samples.
 
@@
 
-Data collection was likely conducted by government employees or contractors. Specific details about their compensation are not provided.
+Not applicable, since the dataset was compliation of data from other resouces related to cyber atatcks and cyber security.
 
@@
 
-The data collection covers several years, as indicated by the datasets (e.g., 2001-2021 for the leading causes of death).
+The data collection covers several years, as indicated by the datasets (e.g., 2004-2020 for the data breach incidents).
 
@@
 
-The data were collected from environmental monitoring stations and health records, not directly from individuals. It is a secondary data source.
+Not applicable
 
@@
 
-The datasets have likely been used for academic research and study before. They have been used to analyze the relationship between mortality rates and causes of certain diseases.
+	The dataset has been used in academic studies to analyze and understand cybersecurity incident trends. The author themselves have published an academic paper based on the dataset.
 
@@
 
-No specific repository is mentioned, but the datasets are available through the Alberta open data portal.
+The University of Queensland’s eSpace repository contains links to publications and related datasets. The dataset and related works can be accessed at UQ eSpace.
 
@@
 
-The data could be used for further environmental health studies, public health planning, and educational purposes.
+The data could be used for further statistical analysis in the field of cyber risks and cybersecurity.It may also be used for educational purposes, to train machine learning models, or for benchmarking organizational cybersecurity practices.
 
@@
 
-The users should consider potential biases and reporting in data collection.Also they should be aware of geographic
-coverage to avoid misinterpretations.
+The users should consider potential biases and reporting in data collection.Also they should be aware of geographic coverage to avoid misinterpretations.
 
@@
 
-The data should  not be used for identifying or targeting specific individuals.
+The dataset should not be used for tasks that could compromise the privacy or security of the affected organizations.
 
@@
 
-The datasets are valuable for understanding trends and correlations in environmental and public health. They should be used appropriately, considering their scope and limitations.
+The datasets are valuable for understanding trends and correlations in cyber incidents. They should be used appropriately, considering their scope and limitations.
 
@@
 
-The datasets are publicly available and can be accessed by third parties. They are distributed under the Open Government Licence - Alberta, which allows for use and distribution by anyone.
+The dataset is already in the public domain and is available to third parties for research and analysis. It can be accessed globally, subject to the terms of use set by the University of Queensland.
 
@@
 
-The datasets are available for download through the Alberta open data portal. There is no mention of a DOI or distribution through other platforms like GitHub.
+The dataset is distributed online through the University of Queensland's data collection portal. It does have a DOI, which can be used to cite the dataset in academic work.
 
@@
 
-The datasets are already available for public access.
+The dataset is already available and has been since its publication year in 2020.
 
@@
 
-The datasets are distributed under the Open Government Licence - Alberta, which permits a wide range of uses. The license details can be found on the Alberta open data portal.
+Yes, the dataset is distributed under a license that permits reuse with acknowledgment, and the terms are outlined in the dataset’s access conditions on the University of Queensland’s website.
 
@@
 
-There are no specified export controls or regulatory restrictions beyond the Open Government Licence - Alberta.
+There are no export controls or regulatory restrictions mentioned for this dataset.
 
@@
 
-The datasets are maintained by the respective departments of the Alberta government – Jobs, Economy and Northern Development, Service Alberta, and Environment and Parks.
+The dataset will continue to be hosted and maintained by the University of Queensland, ensuring its accessibility for ongoing and future research.
 
@@
 
-Contact information is likely available through the Alberta government’s open data portal.
+Enquiries about the dataset can be directed to the corresponding author of the research or the data management team at the University of Queensland.
+Contact can typically be made through the contact details provided on the dataset's webpage.
 
@@
 
-The datasets are updated annually. Specific communication channels for updates are not mentioned.
+Updates to the dataset, if any, will be managed by the research team at the University of Queensland.
+They would be responsible for correcting any errors, adding new data, or removing outdated instances.
 

### 2024-04-10 paper/paper.qmd
Adding datasheet

@@
 abstract: "Undoubtedly as technology is progressing and we are stepping more closer to the digital world, the digital threats are also increasingly rapidly.
-This studys discusses the influence of organizational size, sector, and cybersecurity strategies on cyberattack impacts.
+This study discusses the influence of organizational size, sector, and cybersecurity strategies on cyberattack impacts.
 Utilizing a dataset from the University of Queensland about data breaches and ransomware attacks from 2004 to 2019, we analyze the trends and correlations and study how different factors contribute to the severity and frequency of these incidents.
 Our analysis reveals key insights: larger organizations report more cyber incidents, sectors like healthcare and education are frequent targets, and cybersecurity frameworks significantly help in reducing the user impact.
-These findings highlight critical areas for how organizations can improve their security and protect their data.
+These findings highlight critical areas for how organizations can improve their security and protect their data."
 header-includes:
